
---
title: "Farmer_list_matching"
output: html_document
date: "2023-07-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the personal information file of the PDC 1 from the case.

```{r}

baseline <- read_excel("pi_syngenta_tomatoes.xlsx")
baseline$identifier <- gsub("-", "", baseline$identifier)

full_data <- read_excel("kenya-syngenta_tomatoes.xlsx")
full_data <- full_data %>% select(c("identifier", "hh_farmer_gender"))

baseline <- baseline %>% left_join( full_data, by="identifier")
# remove empty row
baseline <- baseline %>%
  filter(identifier != "nh3dyaxd8p86")

# Remove duplicates from baseline based on mobile_number_farmer
baseline <- baseline %>%
  distinct(mobile_number_farmer, .keep_all = TRUE)
```

```{r}
table(baseline$hh_farmer_gender, baseline$pi_location_cascade_county)
```

Import the available farmer list for endline.

```{r, include = FALSE}
library(readxl)
library(dplyr)
farmer_list_endline_kirinyaga_1 <- read_excel("farmer_list_endline_syngenta.xlsx", sheet = 2)
farmer_list_endline_kirinyaga_1$location <- "Kirinyaga"
farmer_list_endline_kirinyaga_1$service <- "Loan"
farmer_list_endline_kirinyaga_1$`Year of Engagement` <- NA

farmer_list_endline_kirinyaga_2 <- read_excel("farmer_list_endline_syngenta.xlsx", sheet = 5)
farmer_list_endline_kirinyaga_2$location <- "Kirinyaga"
farmer_list_endline_kirinyaga_2$service <- "Training only"

farmer_list_endline_embu_1 <- read_excel("farmer_list_endline_syngenta.xlsx", sheet = 3)
farmer_list_endline_embu_1$location <- "Embu"
farmer_list_endline_embu_1$service <- "Loan"
farmer_list_endline_embu_1$`Year of Engagement` <- NA

farmer_list_endline_embu_2 <- read_excel("farmer_list_endline_syngenta.xlsx", sheet = 6)
farmer_list_endline_embu_2$location <- "Embu"
farmer_list_endline_embu_2$service <- "Training only"

farmer_list_endline <- rbind(farmer_list_endline_kirinyaga_1, farmer_list_endline_kirinyaga_2, farmer_list_endline_embu_1, farmer_list_endline_embu_2)

farmer_list_endline$`Mobile Number`<- paste0("0", farmer_list_endline$`Mobile Number`)

# remove duplicates from farmer_list_endline based on mobile_number_farmer
farmer_list_endline <- farmer_list_endline %>%
  distinct(`Mobile Number`, .keep_all = TRUE)
```

```{r}
table(farmer_list_endline$location, farmer_list_endline$service)
```


```{r, include = FALSE}
# Rename the Mobile number column to mobile_number_farmer
colnames(farmer_list_endline)[colnames(farmer_list_endline) == "Mobile Number"] <- "mobile_number_farmer"

```

## Matching 

We can try and match farmers with a left join function


```{r}
merged_data <- baseline %>% left_join(farmer_list_endline, by="mobile_number_farmer")
```

```{r}
# create now column to indicate whether there is a match or not 
merged_data$match <- ifelse(is.na(merged_data$Ward), FALSE, TRUE)
```

```{r}
table(merged_data$match, merged_data$service, merged_data$pi_location_cascade_county)
```



## Match to retrieve sample = starting from farmer list 

```{r}
# Merge baseline on mobile number 
farmer_list_endline <- farmer_list_endline %>% left_join(baseline, by = "mobile_number_farmer")

farmer_list_endline <- farmer_list_endline %>% select(-c("pi_location_cascade_county", "pi_location_cascade_sub_county","pi_location_other", "pi_location_other_first_admin", "pi_location_other_second_admin", "pi_location_other_third_admin", "pi_location_other_village", "pi_geolocation_latitude", "pi_geolocation_longitude", "pi_geolocation_elevation","hh_farmer_gender", "name_of_farmer"))

farmer_list_endline$match <- ifelse(is.na(farmer_list_endline$identifier), FALSE, TRUE)
colnames(farmer_list_endline)[colnames(farmer_list_endline) == "identifier"] <- "match_phone"

```

Make the sample

```{r}
sample <- farmer_list_endline[farmer_list_endline$match == TRUE,]

```

Make random sample of 31 farmers Embu getting loan.

```{r}
embu_loan <- farmer_list_endline[(farmer_list_endline$location == "Embu" & farmer_list_endline$service == "Loan" & farmer_list_endline$match == FALSE),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 31

# Get the total number of rows in the dataframe
total_rows <- nrow(embu_loan)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
embu_loan_sample <- embu_loan[random_row_indices, ]
```

Make random sample of 37 farmers Embu getting only training.

```{r}
embu_training <- farmer_list_endline[(farmer_list_endline$location == "Embu" & farmer_list_endline$service == "Training only" & farmer_list_endline$match == FALSE),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 37

# Get the total number of rows in the dataframe
total_rows <- nrow(embu_training)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
embu_training_sample <- embu_training[random_row_indices, ]
```

Make random sample of 80 farmers from Kirinyaga receiving loan.

```{r}
kirinyaga_loan <- farmer_list_endline[(farmer_list_endline$location == "Kirinyaga" & farmer_list_endline$service == "Loan" & farmer_list_endline$match == FALSE),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 80

# Get the total number of rows in the dataframe
total_rows <- nrow(kirinyaga_loan)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
kirinyaga_loan_sample <- kirinyaga_loan[random_row_indices, ]
```

Make random sample of 77 farmers from Kirinyaga receiving training only

```{r}
kirinyaga_training <- farmer_list_endline[(farmer_list_endline$location == "Kirinyaga" & farmer_list_endline$service == "Training only" & farmer_list_endline$match == FALSE),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 77

# Get the total number of rows in the dataframe
total_rows <- nrow(kirinyaga_training)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
kirinyaga_training_sample <- kirinyaga_training[random_row_indices, ]
```

Merge the different random sets with the recurring farmers 

```{r}
sample_tomato <- rbind(sample, embu_loan_sample, embu_training_sample, kirinyaga_loan_sample, kirinyaga_training_sample)
```

Export the full farmer list and the sample.

```{r}
library(openxlsx)
write.xlsx(farmer_list_endline, "Farmer_list_syngenta_tomato_endline.xlsx")
write.xlsx(sample_tomato, "sample_syngenta_tomato_endline.xlsx")
```

