---
title: "Data_clearning_Tanzania_endline"
output: html_document
date: "2023-08-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is developed to clean data from Farmfit cases in a standardized way.

### Preparation

Load the necessary packages.

```{r, include = FALSE}
library(here); library(readxl); library(openxlsx);
library(plyr); library(dplyr); library(tidyr); library(tidylog); library(tidyselect)
library(stringr); library(data.table); library(reshape2)
library(zoo);library(splitstackshape) ;library(tidyverse); library(magrittr)
```

Fill in the basic case information.

```{r}
case = "USSL"
year = "2023"
sdm_crop = "maize"
country = "Tanzania"
```

Make a new Github map with the project name where this R Markdown file is stored, together with the raw data and the survey form downloaded from Flow. Fill in the right file names in the code below so the variables data_filename and survey_filename refer to the correct file.

```{r}
data_filename <- "DATA_ANALYSIS-Tz-endline.xlsx"
survey_filename <- "SURVEY_FORM-Tz-endline.xlsx"
```

Describe the files that will be created in the script here.

```{r}
pi_filename <- paste("pi_",year,"_",case,"_",sdm_crop,".xlsx", sep="")
data_delivery <- paste(year,"_",case,"_anom_",sdm_crop,"_endline",".xlsx", sep = "")
```

Import the data files that can be found in the map Analytics-Guild -\> IDH -\> Templates and input files for data delivery. Copy them in the project map.

```{r}
question_library <- read_excel("question library format v4.1.1.xlsx",
                               sheet = "Full Survey", skip = 2) 
vars_transformed <- read_excel("variables with transformation_2.xlsx") 
household_demographics  <- read_excel("household_demographics_var_2.xlsx") 
vars_dashboard <- read_excel("variable names input dashboard_2.xlsx") 

```

Define functions that will be needed later in the script.

```{r}
# DATA TYPE Set factors to integers
factor_to_int <- function(x){
  as.numeric(as.character(x))
}
# Find and replace outliers
outlier_detection <- function(x){
  ifelse(
    x > (mean(x, na.rm=TRUE) + sd(x, na.rm=TRUE)*3) |
      x < (mean(x, na.rm=TRUE) - sd(x, na.rm=TRUE)*3), 
    9997,
    x
  )
}
options(scipen = 999)  
# count values without NA
count_n <- function(x){sum(!is.na(x))}
```

### Import data

We read data from Flow twice. On Data we will perform cleaning. Data_raw will be added to the data delivery as the full raw data file without transformations. Only the personal information variables are removed.

```{r, include = FALSE}
# Data for transformation
Data <- read_excel(data_filename)

# Raw data 
Data_raw <- read_excel(data_filename)
Data_raw <- Data_raw %>%
  select(-starts_with("pi_"), -name_of_farmer, -mobile_number_farmer)

# Read file with survey structure
survey_questions <- read_excel((survey_filename),sheet ="Full Survey", skip=2)
```

Adjust the survey file to be ready for analysis. We now have a clear overview of the survey questions asked in this case.

```{r}
survey_questions %<>% 
  # Fill excel merged cells from the section
  mutate(Title = na.locf(Title, na.rm = FALSE)) %>%
  # Select right columns
  select("Title", "Variable name", "Text",
         "Question type", "Options",
         "Allow multiple") %>%
  # Trim trailing spaces
  mutate_if(is.character, str_trim) %>%
  # Rename columns
  rename("section" = "Title", 
         "variable" = "Variable name", 
         "question" = "Text",
         "type" = "Question type", 
         "options" = "Options",
         "multiple" = "Allow multiple") %>%
  # All variables to lowercase
  mutate_all(tolower) %>%
  # Filter the empty variables
  filter(!is.na(variable)) 
```

What are the numerical columns? How many participants?

```{r}
# Collect the different variables, numerical questions:
numerical_columns <- survey_questions %>%
  filter(type == "number") %>%
  select("variable") %>% pull()

## ---- Count initial number of participants
nr_participants_raw <- length(unique(Data$Identifier))
```

## Check how it compares with baseline sample

```{r}

#import baseline data
baseline <- read_excel("Baseline data/pi_ussl_tanzania.xlsx")
baseline$identifier <- gsub("-", "", baseline$identifier)

full_data <- read_excel("Baseline data/2021_tanzania-ussl.xlsx")
full_data <- full_data %>% select(c("identifier","pi_location_cascade_region", "hh_farmer_gender"))

baseline <- baseline %>% left_join( full_data, by="identifier")

```

```{r}
# Some data cleaning in baseline 
baseline <- baseline %>% mutate(mobile_number_farmer = ifelse(identifier == "n9jr-aa5a-68kq", "0786909569", mobile_number_farmer))
baseline <- baseline %>% mutate(mobile_number_farmer = ifelse(identifier == "75wr-y84m-f6wj", "0785664956", mobile_number_farmer))
baseline$mobile_number_farmer <- gsub("-", "", baseline$mobile_number_farmer)
baseline <- baseline %>%drop_na(mobile_number_farmer)


```

Compare with endline data

```{r}
result <- baseline %>%
  inner_join(Data, by = "mobile_number_farmer") %>%
  summarise(equal_observations = n())
```

### Changes to dataset and codebook

First quick data cleaning steps:

-   Date format

-   Remove irrelevant columns

-   Names to lower case

-   Remove farmers that do not want to participate

-   Data types

-   "other" columns

```{r, include = FALSE}
Data <- Data %>%
  
  # Set the date to date format
  mutate(`Submission Date` = as.Date(`Submission Date`, format = "%d-%m-%Y")) %>%
  
  # Remove irrelevant columns
  # ONLY DELETE SUBMITTER AFTER THE PRODUCTIVITY CHECK! YOU WANT THIS VARIABLE IN FOR OULIER CHECKING
  select(-c(`Display Name`,`Device identifier`,`Instance`,
            #`Submitter`, 
            `Form version`)) %>%
  select(-contains("--option--")) %>%
  
  # All column names to lowercase
  rename_all(funs(tolower)) %>%
  
  # Rename option variables
  rename_all( funs(gsub("--other--", "_other2", ., perl=T))) %>%

  # All variables to lowercase
  mutate_all(tolower) %>%
  
  # Remove to farmers that didn't participate
  mutate(ic_informed_consent = na.locf(ic_informed_consent, na.rm = FALSE)) %>%
  filter(ic_informed_consent == "accepted to participate")  %>%
  
  # Numerical columns to numeric data type
  mutate_at(vars(numerical_columns), funs(as.numeric)) %>%
    
    #General case
    mutate(f_harvest_num = as.numeric(f_harvest_num)) %>%
    ## In some cases, the crop has been added to this variable name. Variable will then be f_[CROP]_harvest_num. This occurs when there are 2 focus crops in the survey. Adjustment is required in this case, like in the example below for a cocoa case. 
    #mutate(f_cocoa_harvest_num = as.numeric(f_cocoa_harvest_num)) %>% 
  
  # Remove all "other" option text ("other, please specify")
  # mutate_each(funs(str_remove(., "other, please specify"))) %>%
  mutate_if(is.character, funs(gsub("other, please specify", NA,.))) %>% 
  mutate_if(is.character, funs(gsub("\\|$","",.))) %>% 
  
  # Remove trailing spaces
  mutate_if(is.character, str_trim) %>% 
  
  # Remove punctuation and special characters - EXCEPT for "|"
  mutate_if(is.character, funs(gsub("[^\\|[:^punct:]]", "", ., perl=T)))
```

Adjust the **farm sizes** to acres.

```{r}
# First check what are the unit measurements reported in this case.
table(Data$f_unit_land)
table(Data$f_unit_land_other2)
```

The code below transforms hectares to acres. When another option is given as unit of measurement, the code needs to be adjusted so that everything is in acres in the variables f_size_acre and f_focus_crop_size_acre.

```{r, include = FALSE}

Data <- Data %>%
  # cleaning wrong value of f_focus_crop_size. Should be 5  instead of 48
  mutate(f_focus_crop_size = ifelse(identifier == "4fvmuv7dnks2", 5 , f_focus_crop_size))

# Mutate the different units to acres. We need a farm size in acres for the total farm size, and the size dedicated to focus crop.
Data <- Data %>%
# Combine unit columns
  unite("f_unit_land", c(f_unit_land, f_unit_land_other2), 
        na.rm = TRUE, remove = FALSE, sep=" ") %>%
  # Convert land size to acres if in hectares
  mutate(f_size_acre = ifelse(f_unit_land == "hectares", f_size, f_size * 2.471)) %>%
  mutate(f_focus_crop_size_acre = ifelse(f_unit_land == "hectares", f_focus_crop_size, f_focus_crop_size * 2.471)) %>%
  # Fill missing values in focus crop size
  mutate(f_focus_crop_size_acre = na.locf(f_focus_crop_size_acre, na.rm = FALSE)) %>%
  # Calculate size in hectares or acres based on unit
  mutate(f_size_hectare = ifelse(f_unit_land == "hectares", f_size, f_size_acre / 2.471)) %>%
  mutate(f_focus_crop_size_hectare = ifelse(f_unit_land == "hectares", f_focus_crop_size, f_focus_crop_size_acre / 2.471))
```

Mutate the birth year variables to **age.** Check whether they are included in this case or not.

```{r, include = FALSE}
Data <- Data %>%
  mutate(cal_hh_farmer_age = 2023 - hh_farmer_birthyear,
           cal_hh_head_age = 2023-hh_head_birthyear) #update to current year, check whether applicable, not always included.
```

### Other variables

Identify the "other" variables, and combine the values with their parent. Then the "other" variables are removed.

```{r, include = FALSE}
variables_other <- Data %>% 
  select(ends_with("_other"), ends_with("_other2")) %>% 
  names() 

for(other in variables_other){
  
  if(any(names(Data) %in% gsub("_other$|_other2$", "", other))){
    
    # Fuzzy match? - many spelling mistakes in "other" questions
    
    # Combine "other" variables with their parents
    Data <- Data %>%
      unite(!!gsub("_other$|_other2$", "", other), 
            c(gsub("_other$|_other2$", "", other), all_of(other)), 
            sep="|", remove=FALSE, na.rm=TRUE) %>%
      select(-other)
  }
}
```

### 

### Context specific adjustments

Check for obvious outliers first by running the full code EXCEPT for running the outlier function below. Download the productivity_check.xlsx. file that is created at the end of the script. Use the filter options and arrange columns from high to low, low to high, etc. to detect outliers manually. Also use the intake form! Highlight the cells you have questions about or that are obvious outliers. Consult the data collection supervisors to make adjustments. The changes in the data can be made in the code block below.

Note that you should keep the variable "submitter" for the time being so we can trace back which enumerator has entried that data. This variable is normally deleted in the script above (cmd+F "submitter"). Make sure you delete this variable in the end when sharing the data.

Farm size

```{r}

Data <- Data %>%
  mutate(f_focus_crop_size = ifelse(identifier == "4fvmuv7dnks2", 5, f_focus_crop_size))
```

Price

```{r}
# Price

Data <- Data %>% mutate(f_focus_price = ifelse(identifier == "rqjag5qbh6pp", f_focus_price / 10, f_focus_price))


median_price <- Data %>%
  filter(f_focus_measurement_sold == "100 kg bags") %>%
  summarize(median_price = median(f_focus_price, na.rm = TRUE))

Data <- Data %>%
  mutate(
    f_focus_price = case_when(
      identifier %in% c("7hhj95g8v6xc", "b6he2p6uucb1", "pumunnkrsw4e") ~ median_price$median_price,
      identifier == "hg01f007577k" ~ 85000,
      identifier == "e15k5fgnu2m1" ~ 85000,
      identifier == "cw9gwcdajwc7" ~ 90000,
      TRUE ~ f_focus_price
    )
  )

```

Clean focus crop measurement unit produced

```{r}

Data <- Data %>%
  mutate(
    f_focus_measurement_prod = case_when(
      f_focus_measurement_prod %in% c("", "NA") & f_focus_measurement_sold == "100 kg bags" ~ "100 kg bags",
      f_focus_measurement_prod == "108kg" ~ "108 kg bags",
      f_focus_measurement_prod == "120kg bags" ~ "120 kg bags",
      identifier %in% c("p8mxbt8pdspd", "mnpy971bpgt5", "bcg1q3qp5k27", "5k5q4xc6b061") ~ "100 kg bags",
      TRUE ~ f_focus_measurement_prod
    ),
    f_focus_quant_prod = ifelse(identifier == "b6he2p6uucb1", f_focus_quant_prod / 108, f_focus_quant_prod),
    f_focus_measurement_prod = ifelse(f_focus_measurement_prod == "120 kg of bags", "120 kg bags", f_focus_measurement_prod),
    f_focus_measurement_prod = ifelse(identifier == "495mytja41t6", "", f_focus_measurement_prod),
    f_focus_measurement_sold = ifelse(identifier == "495mytja41t6", "", f_focus_measurement_sold),
    f_focus_measurement_prod = ifelse(identifier == "5k5q4xc6b061", "100 kg bags", f_focus_measurement_prod),
    f_focus_price = ifelse(identifier == "5k5q4xc6b061", median_price$median_price, f_focus_price),
    f_focus_measurement_prod = ifelse(
      (f_focus_price %in% c(60000, 70000, 80000, 85000, 90000, 95000)) & grepl("gunia", f_focus_measurement_prod),
      "100 kg bags",
      f_focus_measurement_prod
    ),
    f_focus_measurement_prod = case_when(
      f_focus_measurement_prod %in% c("120", "120 kg if bags", "120 kg of bags ", "120 kg of bags") ~ "120 kg bags",
      f_focus_measurement_prod == "6bags" ~ "",
      TRUE ~ f_focus_measurement_prod
    )
  )


```

Clean focus crop measurement unit sold

```{r}

# Load the necessary library
library(dplyr)

# Clean the data
Data <- Data %>%
  mutate(
    f_focus_measurement_sold = case_when(
      f_focus_quant_sold == 0 & !(f_focus_measurement_sold %in% c("100 kg bags", "108kg", "i don't know", "i prefer not to say")) ~ "0, unknown or not applicable",
      f_focus_measurement_sold %in% c("120", "120kg bags", "120 kg if bags", "120 kg of bags", "120kg bags") ~ "120 kg bags",
      f_focus_measurement_sold %in% c("108 kg", "108kg") ~ "108 kg bags",
      (f_focus_price %in% c(60000, 70000, 80000, 85000, 90000, 95000)) & grepl("gunia", f_focus_measurement_sold) ~ "100 kg bags",
      identifier == "8svhjs4vq22r" ~ "108 kg bags",
      identifier == "kh8jhf0kk6n1" ~ "100 kg bags",
      identifier == "835381j3wd9g" ~ "100 kg bags",
      identifier == "5k5q4xc6b061" ~ "100 kg bags",
      identifier == "495mytja41t6" ~ "",
      identifier == "197cu3yn9x3t" ~ "100 kg bags",
      identifier == "rqjag5qbh6pp" ~ "100 kg bags",
      TRUE ~ f_focus_measurement_sold
    ),
    f_focus_measurement_prod = case_when(
      identifier == "rqjag5qbh6pp" ~ "100 kg bags",
      TRUE ~ f_focus_measurement_prod
    )
  )

```

Consumption share

```{r}
library(dplyr)

Data <- Data %>%
  mutate(
    f_focus_own_consumption_share = case_when(
      f_focus_own_consumption_share == "2650 between a quarter and half of it" ~ "26-50%",
      f_focus_own_consumption_share == "025a small part" ~ "0-25%",
      f_focus_own_consumption_share == "100 all of it" ~ "100%",
      f_focus_own_consumption_share == "5175 more than half but less than 75" ~ "51-75%",
      f_focus_own_consumption_share == "75100 most of it" ~ "76-100%",
      TRUE ~ f_focus_own_consumption_share
    )
  )

Data <- Data %>%
  mutate(
    f_focus_own_consumption_share_v2 = ifelse(
      !grepl("[^0-9.]", f_focus_own_consumption_share),
      as.numeric(f_focus_own_consumption_share),
      NA
    )
  ) %>%
  mutate(
    f_focus_own_consumption_share_v2 = f_focus_own_consumption_share_v2 / f_focus_quant_prod
  ) %>%
  mutate(
    f_focus_own_consumption_share = case_when(
      f_focus_own_consumption_share_v2 >= 0 & f_focus_own_consumption_share_v2 <= 0.25 ~ "0-25%",
      f_focus_own_consumption_share_v2 > 0.25 & f_focus_own_consumption_share_v2 <= 0.5 ~ "26-50%",
      f_focus_own_consumption_share_v2 > 0.5 & f_focus_own_consumption_share_v2 <= 0.75 ~ "51-75%",
      f_focus_own_consumption_share_v2 > 0.75 & f_focus_own_consumption_share_v2 < 0.999 ~ "76-100%",
      f_focus_own_consumption_share_v2 == 1 ~ "100%",
      TRUE ~ f_focus_own_consumption_share
    )
  ) %>%
 select(-f_focus_own_consumption_share_v2)



```

Lost share

```{r}
library(dplyr)

Data <- Data %>%
  mutate(
    f_focus_lost_share = case_when(
      f_focus_lost_share == "2650 between a quarter and half of it" ~ "26-50%",
      f_focus_lost_share == "025a small part" ~ "0-25%",
      f_focus_lost_share == "100 all of it" ~ "100%",
      f_focus_lost_share == "5175 more than half but less than 75" ~ "51-75%",
      f_focus_lost_share == "75100 most of it" ~ "76-100%",
      TRUE ~ f_focus_lost_share
    )
  )

Data <- Data %>%
  mutate(
    f_focus_lost_share_v2 = ifelse(
      !grepl("[^0-9.]", f_focus_lost_share),
      as.numeric(f_focus_lost_share),
      NA
    )
  ) %>%
  mutate(
    f_focus_lost_share_v2 = f_focus_lost_share_v2 / f_focus_quant_prod
  ) %>%
  mutate(
    f_focus_lost_share = case_when(
      f_focus_lost_share_v2 >= 0 & f_focus_lost_share_v2 <= 0.25 ~ "0-25%",
      f_focus_lost_share_v2 > 0.25 & f_focus_lost_share_v2 <= 0.5 ~ "26-50%",
      f_focus_lost_share_v2 > 0.5 & f_focus_lost_share_v2 <= 0.75 ~ "51-75%",
      f_focus_lost_share_v2 > 0.75 & f_focus_lost_share_v2 < 0.999 ~ "76-100%",
      f_focus_lost_share_v2 == 1 ~ "100%",
      TRUE ~ f_focus_lost_share
    )
  ) %>%
 select(-f_focus_lost_share_v2)




```

Dependencies

```{r}
Data <- Data %>%
  mutate(
    cs_access_finance_change_linkage_sdm = ifelse(grepl("decrease", cs_access_finance_change, ignore.case = TRUE), "", cs_access_finance_change_linkage_sdm),
    cs_climate_adaptation_change_linkage_sdm = ifelse(grepl("decreased", cs_climate_adaptation_change, ignore.case = TRUE), "", cs_climate_adaptation_change_linkage_sdm),
    cs_crop_income_change_linkage_sdm = ifelse(grepl("decrease", cs_crop_income_change, ignore.case = TRUE), "", cs_crop_income_change_linkage_sdm),
    cs_crop_costs_change_linkage_sdm = ifelse(grepl("increase", cs_crop_costs_change, ignore.case = TRUE), "", cs_crop_costs_change_linkage_sdm),
    cs_crop_loss_change_linkage_sdm = ifelse(grepl("increase", cs_crop_loss_change, ignore.case = TRUE), "", cs_crop_loss_change_linkage_sdm)
  )

```

### Handle outliers.

```{r, include = FALSE}

# OUTLIERS:
#Identify numerical columns again because of change in names above
numerical_columns <- survey_questions %>%
filter(type == "number") %>%
select("variable") %>% pull()

```

Only run the next chunck of code AFTER you did the productivity check! See explanation above.

```{r, include = FALSE}
Data <- Data %>%
    mutate_at(vars(numerical_columns), funs(outlier_detection))
```

Put values that are 9999/"i don't know" or 9998/"i prefer not to say/9997 to NA.

```{r, include=FALSE}
Data <- Data %>%
    mutate_if(is.numeric, list(~na_if(., 9999))) %>%
    mutate_if(is.numeric, list(~na_if(., 9998))) %>%
    mutate_if(is.numeric, list(~na_if(., 9997))) %>%
    mutate_if(is.character, list(~na_if(., "i don't know"))) %>%
    mutate_if(is.character, list(~na_if(., "i prefer not to say"))) 
```

### Some statistics

```{r}
# Number of participants who gave informed consent 
nr_participants_ic <- length(unique(Data$identifier))

# NUMBER OF SEASONS
number_of_seasons <- max(Data$f_harvest_num, na.rm=TRUE)
```

# Actual income calculation

### 1) Transform unit measurements

Each unit measurement variables needs to be aligned. We transform every variable to kg.

First check what are the unit measurement reported in this case. Use the code below, or check the Lumen dashboard.

```{r}
table(Data$f_focus_measurement_prod)
table(Data$f_focus_measurement_sold)
```

To do so, we extract the numerical values from the measurement units for quantity produced, sold, consumed and lost. Always check by hand whether this is going well. We extract the numbers from the measurement units to calculate total quantities. If there are other measurement units, not related to kg, put a value for those in the variable: cal_focus_measurement_prod

**Measurement unit production**

```{r, include = FALSE}
Data <- Data %>%
##Measurement unit for production
  mutate(cal_focus_measurement_prod := ifelse(grepl("[[:digit:]]", f_focus_measurement_prod), 
      extract_numeric(f_focus_measurement_prod), 
      ifelse(f_focus_measurement_prod %in% c("KG","kg","kilo","kgs","kilogram", "kilograms kg"), 1, 0)))
```

**Measurement unit sales**

```{r, include = FALSE}
Data <- Data %>%
##Measurement unit for sales focus crop
  mutate(cal_focus_measurement_sold := ifelse(
      grepl("[[:digit:]]", f_focus_measurement_sold), 
      extract_numeric(f_focus_measurement_sold), 
      ifelse(f_focus_measurement_sold %in% c("KG","kg","kilo","kgs","kilogram", "kilograms kg"), 1, 0)))

```

### 2) Quantities produced, sold, lost and own consumption

In this section the quantities produced, sold, lost and own consumption are cleaned.

For quantities produced and sold, we build further on the calculated measurement units in previous section.

```{r, include = FALSE}
##Calculate the quantities focus crop
Data <- Data %>%
  mutate(cal_focus_quant_prod_kg = f_focus_quant_prod * cal_focus_measurement_prod) %>%
  mutate(cal_focus_quant_sold_kg = f_focus_quant_sold * cal_focus_measurement_sold) 
  
  # also calculate the unit price 
Data <- Data %>%
  mutate(cal_focus_price = ifelse(cal_focus_measurement_prod != 0, f_focus_price/cal_focus_measurement_sold, f_focus_price),
    cal_focus_price = ifelse(is.na(cal_focus_quant_sold_kg) | cal_focus_quant_sold_kg == 0, NA, cal_focus_price))

```

The quantities lost and own consumption of the focus crop, we evaluated in buckets:

-   0% (none of it)

-   0-25% (a small part)

-   26-50% (Between a quarter and half of it)

-   51-75% (More than half but less than 75%)

-   75-100% (Most of it)

-   100% (all of it)

We calculate the actual quantities by using the codes belonging to the options in this question and the quantity produced.

```{r}
table(Data$f_focus_lost_share)
```

```{r}
#Data <- Data %>%
#  mutate(cal_focus_quant_own_consumption_kg = f_focus_own_consumption_share * cal_focus_quant_prod_kg) %>%
#  mutate(cal_focus_quant_lost_kg = f_focus_lost_share * cal_focus_quant_prod_kg)
```

Check whether the calculations went well. We make a subset of the data with relevant variables. Click on the dataset to sort the columns to detect strange numbers and solve if necessary.

```{r, include = FALSE}

quantities_produced_subset <- Data %>% select(f_focus_measurement_prod, cal_focus_measurement_prod, f_focus_quant_prod, cal_focus_quant_prod_kg)

quantities_sold_subset <- Data %>% select(f_focus_measurement_sold, cal_focus_measurement_sold, f_focus_quant_sold, cal_focus_quant_sold_kg)

price_subset <- Data %>% select(f_focus_price, cal_focus_measurement_sold, cal_focus_price)
```

### 3) Focus crop revenue

```{r, include = FALSE}
Data <- Data %>%
  mutate(cal_focus_revenue = cal_focus_quant_sold_kg * cal_focus_price) 
   
##X) STANDARD APPROACH Inbetween, adjust for repeated question groups (IF APPLICABLE)  ----  
     ##Adress values for those that have multiple seasons

Data <- Data %>%
  #Identify duplicates
  group_by(identifier, f_focus_rev_timeperiod) %>% 
  mutate(count_dup = n()) %>% # Count duplicates within each group
  mutate(count_dup = ifelse(is.na(f_focus_rev_timeperiod), NA, count_dup)) %>% # Check whether we don't make mistakes here. For some cases, farmers enter twice wrongly, while in other cases entering the same values for 2 seasons is ok.
  
  #Identify those with multiple seasons
  group_by(identifier, f_focus_rev_timeperiod) %>% 
  mutate(count_seasons = n()) %>% # Count the number of seasons for each group
  mutate(count_seasons = ifelse(is.na(f_focus_rev_timeperiod) | count_seasons > 1, NA, count_seasons)) %>%

  #Replace values with 0 for those that entered the data twice 
  mutate(`repeat no` = as.numeric(`repeat no`)) %>%
  mutate(cal_focus_quant_prod_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_prod_kg)) %>%
  mutate(cal_focus_quant_sold_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_sold_kg)) %>%
  #mutate(cal_focus_quant_own_consumption_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_own_consumption_kg )) %>%
  #mutate(cal_focus_quant_lost_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_lost_kg)) %>%
  mutate(cal_focus_revenue = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_revenue)) %>%

  #Sum the quantities for those that entered data for multiple seasons
  group_by(identifier, count_seasons) %>%
  mutate(cal_focus_quant_prod_kg =  sum(cal_focus_quant_prod_kg)) %>%
  mutate(cal_focus_quant_sold_kg =  sum(cal_focus_quant_sold_kg)) %>%
  #mutate(cal_focus_quant_own_consumption_kg =  sum(cal_focus_quant_own_consumption_kg)) %>%
  #mutate(cal_focus_quant_lost_kg =  sum(cal_focus_quant_lost_kg)) %>%
  mutate(cal_focus_revenue = sum(cal_focus_revenue)) %>%

  #Ensure we have 1 row of data for each farmer
  mutate(cal_focus_quant_prod_kg = ifelse(`repeat no` >1 , NA, cal_focus_quant_prod_kg)) %>%
  mutate(cal_focus_quant_sold_kg = ifelse(`repeat no` >1  , NA, cal_focus_quant_sold_kg)) %>%
  #mutate(cal_focus_quant_own_consumption_kg = ifelse(`repeat no` >1  , NA, cal_focus_quant_own_consumption_kg )) %>%
  #mutate(cal_focus_quant_lost_kg = ifelse(`repeat no` >1 , NA, cal_focus_quant_lost_kg)) %>%
  mutate(cal_focus_revenue = ifelse (`repeat no`>1, NA, cal_focus_revenue)) %>%
  ungroup() %>%
  select(-count_dup, -count_seasons)

##Replace missing values in focus revenue by zero
Data <- Data %>%
  mutate(cal_focus_revenue = ifelse(is.na(cal_focus_revenue), 0, cal_focus_revenue)) 

```

### 4) Productivity

Calculate the productivity variable by dividing the production by the farm size dedicated to the focus crop.

```{r, include = FALSE}

Data <- Data %>%
  mutate(cal_focus_productivity_acre = cal_focus_quant_prod_kg/f_focus_crop_size_acre) 
```

### 5) Labour costs

There are different approaches to chose from in terms of the labour cost questions. Check which approach applies for the case and delete the others.

```{r}
# APPROACH 1: Labour cost per month 
# First create a new variable month_count to indicate in how many months labour was hired

#Data <- Data %>%
#  mutate(
#    month_count = case_when(
#      f_focus_labour_hired_months == "all months" ~ 12,
#      TRUE ~ str_count(f_focus_labour_hired_months, '\\|') + 1  # Count the number of "|" and add 1
#    )
#  )

#Calculate the yearly labour cost
#Data <- Data %>%
#  mutate(cal_labour_cost = f_focus_labour_costs_month*month_count)

# FILL NA VALUES WITH 0
#Data$cal_labour_cost <- Data$cal_labour_cost %>% replace(is.na(.), 0)

```

```{r}
# APPROACH 2: Labour cost per season
Data <- Data %>% 
  mutate(cal_labour_cost = f_focus_labour_costs_season * f_harvest_num)

# FILL NA VALUES WITH 0
Data$cal_labour_cost <- Data$cal_labour_cost %>% replace(is.na(.), 0)

```

```{r}
# APPROACH 3: Labour cost per year 
#Data <- Data %>% 
#  mutate(cal_labour_cost = f_focus_labour_costs_year)

# FILL NA VALUES WITH 0
#Data$cal_labour_cost <- Data$cal_labour_cost %>% replace(is.na(.), 0)

```

### 6) Input costs

There are 2 approaches for assessing input costs: cost per season, and cost per year. Check for this case which approach applies.

```{r, include = FALSE}
# APPRROACH 1: cost per year
# First fill NA values with 0 so that calculation goes well
#Data <- Data %>%
#  mutate(f_inputs_costs_seeds = ifelse(is.na(f_inputs_costs_seeds), 0, f_inputs_costs_seeds)) %>%
#  mutate(f_inputs_costs_fertilizer = ifelse(is.na(f_inputs_costs_fertilizer), 0, f_inputs_costs_fertilizer)) %>%
#  mutate(f_inputs_costs_chemicals = ifelse(is.na(f_inputs_costs_chemicals), 0, f_inputs_costs_chemicals))
#Do the calculation
#Data <- Data %>% 
#  mutate(cal_input_costs = f_inputs_costs_seeds + f_inputs_costs_fertilizer + f_inputs_costs_chemicals)

# APPROACH 2: cost per season
# first fill NA values with 0
Data <- Data %>%
  mutate(f_inputs_costs_seeds_season = ifelse(is.na(f_inputs_costs_seeds_season), 0, f_inputs_costs_seeds_season)) %>%
  mutate(f_inputs_costs_fertilizer_season = ifelse(is.na(f_inputs_costs_fertilizer_season), 0, f_inputs_costs_fertilizer_season)) %>%
  mutate(f_inputs_costs_chemicals = ifelse(is.na(f_inputs_costs_chemicals), 0, f_inputs_costs_chemicals)) %>%
  mutate(f_harvest_num = ifelse(is.na(f_harvest_num), 0, f_harvest_num))
# calculation
Data <- Data %>% 
  mutate(cal_inputs_costs = (f_inputs_costs_seeds_season + f_inputs_costs_fertilizer_season + f_inputs_costs_chemicals) * f_harvest_num)
```

### 7) Other costs

```{r}
# Fill all NA values 
variables_to_replace <- c("f_focus_costs_irrigation", "f_focus_costs_energy",
                          "f_costs_land_focuscrop", "f_focus_costs_storage",
                          "f_focus_costs_marketing")

# Replace NA values with 0 for the specified variables using dplyr
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

# Calculation
Data <- Data %>%
  mutate(cal_focus_other_costs = f_focus_costs_irrigation + f_focus_costs_energy + f_costs_land_focuscrop + f_focus_costs_storage + f_focus_costs_marketing)
```

### 8) Total cost focus crop

```{r}
Data <- Data %>%
    mutate(cal_focus_cost = cal_labour_cost + cal_inputs_costs + cal_focus_other_costs) 
```

### 9) Net-income focus crop

```{r}


#for this case, put cal_focus_income at mssign if f_focus_quant_sold is equal to 0.
Data <- Data %>%
  mutate(cal_focus_income = ifelse(cal_focus_quant_sold_kg == 0, NA, cal_focus_revenue - cal_focus_cost))

```

### 10) Net-income other crops

```{r, include = FALSE}
# Fill NA's 
variables_to_replace <- c("f_othermaincrop_1_costs_labour", "f_othermaincrop_1_costs_inputs",
                          "f_othermaincrop_1_costs_transport", "f_othermaincrop_2_costs_labour"                       ,"f_othermaincrop_2_costs_inputs", "f_othermaincrop_2_costs_transport")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

# calculation
Data <- Data %>%
  mutate(cal_othermaincrop_1_cost = f_othermaincrop_1_costs_labour + f_othermaincrop_1_costs_inputs + f_othermaincrop_1_costs_transport) %>%
  mutate(cal_othermaincrop_2_cost = f_othermaincrop_2_costs_labour + f_othermaincrop_2_costs_inputs + f_othermaincrop_2_costs_transport)

```

Following code calculates production value for other crops. This is only done in Living income studies! Check whether the code applies to this case.

```{r, include = FALSE}
#variables_to_replace <- c("f_othermaincrop_1_quant_prod", "f_othermaincrop_1_price",
#                          "f_othermaincrop_2_quant_prod", "f_othermaincrop_2_price")
#Data <- Data %>%
#  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

#Data <- Data %>%
#  mutate(cal_othermaincrop_production_value = f_othermaincrop_1_quant_prod * f_othermaincrop_1_price + f_othermaincrop_2_quant_prod * f_othermaincrop_2_price)
```

Revenues from other crops:

```{r}
variables_to_replace <- c("f_othermaincrop_1_inc_sold", "f_othermaincrop_2_inc_sold",
                          "f_other_crop_income")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

Data <- Data %>%
  mutate(cal_othermaincrop_1_inc_sold = f_othermaincrop_1_inc_sold) %>%
  mutate(cal_othermaincrop_2_inc_sold = f_othermaincrop_2_inc_sold) %>%
  mutate(cal_other_crop_income = f_other_crop_income)
```

Net income calculation:

```{r}
Data <- Data %>%
  mutate(cal_other_crop_income = cal_othermaincrop_1_inc_sold - cal_othermaincrop_1_cost + cal_othermaincrop_2_inc_sold -  cal_othermaincrop_2_cost + cal_other_crop_income)
```

```{r}
# Subset to check calculations 
subset <- Data %>% select(c(cal_othermaincrop_1_inc_sold, cal_othermaincrop_1_cost, cal_othermaincrop_2_inc_sold, cal_othermaincrop_2_cost, cal_other_crop_income, cal_other_crop_income))
```

### 11) Net-income livestock

```{r}
variables_to_replace <- c("f_livestock_income_total", "f_livestock_nr_hired_labourers",
                          "f_livestock_days_hiredlabour", "f_livestock_wages_hiredlabour",                               "f_livestock_costs_other")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

Data <- Data %>%
  mutate(cal_livestock_revenue = f_livestock_income_total) %>%
  mutate(cal_livestock_labour_cost = (f_livestock_nr_hired_labourers * f_livestock_days_hiredlabour * f_livestock_wages_hiredlabour)) %>%
  mutate(cal_livestock_costs_other = f_livestock_costs_other) %>%
  mutate(cal_livestock_cost = cal_livestock_costs_other + cal_livestock_labour_cost)

# Net income
Data <- Data %>%
  mutate(cal_livestock_income = cal_livestock_revenue - cal_livestock_cost)

```

### 12) General farm costs

```{r}
variables_to_replace <- c("f_equip_costs", "hh_loan_one_size",
                          "hh_loan_one_value_interest", "hh_loan_two_size",                                              "hh_loan_two_value_interest")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

Data <- Data %>% 
  mutate(cal_equipment_costs = f_equip_costs) %>%
  mutate(cal_farm_costs_loans = hh_loan_one_size * (hh_loan_one_value_interest/100) + hh_loan_two_size * (hh_loan_two_value_interest/100)) 

# total costs 
Data <- Data %>% mutate(cal_farm_costs_general = cal_equipment_costs + cal_farm_costs_loans)

```

### 13) Off-farm and other income

```{r}
variables_to_replace <- c("f_nonfarm_enterpr_income_month", "f_income_offfarmlabour_month", "f_income_other_month")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

Data <- Data %>%
  mutate(cal_offfarm_income = 12*(f_nonfarm_enterpr_income_month + f_income_offfarmlabour_month + f_income_other_month))
```

### 14) Farm income calculation

```{r}
Data <- Data %>%
  mutate(cal_farm_income = cal_focus_income + cal_other_crop_income + cal_livestock_income - cal_farm_costs_general,
    cal_farm_income = ifelse(is.na(cal_focus_income), NA, cal_farm_income)) %>%
 mutate(cal_farm_income = ifelse(is.na(farmer_present), NA, cal_farm_income))


```

### 15) Actual income calculation

There are three approaches to calculating the actual income. Please check which approach is suitable for the case!

```{r}
# income share variable needs to be transformed
Data$f_focus_income_share <- as.numeric(Data$f_focus_income_share)
Data$f_othermaincrop_1_income_share <- as.numeric(Data$f_othermaincrop_1_income_share)
Data$f_othermaincrop_2_income_share <- as.numeric(Data$f_othermaincrop_2_income_share)

Data <- Data %>%
  mutate(
    f_other_crop_income_share = ifelse(f_other_crop_income_share == "less than 10", "0", f_other_crop_income_share),
    f_other_crop_income_share_v2 = as.numeric(gsub("[^0-9.]", "", f_other_crop_income_share)) / 10,
    f_other_crop_income_share = f_other_crop_income_share_v2
  ) %>%
  select(-f_other_crop_income_share_v2)

Data$f_other_crop_income_share <- as.numeric(Data$f_other_crop_income_share)
Data$f_livestock_income_share <- as.numeric(Data$f_livestock_income_share)

```

```{r}
Data <- Data %>% 
  # First approach for actual income, based on as many direct inputed values of the farmer.
  mutate(
    cal_actual_income = cal_focus_income + cal_other_crop_income + cal_livestock_income - cal_farm_costs_general + cal_offfarm_income,
    # Set cal_actual_income to missing (NA) if cal_focus_income is 0 or missing
    cal_actual_income = ifelse(is.na(cal_focus_income) | cal_focus_income == 0, NA, cal_actual_income)
  )
```

```{r}
Data <- Data %>% 
  # Second approach for actual income, based on the income share of farmers
  mutate(
    cal_actual_income_v2 = cal_focus_income * (10/f_focus_income_share),
    # Set cal_actual_income_v2 to missing (NA) if cal_focus_income is 0 or missing
    cal_actual_income_v2 = ifelse(is.na(cal_focus_income) | cal_focus_income == 0, NA, cal_actual_income_v2))

```

```{r}
# Third approach for actual. income , based on income share focus crop and income share other crops
# NOT USED BECAUSE THE SUM DID NOT SEEM RELIABLE
#Data <- Data %>%
#  mutate(cal_actual_income_v3 = (cal_focus_income + cal_other_crop_income +cal_livestock_income) * (1/(f_focus_income_share+f_othermaincrop_1_income_share+f_othermaincrop_2_income_share+ f_other_crop_income_share+ f_livestock_income_share)))
         
```

```{r}
# make a check file to see whether everything went correctly 

subset <- Data %>% 
  select(cal_focus_quant_prod_kg, cal_focus_quant_sold_kg, cal_focus_price, 
#         cal_focus_quant_own_consumption_kg, cal_focus_quant_lost_kg, 
         cal_focus_revenue, cal_focus_productivity_acre,cal_labour_cost, cal_inputs_costs, cal_focus_other_costs, cal_focus_cost, cal_focus_income, cal_othermaincrop_1_cost, cal_othermaincrop_2_cost, #cal_othermaincrop_production_value,
cal_othermaincrop_1_inc_sold, cal_othermaincrop_2_inc_sold,cal_other_crop_income, cal_livestock_income, cal_equipment_costs,cal_farm_costs_loans, cal_farm_costs_general, cal_offfarm_income , cal_actual_income) #cal_actual_income_v2, cal_actual_income_v3)

write.xlsx(subset, "calculated_variables_script2.0.xlsx")
```

## Net Promotor score

```{r, include = FALSE}
hh_farmer_gender <- c("all farmers")
    
    NSP_total <- Data %>%
      select(`repeat no`, cs_recommendation) %>% 
      filter(`repeat no` == 1)%>%
      mutate(cs_recommendation = ifelse(is.na(cs_recommendation), "no input",cs_recommendation)) %>%
      mutate(detractor = ifelse(cs_recommendation == "not likely"| cs_recommendation == "somewhat likely"| cs_recommendation == "likely",1,0),
             promoter = ifelse(cs_recommendation == "very likely", 1,0),
             passive = ifelse(cs_recommendation == "most likely"| cs_recommendation == "i dont know",1,0),
             nr_recommenders = ifelse(cs_recommendation == "no input", 0,1)) %>%
      summarize(nr_promoters = sum(promoter),
                nr_detractors = sum(detractor),
                nr_passive = sum(passive),
                nr_farmers = sum(nr_recommenders)) %>%
      mutate(nsp_total = round((nr_promoters/nr_farmers) - (nr_detractors/nr_farmers),2),
             hh_farmer_gender = hh_farmer_gender)
    
    NSP_by_gender <- Data %>%
      select(`repeat no`, cs_recommendation, hh_farmer_gender) %>% 
      filter(`repeat no` == 1)%>%
      mutate(cs_recommendation = ifelse(is.na(cs_recommendation), "no input",cs_recommendation)) %>%
      mutate(detractor = ifelse(cs_recommendation == "not likely"| cs_recommendation == "somewhat likely"| cs_recommendation == "likely",1,0),
             promoter = ifelse(cs_recommendation == "very likely", 1,0),
             passive = ifelse(cs_recommendation == "most likely"| cs_recommendation == "i dont know",1,0),
             nr_recommenders = ifelse(cs_recommendation == "no input", 0,1)) %>%
      group_by(hh_farmer_gender) %>%
      summarize(nr_promoters = sum(promoter),
                nr_detractors = sum(detractor),
                nr_passive = sum(passive),
                nr_farmers = sum(nr_recommenders)) %>%
      mutate(nsp_gender = round((nr_promoters/nr_farmers) - (nr_detractors/nr_farmers),2))
    
    NSP <- full_join(NSP_total, NSP_by_gender)
    NSP <- NSP %>% select(hh_farmer_gender, nr_farmers, nr_promoters, nr_detractors,nr_passive, nsp_total, nsp_gender)
    
```

# Summary statistics

### All Farmers: numerical descriptives

```{r, include = FALSE}
farmer_type <- c("all farmers")
all_var <- Data %>% 
  filter(`repeat no` == 1) %>%
  select(`repeat no`) %>% 
  table() %>% 
  melt(c("farmer type"), value.name="n") %>%
  mutate(`farmer type` = farmer_type)

numerical_descriptives_all_farmers <- Data %>%
  select_if(is.numeric) %>%
  summarise_each(funs(round(mean(., na.rm = TRUE),2))) %>%
  melt(value.name="mean") %>%
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(round(sd(., na.rm = TRUE),2))) %>%
      melt(value.name="sd")) %>%
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(min(., na.rm = TRUE))) %>%
      melt(value.name="min")
  ) %>% 
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(round(max(., na.rm = TRUE),2))) %>%
      melt(value.name="max")
  ) %>% 
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(count_n))  %>%
      melt(value.name="freq")
  ) %>%
  left_join(survey_questions) %>%
  select("variable", "freq", "mean","sd", "min","max") 

```

### All farmers: single categorical descriptives

```{r, include = FALSE}
single_mc <- survey_questions %>% 
  filter(type=="option") %>% 
  filter(is.na(multiple)) %>% 
  pull(variable)

single_categorical_descriptives_all_farmers <- all_var

for(i in single_mc){
  
  if(any(names(Data) %in% i)){
    
    frequencies <- Data %>% 
      filter(`repeat no`==1) %>%
      select(all_of(i)) %>%       
      table() %>% 
      melt(c("category"),value.name = "freq") %>%
      mutate(n = nr_participants_raw) %>%
      mutate(`farmer type` = farmer_type) %>%
      mutate("variable" = i) %>%
      left_join(all_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    single_categorical_descriptives_all_farmers <- single_categorical_descriptives_all_farmers %>% 
      full_join(frequencies) %>%
      select(variable, `farmer type`, n, category, freq, "%")
    
  }
}
```

### All farmers: multiple categorical descriptives

```{r, include = FALSE}
multiple_mc <- survey_questions %>%
  filter(type == "option") %>%
  filter(multiple == "yes") %>%
  pull(variable)

multiple_categorical_descriptives_all_farmers <- all_var

for (j in multiple_mc) {
  print(j)
  
  # Skip the variable "f_labour_otheractivity_labour_type"
  if (j == "f_labour_otheractivity_labour_type") {
    next
  }
  if (j == "hh_loan_one_agri_purpose") {
    next
  }
  if (j == "hh_loan_mulitple_largest_purpose") {
    next
  }
  if (any(names(Data) %in% j)) {
    frequencies <- Data %>%
      select(all_of(j)) %>%
      cSplit(j, "|") %>%
      gather(key, value) %>%
      select(-key) %>%
      table() %>%
      melt(c("category"), value.name = "freq") %>%
      mutate(n = nr_participants_raw) %>%
      mutate("variable" = j) %>%
      left_join(all_var) %>%
      mutate("%" = round(freq / n, 2) * 100) %>%
      mutate(category = as.character(category))
    
    multiple_categorical_descriptives_all_farmers <- multiple_categorical_descriptives_all_farmers %>%
      full_join(frequencies) %>%
      select(variable, `farmer type`, n, category, freq, "%")
  }
}

```

### By gender: numerical descriptives

```{r, include=FALSE}
gender_var <- Data %>% 
  select(hh_farmer_gender) %>% 
  table() %>% 
  melt("hh_farmer_gender", value.name="n")

numerical_descriptives_by_gender <- Data %>%
  group_by(hh_farmer_gender) %>%
  select_if(is.numeric) %>%
  summarise_each(funs(round(mean(., na.rm = TRUE),2))) %>%
  melt(value.name="mean") %>%
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(round(sd(., na.rm = TRUE),2))) %>%
      melt(value.name="sd")) %>%
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(min(., na.rm = TRUE))) %>%
      melt(value.name="min")
  ) %>% 
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(round(max(., na.rm = TRUE),2))) %>%
      melt(value.name="max")
  ) %>% 
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(count_n))  %>%
      melt(value.name="freq")
  ) %>%
  left_join(survey_questions) %>%
  left_join(gender_var) %>%
  select("variable","hh_farmer_gender","n", "freq", "mean","sd", "min","max") 

```

### By gender: single categorical descriptives

```{r, include = FALSE}
single_mc <- survey_questions %>% 
  filter(type=="option") %>% 
  filter(is.na(multiple)) %>% 
  filter(variable != "hh_farmer_gender") %>%
  pull(variable)

single_categorical_descriptives_by_gender <- gender_var

for(i in single_mc){
  
  if(any(names(Data) %in% i)){
    
    frequencies <- Data %>% 
      select(hh_farmer_gender, all_of(i)) %>% 
      table() %>% 
      melt(c("hh_farmer_gender","category"), value.name="freq") %>%
      mutate("variable" = i) %>%
      left_join(gender_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    single_categorical_descriptives_by_gender <- single_categorical_descriptives_by_gender %>% 
      full_join(frequencies) %>%
      select(variable, hh_farmer_gender, n, category, freq, "%")
  }
}
```

### By gender: multiple categorical descriptives

```{r, include = FALSE}
multiple_mc <- survey_questions %>% 
  filter(type=="option") %>% 
  filter(multiple == "yes") %>%
  pull(variable)

multiple_categorical_descriptives_by_gender <- gender_var

for(j in multiple_mc){
  print(j)
  if (j == "f_labour_otheractivity_labour_type") {
    next
  }
  if (j == "hh_loan_mulitple_largest_purpose") {
    next
  }
  if(any(names(Data) %in% j)){
    
    frequencies <- Data %>% 
      select(hh_farmer_gender, all_of(j)) %>% 
      cSplit(j, "|") %>%
      gather(key, value, -hh_farmer_gender) %>%
      select(-key) %>% table() %>%
      melt(c("hh_farmer_gender","category"), value.name="freq") %>%
      mutate("variable" = j) %>%
      left_join(gender_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    multiple_categorical_descriptives_by_gender <- multiple_categorical_descriptives_by_gender %>% 
      full_join(frequencies) %>%
      select(variable, hh_farmer_gender, n, category, freq, "%")
  }
}
```

# Remove repeated rows

Remove repeated rows for household members and create columns for each household member (max household number is 15). Delete the unnecessary, only if household roster questions are asked, check also which ones are used.

```{r, include = FALSE}
if("identifier" %in% colnames(Data)){     
    Data <- Data %>%
      group_by(identifier) %>%
      mutate(id = row_number())
    
    for (nr in c(1:15)){
      
      ### ---- delete the unnecessary --- only if household roster questions are asked, check also which ones are used
      hh_focuscrop_nr <- paste0("hh_focuscrop_", nr)
      
      Data <- Data %>%
        #Delete the rows that do not apply
        mutate(
          !!sym(hh_focuscrop_nr) := ifelse(id == nr, hh_focuscrop_hrs, NA)) %>%
        
        #Delete the rows that do not apply
        fill(!!sym(hh_focuscrop_nr)) %>%
        fill(!!sym(hh_focuscrop_nr), .direction = "up") %>%
        
        #Delete the rows that do not apply
        mutate(
          !!sym(hh_focuscrop_nr) := ifelse(id >1 , NA, !!sym(hh_focuscrop_nr) ))
      
    }
  }
  
  Data <- Data %>%
    filter(id == 1) %>%
    select(-id,
           #-hh_member_birthyear,
           #-hh_member_gender,
           #-hh_member_education,
           
           -hh_focuscrop_hrs)
  
```

# Add match - column

When you are dealing with an endline case, we create a "match" column to indicate whether or not the farmer is recurring. 
PLEASE IGNORE THIS SECTION WHEN YOU ARE DEALING WITH A BASELINE CASE! 

In the 5200 Farmfit Github folder, you can find a folder named "PDC 2.0 sampling". Look here for the pi_ file of the baseline and copy to the case folder. We load in this dataset to match with current farmers and see which ones are recurring. 

First load in the baseline farmers and only keep the phone numbers. 

```{r}
baseline_farmers <- read_excel("pi_ussl_tanzania.xlsx")
# Only keep the phone number column for matching
baseline_farmers <- baseline_farmers %>% select(identifier, mobile_number_farmer)

# Retain only unique rows based on mobile_number_farmer
baseline_farmers <- baseline_farmers %>%
  distinct(mobile_number_farmer, .keep_all = TRUE)
```

Now we match the Data with these phone numbers. 

```{r}
Data <- Data %>% left_join(baseline_farmers, by="mobile_number_farmer")
```
And add a column to indicate the match
```{r}
# Rename columns that changed due to merging
colnames(Data)[colnames(Data) == "identifier.y"] <- "cal_match_baseline"
Data$cal_match_baseline <- gsub("-", "", Data$cal_match_baseline)

colnames(Data)[colnames(Data) == "identifier.x"] <- "identifier"
```

# Check completeness of the codebook

All variables that are in the data should get a description in the codebook. That's why we check this and correct/ add where missing. We also check whether we have all the variables that are required for the data delivery.

```{r, include = FALSE}
question_library <- question_library %>%
    dplyr::rename(section = "Title",
                  question = "Text", 
                  variable = "Variable name",
                  type = "Question type",
                  options = "Options",
                  multiple = "Allow multiple") %>%
    mutate_all(tolower) %>%
    select(section,
           question,
           variable,
           type,
           options,
           multiple) %>%
    fill(section)
  
unmatched <- list()
```

Let's check the variable names. First we need a dataframe with all variable names, and

```{r, include = FALSE}
#Make a dataframe listing all variables 
variable <-ls(Data)
data <- data.frame(variable)

## ---- Variables ending with other  ---

#Check if data variable is in codebook        
data$compare <-data$variable  %in%  survey_questions$variable 

#if it's a "other variable", do not remove
data <- data %>%
  mutate(compare = ifelse(grepl('_other$',variable) & compare == FALSE , FALSE, TRUE)) %>%
  filter(compare == FALSE) 

survey_questions <- bind_rows(survey_questions, data)
survey_questions <- survey_questions %>%
  arrange(variable)  %>%
  fill(question,
       section) %>%
  mutate(type = ifelse(!is.na(compare), "free text",type)) %>%
  
  mutate(variable_2 = str_remove(variable,'_other$')) %>%
  mutate(same = ifelse(variable_2== lag(variable_2), 1, 0)) %>%
  
  mutate(section = ifelse(!is.na(compare) & same == 0 , "unknown", section),
         question = ifelse(!is.na(compare) & same == 0 , "no question available", question)) %>%
  select(-compare, -variable_2, -same)

```

Check if it is a calculated variable or a variable from the question library.

```{r}
variable <-ls(Data)
data <- data.frame(variable)
data$compare <-data$variable  %in%  survey_questions$variable 
data <- data %>%
  filter(compare == FALSE) %>%
  select(-compare)

data$compare <-data$variable  %in%  vars_transformed$variable 

#check whether there is a match by checking out the "data" dataset MANUALLY
#if a variable is not in the vars_transformed, update the excel with variable transformations
```

Append the survey questions with the calculated variables

```{r}
#One changes are applied, append the codebook with information from the calculated variables (vars_transformed)
library(dplyr)

# Assuming "data" and "vars_transformed" are your dataframes

vars_calculated <- data$variable

calculated_variables <- vars_transformed %>%
  filter(variable %in% vars_calculated)

survey_questions <- survey_questions %>%
  bind_rows(calculated_variables)

```

Check whether its a household demographics variable

```{r}
#Check whether already in the codebook
variable <-ls(Data)
data <- data.frame(variable)
data$compare <-data$variable  %in%  survey_questions$variable 
data <- data %>%
  filter(compare == FALSE) 

```

If indeed household demographics variable, append the codebook with that information

```{r}
data <- data %>%
  filter(variable %in% household_demographics$variable) %>%
  select(-compare)

data <- merge(data, household_demographics, by.x = "variable")
survey_questions <- bind_rows(survey_questions, data)
```

```{r, include = FALSE}
# --- Check household demographic questions ---
  hh_demo <- survey_questions %>%
    filter(str_detect(variable, 'hh_member_')) %>%
    arrange(variable) %>%
    fill(section, options)
  
  survey_questions <- survey_questions %>%
    filter(!str_detect(variable, 'hh_member_')) 
  survey_questions <- bind_rows(survey_questions, hh_demo)
  
```

Check whether we have all the required variables for the portal

```{r}
# ---- dashboard variables ----
  #All variables in delivered data into dataframe
  
  variable <-ls(Data)
  data <- data.frame(variable)
  
  #Check if data variable is dashboard list     
  vars_dashboard$compare <- vars_dashboard$latest_var %in% data$variable 
  
  data <- vars_dashboard %>%
    filter(!is.na(latest_var)) %>%
    filter(compare == FALSE) %>%
    filter(is.na(note))
  

```

Prepare final list of unmatched vars

-   Check these by hand and ensure the variable names are aligned with

1)  the question library

2)  list of calculated variables

3)  Household demographic variables.

Ensure that all variables part of the cleaned data or dashboard variable names are described in the codebook.

```{r, include = FALSE}
variable <- ls(Data)
data <- data.frame(variable) %>%
  filter(
    !variable %in% c(survey_questions$variable, vars_transformed$variable, question_library$variable)
  )

```

## Anonymise data for further use

```{r, include = FALSE}
Data <- Data %>%
          rename(location_cascade_region = pi_location_cascade) %>%  ### the variable on the right differs from case to case!!!
          select(
      -c(starts_with("pi_"), -contains("county")),
      -c(name_of_farmer, mobile_number_farmer)) %>%
          rename(pi_location_cascade_first_level = location_cascade_region)
```

# Combine the inputs for data delivery

```{r}

sets <- list(
    "Codebook" = survey_questions,
    "Cleaned Data" = Data,
    "Raw Data (anonymised)" = Data_raw, 
    "Num. desc. all farmers" = numerical_descriptives_all_farmers,
    "Cat. desc. single all farmers" = single_categorical_descriptives_all_farmers,
    "Cat. desc. multi all farmers" = multiple_categorical_descriptives_all_farmers,
    "Num. desc. by gender" = numerical_descriptives_by_gender,
    "Cat. desc. single by gender" = single_categorical_descriptives_by_gender,
    "Cat. desc. multi by gender" = multiple_categorical_descriptives_by_gender,
    "Net promoter score" = NSP)
  
  write.xlsx (sets, file = (paste0(data_delivery)))
```

# Get personal information sheet

```{r, include = FALSE}
## ---- Extract personal information ----
Data_raw <- read_excel(data_filename)

Data_raw <- Data_raw %>%
  select(-contains("--option--")) %>%
  rename_all(funs(tolower)) %>%
  filter(`repeat no`==1) %>%
  select(identifier,
         name_of_farmer,
         mobile_number_farmer,
         contains("pi_")) #-contains("ppi")) 


pi_info<- list("Personal information" = Data_raw)
write.xlsx(pi_info, file="pi_2023_Tanzania_maize.xlsx")

```

```{r}
#Productivity check

productivity <- Data %>%
    mutate(focus_product_min_sold = cal_focus_quant_prod_kg - cal_focus_quant_sold_kg) %>%

    select(identifier,
           #submitter,
           #`submission date`, ##MAKE SURE THIS VARIABLE IS NOT DELETED IN THE FIRST PIECE OF THE CODE.It should be deleted before sharing with IDH, but it should be kept when checking the productivity numbers.
           starts_with("pi_location_cascade_level"), #include the applicable variable, this differs per case
           f_size,
           f_size_acre,
           f_focus_crop_size,
           f_focus_crop_size_acre,
           cal_focus_productivity_acre,
           f_focus_quant_prod,
           f_focus_measurement_prod,
           f_focus_quant_sold,
           f_focus_measurement_sold,
           focus_product_min_sold,
           f_focus_price,
           f_focus_income_share,
           f_focus_own_consumption_share,
           f_focus_lost_share,
           f_livestock_income_total,
           f_livestock_nr_labourers,
           f_livestock_nr_hired_labourers,
           f_livestock_days_hiredlabour,
           f_livestock_wages_hiredlabour,
           f_livestock_costs_other,
           f_income_offfarmlabour_month,
           f_focus_production_costs_season,
           f_labour_hired_share,
           f_focus_labour_wage_per_day,
           f_focus_labour_costs_month,
           f_focus_labour_costs_season,
           f_focus_labour_costs_year,
           f_inputs_costs_seeds_season,
           f_inputs_costs_fertilizer_season,
           f_inputs_costs_chemicals,
           f_coop_fee
           )
  
write.xlsx (productivity, file = "productivity_check.xlsx", overwrite = TRUE)
```
