---
title: "Data_cleaning_nigeria_palm_oil"
output: html_document
date: "2025-04-07"
---

```{r setup, include=FALSE}
rm(list = ls(), envir = .GlobalEnv)
knitr::opts_chunk$set(echo = TRUE)
```

This script is developed to clean data from IDH cases in a standardized way.

### Preparation

Load the necessary packages.

```{r, include = FALSE}
library(here); library(readxl); library(openxlsx);
library(plyr); library(dplyr); library(tidyr); library(tidylog); library(tidyselect)
library(stringr); library(data.table); library(reshape2)
library(zoo);library(splitstackshape) ;library(tidyverse); library(magrittr)
```

Fill in the basic case information.

```{r}
case = "Okomu"
year = "2025"
focus_crop = "palm_oil"
country = "Nigeria"
```

Make a new Github map with the project name where this R Markdown file is stored, together with the raw data and the survey form downloaded from Flow. Fill in the right file names in the code below so the variables data_filename and survey_filename refer to the correct file.

```{r}
data_filename <- "Farmer_survey_Okomu_Nigeria_-_Palm_oil_-_all_versions_-_English_-_2025-04-07-12-26-10 (1).xlsx"
survey_filename <- "azgx92zfMV8LeoNoVnsQvN.xlsx"
```

Describe the files that will be created in the script here.

```{r}
pi_filename <- paste("pi_",year,"_",case,"_",focus_crop,".xlsx", sep="")
data_delivery <- paste(year,"_",case,"_anom_",focus_crop,"_baseline",".xlsx", sep = "")
```

Import the data files that can be found in the map Analytics-Guild -\> IDH -\> Templates and input files for data delivery. Copy them in the project map.

```{r}
question_library <- readxl::read_excel("IDH Farmer Survey Builder.xlsx", sheet = "survey")

vars_transformed <- read_excel("variables with transformation_2.xlsx") 
household_demographics  <- read_excel("household_demographics_var_2.xlsx") 
vars_dashboard <- read_excel("variable names input dashboard_2.xlsx") 

```


Define functions that will be needed later in the script.

```{r}
# DATA TYPE Set factors to integers
factor_to_int <- function(x){
  if (is.factor(x) || is.character(x)) {
    return(as.numeric(as.character(x)))
  } else {
    return(as.numeric(x))
  }
}

# Find and replace outliers
outlier_detection <- function(x){
  ifelse(
    x > (mean(x, na.rm=TRUE) + sd(x, na.rm=TRUE)*3) |
      x < (mean(x, na.rm=TRUE) - sd(x, na.rm=TRUE)*3), 
    9997,
    x
  )
}

# Avoid scientific notation in printed numbers
options(scipen = 999)  

# count values without NA
count_n <- function(x){sum(!is.na(x))}
```

### Import data

We read the data downloaded. The below is for a KOBO format. On Data we will perform cleaning. Data_raw will be added to the data delivery as the full raw data file without transformations. Only the personal information variables are removed.

**Note**: specify the correct sheet name

```{r, include = FALSE}
# Data for transformation
Data <- read_excel(data_filename, sheet = "Farmer survey Okomu Nigeria")
focus_repeat_revenue <- read_excel(data_filename, sheet = "focus_repeat_revenue")

# Read file with survey structure
survey_questions <- read_excel((survey_filename),sheet ="survey")
questions_choices <- read_excel ((survey_filename), sheet ="choices")
```

Execute mapping so that columns are not questions but variable names. 

```{r}
Data <- Data %>%
  rename(`Is this farmer part of the Okomu database ?` = `Is this farmer part of the Okomu database ?...14`)
```

```{r}
library(stringdist)

# Prepare clean mapping
question_to_variable <- survey_questions %>%
  filter(!is.na(`label::English`)) %>%
  mutate(
    label_clean = tolower(trimws(`label::English`)),
    name_clean = trimws(name)
  ) %>%
  select(label_clean, name_clean)

# Create a function to rename a column if it starts with a label
rename_column <- function(col_name, mapping) {
  col_name_clean <- tolower(trimws(col_name))
  
  # Try to find a match where the label is a prefix of the column name
  matches <- mapping$label_clean[sapply(mapping$label_clean, function(lbl) startsWith(col_name_clean, lbl))]
  
  if (length(matches) > 0) {
    best_label <- matches[which.max(nchar(matches))]  # choose longest matching label
    var_name <- mapping %>% filter(label_clean == best_label) %>% pull(name_clean)
    
    suffix <- substr(col_name, nchar(best_label) + 2, nchar(col_name))  # +2 accounts for the slash
    return(paste0(var_name, "/", suffix))
  }
  
  return(col_name)  # fallback
}

# Apply the renaming logic
new_names <- sapply(names(Data), rename_column, mapping = question_to_variable)
names(Data) <- new_names

# Step 4: Remove columns that are entirely NA
Data <- Data %>%
  select(where(~ !all(is.na(.))))

#Remove "/" at the end of the variable name
names(Data) <- gsub("/$", "", names(Data))

#Delete columns from previous versions of the survey
Data <- Data[, -c(535:542)]

```

```{r}
# Prepare the mapping
question_to_variable <- survey_questions %>%
  filter(!is.na(`label::English`)) %>%
  select(`label::English`, name)

mapping_dict <- setNames(question_to_variable$name, question_to_variable$`label::English`)

# Rename columns in Data, including support for question_text/answer_option format
new_names <- names(Data)

new_names <- sapply(new_names, function(col_name) {
  if (col_name %in% names(mapping_dict)) {
    # Exact match
    return(mapping_dict[[col_name]])
  } else if (grepl("/", col_name, fixed = TRUE)) {
    # Handle multiple choice columns
    parts <- strsplit(col_name, "/", fixed = TRUE)[[1]]
    question_part <- parts[1]
    option_part <- parts[2]
    
    if (question_part %in% names(mapping_dict)) {
      return(paste0(mapping_dict[[question_part]], "/", option_part))
    }
  }
  return(col_name)  # Return as-is if no match
})

# Apply new column names
names(Data) <- new_names

# Verify the result
head(Data)
```
```{r}
# Rename columns at position 105 and 106
names(Data)[113:114] <- c(
  "f_focus_production_costs_types/Nonmech_Buying_renting",
  "f_focus_production_costs_types/Mech_Buying_renting"
)
names(Data)[113:114]

```

```{r}
# Check for any duplicate names
duplicated_names <- names(Data)[duplicated(names(Data))]
duplicated_names
```


```{r}
# Raw data 
Data_raw <- Data

#Anonymise
Data_raw <- Data_raw %>%
  select(-starts_with("pi_"), -name_of_farmer, -mobile_number_farmer)

Data_focus_repeat_raw <- read_excel(data_filename, sheet = "focus_repeat_revenue")

## ---- Extract personal information ----
pi_information <- Data %>%
 # select(-contains("--option--")) %>% #how to deal with that in kobo format?
  rename_all(funs(tolower)) %>%
  select(uuid,
         name_of_farmer,
         mobile_number_farmer,
         contains("pi_")) #-contains("ppi")) 

pi_info<- list("Personal information" = pi_information)

write.xlsx(pi_info, file=pi_filename) #pi_filename was created in the beginning of the script

```

Adjust the survey file to be ready for analysis. We now have a clear overview of the survey questions asked in this case. The below code is for Kobo format

```{r, include = FALSE}
# SURVEY QUESTION SHEET

survey_questions %<>%
  # Select relevant columns: type, name, label in English, and relevant
  select(type, name, `label::English`, relevant) %>%
  # Trim any trailing spaces from character columns
  mutate(across(where(is.character), str_trim)) %>%
  # Rename "label: English" to a simpler "label"
  rename(label = `label::English`) %>%
  # Convert all data to lowercase for consistency
  mutate(across(everything(), tolower)) %>%
  # Filter out rows where the 'type' column is missing (empty rows)
  filter(!is.na(type)) %>%
  # Filter out rows where the 'label' column is missing (empty rows)
  filter(!is.na(label)) %>%
  # Filter out rows where the type is 'note', 'begin_group', or 'begin_repeat'
  filter(!type %in% c("note", "begin_group", "begin group", "begin_repeat"))

# CHOICES QUESTION SHEET
questions_choices %<>%
  # Select relevant columns: list_name, name, label in English, and label in Portuguese
  select(list_name, name, `label::English`) %>%
  # Trim any trailing spaces from character columns
  mutate(across(where(is.character), str_trim)) %>%
  # Rename "label: English" to "label" and "list_name" to "choices_name"
  rename(label = `label::English`, choices_name = list_name) %>%
  # Convert all data to lowercase for consistency
  mutate(across(everything(), tolower)) %>%
  # Filter out rows where the 'choices_name' column is missing (empty rows)
  filter(!is.na(choices_name))
```

What are the numerical columns? How many participants?

```{r}
# Collect the different variables, numerical questions:
numerical_columns <- survey_questions %>%
  filter(type %in% c("integer", "decimal")) %>%
  select("name") %>% pull()
numerical_columns
#Ensure those columns in Data are numeric
Data <- Data %>%
  mutate(across(all_of(numerical_columns), ~as.numeric(.), .names = "{.col}"))

#check which columns are now numeric
numeric_check <- sapply(Data[numerical_columns], is.numeric)
print(numeric_check)
## ---- Count initial number of participants
nr_participants_raw <- length(unique(Data$`_uuid`))
nr_participants_raw
```

# Section 1. Cleaning the data

## 1. Rename variables where necessary

```{r}
# Rename all columns with "_season" to "_year" because in this dataset we assessed all revenues costs per year 2024 instead of season
names(Data) <- gsub("_season", "_year", names(Data))

survey_questions <- survey_questions %>%
  mutate(name = gsub("_season", "_year", name))

#We need to replace hh_loan_agri_source_other with hh_loan_agri_size as this was a mistake in the template
Data <- Data %>%
  rename(hh_loan_agri_size = hh_loan_agri_source_other)

survey_questions <- survey_questions %>%
  mutate(name = ifelse(name == "hh_loan_agri_source_other", "hh_loan_agri_size", name))

Data <- Data %>%
  mutate(
    cs_sdm_company_services_years = if_else(uuid == "20250403124310", 4, as.numeric(cs_sdm_company_services_years)),
    cs_sdm_company_services_years = if_else(uuid == "20250402142509", 5, as.numeric(cs_sdm_company_services_years)),
    cs_sdm_company_yrs_selling = if_else(uuid == "20250403124310", 4, as.numeric(cs_sdm_company_yrs_selling))
  )

```

## 2. Standard data cleaning (dates, columns, lowercase, participation, duplicates)

```{r, include = FALSE}
Data <- Data %>%
 
  # Remove irrelevant columns
  select(-c(devicephonenum, deviceid, subscriberid, simid, 
            `__version__`, `_status`, `_submitted_by`, 
            )) %>%
  select(-contains("--option--")) %>%
  
  # All column names to lowercase
  rename_all(tolower) %>%
  
  # Remove farmers who didn’t participate
  filter(surv_consent != "Refused to participate") %>%
  
  # Remove completely empty columns
  select(where(~ !all(is.na(.))))
```

## 3. Other variables

Identify the "other" variables, and combine the values with their parent. Then the "other" variables are removed.

```{r, include = FALSE}
variables_other <- Data %>% 
  select(ends_with("_other"), ends_with("_other2")) %>% 
  names() 

for(other in variables_other){
  
  if(any(names(Data) %in% gsub("_other$|_other2$", "", other))){
    
    # Fuzzy match? - many spelling mistakes in "other" questions
    
    # Combine "other" variables with their parents
    Data <- Data %>%
      unite(!!gsub("_other$|_other2$", "", other), 
            c(gsub("_other$|_other2$", "", other), all_of(other)), 
            sep="|", remove=FALSE, na.rm=TRUE) %>%
      select(-other)
  }
}
```

## 4. Correct prices

```{r}
Data <- Data %>%
  mutate(f_focus_price = case_when(
    uuid == "20250402103135" ~ 1800,
    uuid == "20250403232810" ~ 160000,
    uuid == "20250402120910" ~ 170000,
    uuid == "20250403175416" ~ 180000,
    uuid == "20250403182040" ~ 160000,
    uuid == "20250403111719" ~ 180000,
    uuid == "20250402202201" ~ 13000,
    uuid == "20250405100755" ~ 160000,
    uuid == "20250404174723" ~ 170000,
    uuid == "20250405100736" ~ 170000,
    uuid == "20250404104739" ~ 170000,
    uuid == "20250406132354" ~ 170000,
    uuid == "20250406132324" ~ 170000,
    uuid == "20250406171539" ~ 170000,
    uuid == "20250406183414" ~ 170000,
    uuid == "20250403135132" ~ 170000,
    uuid == "20250406162538" ~ 180000,
    uuid == "20250405134554" ~ 180000,
    uuid == "20250406092649" ~ 180000,
    uuid == "20250406101522" ~ 180000,
    uuid == "20250406213045" ~ 180000,
    uuid == "20250404151319" ~ 180000,
    uuid == "20250403155837" ~ 180000,
    uuid == "20250407094317" ~ 180000,
    uuid == "20250406173204" ~ 200000,
    uuid == "20250406195031" ~ 200000,
    uuid == "20250406191720" ~ 200000,
    uuid == "20250403150424" ~ 220000,
    uuid == "20250404103754" ~ 220000,
    uuid == "20250403135154" ~ 220000,
    uuid == "20250406162605" ~ 1400,
    uuid == "20250404102248" ~ 240000,
    uuid == "20250404212240" ~ 1300,
    uuid == "20250402185816" ~ 35000,
    uuid == "20250404163027" ~ 50000,
    uuid == "20250402155505" ~ 210000,
    uuid == "20250403162721" ~ 275000, 
    uuid == "20250403092515" ~ 180459.7701,# total revenue instead of unit price
    uuid == "20250402152157" ~ 254.491018,# total revenue instead of unit price
    uuid == "20250402114502" ~ 1166.666667,# total revenue instead of unit price
    uuid == "20250402213407" ~ 147,
    uuid == "20250402163841" ~ 157,
    uuid == "20250402183452" ~ 157,
    uuid == "20250402195231" ~ 157,    
    uuid == "20250402180535" ~ 220,     
    uuid == "20250402165052" ~ 700,     
    uuid == "20250404194002" ~ 900, 
    uuid == "20250403165347" ~ 1300,
    uuid == "20250402181509" ~ 1300,   
    uuid == "20250403080923" ~ 1300,       
    uuid == "20250405201822" ~ 1300, 
    uuid == "20250406101541" ~ 1300,    
    uuid == "20250403143947" ~ 1300,    
    uuid == "20250405112335" ~ 1300, 
    uuid == "20250402202201" ~ 1300,     
    uuid == "20250404203517" ~ 1300,  
    uuid == "20250402145632" ~ 1300,  
    uuid == "20250403183341" ~ 1300, 
    uuid == "20250406220011" ~ 1600, 
    uuid == "20250405204231" ~ 1600,  
    uuid == "20250404212654" ~ 2200,  
    uuid == "20250402202519" ~ 3500,  
    uuid == "20250404163027" ~ 150000, 
    uuid == "20250405111424" ~ 150000, 
    uuid == "20250405153409" ~ 150000, 
    uuid == "20250405172217" ~ 150000, 
    uuid == "20250406121517" ~ 180000,
    uuid == "20250406125819" ~ 180000,
    uuid == "20250405210650" ~ 180000,
    uuid == "20250403141913" ~ 180000,
    uuid == "20250406063922" ~ 180000,
    uuid == "20250406081001" ~ 180000,
    uuid == "20250404085210" ~ 210000,
    uuid == "20250402185816" ~ 210000,
    uuid == "20250404212548" ~ 240000,
    uuid == "20250403124311" ~ 240000,
    
    uuid == "20250405153259" ~ 595.4734961,
    uuid == "20250406163024" ~ 595.4734961,
    uuid == "20250406164824" ~ 595.4734961,
    uuid == "20250406193038" ~ 595.4734961,
    uuid == "20250405132506" ~ 595.5330554,
    uuid == "20250402183949" ~ 650.487963565387,
    uuid == "20250404212917" ~ 650.487963565387,
    uuid == "20250404081708" ~ 650.487963565387,
    uuid == "20250405110540" ~ 650.487963565387,
    uuid == "20250406112656" ~ 650.487963565387,
    uuid == "20250402150226" ~ 650.553025374105,
    uuid == "20250405165302" ~ 650.553025374105,
    uuid == "20250405202116" ~ 650.553025374105,
    uuid == "20250403183401" ~ 654.319371727749,
    uuid == "20250405132310" ~ 654.319371727749, 
    uuid == "20250403123410" ~ 778.660436137072, 
    uuid == "20250403174509" ~ 778.660436137072, 
    uuid == "20250402211150" ~ 778.660436137072, 
    uuid == "20250403203121" ~ 778.660436137072,  
    
    uuid == "20250402121653" ~ 778.660436137072,  
    uuid == "20250406150131" ~ 778.660436137072,  
    uuid == "20250403194837" ~ 778.660436137072,
    uuid == "20250405000449" ~ 778.660436137072,
    uuid == "20250404073806" ~ 778.660436137072,
    uuid == "20250404163411" ~ 778.660436137072,
    uuid == "20250404204542" ~ 778.660436137072,
    uuid == "20250406185145" ~ 778.660436137072,
    uuid == "20250405164924" ~ 778.738317757009,
    uuid == "20250406110123" ~ 778.738317757009,
    uuid == "20250405161238" ~ 180000,
    
    TRUE ~ as.numeric(f_focus_price)  # retain existing value for other rows
  ))
```


## 5. Specific adjustments for quantities and measurement units

This is case specific.

```{r}
Data <- Data %>%
  mutate(f_focus_quant_prod = as.numeric(f_focus_quant_prod)) %>%
  mutate(f_focus_quant_prod = if_else(
    uuid == "20250407102812",
    35000,
    f_focus_quant_prod
  )) %>%
  mutate(f_focus_quant_prod = if_else(
    uuid == "20250403232737",
    16,
    f_focus_quant_prod
  )) %>%
  mutate(f_focus_quant_prod = if_else(
    uuid == "20250405161238",
    68,
    f_focus_quant_prod
  )) %>%
  mutate(f_focus_quant_prod = if_else(
    uuid == "20250402155505",
    80,
    f_focus_quant_prod
  )) 
  
  
  
```

```{r}
Data <- Data %>%
  mutate(f_focus_quant_sold = as.numeric(f_focus_quant_sold)) %>%
  mutate(f_focus_quant_sold = if_else(
    uuid == "20250406191720",
    40,
    f_focus_quant_sold
  )) %>%
  mutate(f_focus_quant_sold = if_else(
    uuid == "20250407035159",
    150,
    f_focus_quant_sold
  )) %>%
  mutate(f_focus_quant_sold = if_else(
    uuid == "20250405161238",
    68,
    f_focus_quant_sold
  ))
```

Measurement unit

```{r}
Data <- Data %>%
  mutate(f_focus_measurement_sold = if_else(uuid %in% c("20250402103135", "20250406162605","20250403165347","20250402114502"),
                                            "Number of bunches",f_focus_measurement_sold)) %>%
  mutate(f_focus_measurement_sold = if_else(uuid=="20250405161238", "Tonnes", f_focus_measurement_sold))

Data <- Data %>%
  mutate(f_focus_measurement_prod = if_else(uuid=="20250405161238", "Tonnes", f_focus_measurement_prod))
  
```

## 6. Handle outliers and missing values

Exclude variables such as quantities, price and farm size to be checked manually.

```{r, include = FALSE}
excluded_columns <- c("f_size", "f_focus_crop_size", "f_focus_quant_prod", "f_focus_quant_sold", "f_focus_quant_lost", "f_focus_own_consumption") # Replace with actual variable names

# Identify numerical columns for outlier handling, excluding those ending with "_#"
numerical_columns <- names(Data %>% select_if(is.numeric)) %>%
  .[!grepl("_\\d+$", .)]  %>%  # Exclude columns ending with "_#"
  .[!(. %in% excluded_columns)] %>%  # Exclude specific columns
  .[sapply(Data[.], function(x) length(unique(x)) > 2)]  # Exclude binary columns
```

The following piece of code will delete outliers. 

!! Make sure to only run this code AFTER the productivity check !! 

```{r, include = FALSE}
# Apply the outlier detection function to numerical columns
Data <- Data %>%
    mutate_at(vars(numerical_columns), funs(outlier_detection)) %>%
  
# Replace specific values (9999, 9998, 9997) with NA for numeric columns
  mutate_if(is.numeric, list(~na_if(., 9999))) %>%
  mutate_if(is.numeric, list(~na_if(., 9998))) %>%
  mutate_if(is.numeric, list(~na_if(., 9997)))

# Change "I don't know" and "I prefer not to say" into NAs
Data <- Data %>%
  mutate_if(is.character, list(~na_if(., "i don't know"))) %>%
  mutate_if(is.character, list(~na_if(., "i prefer not to say")))

```

**Data focus revenue repeat**

```{r}
# Identify numerical columns for outlier handling, excluding those ending with "_#"
numerical_columns <- names(focus_repeat_revenue %>% select_if(is.numeric)) %>%
  .[!grepl("_\\d+$", .)]  %>%  # Exclude columns ending with "_#"
  .[sapply(focus_repeat_revenue[.], function(x) length(unique(x)) > 2)]  # Exclude binary columns

# Apply the outlier detection function to numerical columns
focus_repeat_revenue <- focus_repeat_revenue %>%
    mutate_at(vars(numerical_columns), funs(outlier_detection)) %>%

  # Replace specific values (9999, 9998, 9997) with NA for numeric columns
  mutate_if(is.numeric, list(~na_if(., 9999))) %>%
  mutate_if(is.numeric, list(~na_if(., 9998))) %>%
  mutate_if(is.numeric, list(~na_if(., 9997))) %>%

  # Replace "I don't know" or "I prefer not to say" with NA for character columns
  mutate_if(is.character, list(~na_if(., "i don't know"))) %>%
  mutate_if(is.character, list(~na_if(., "i prefer not to say")))
```

## 7. Adjust Farm Size

Adjust the **farm sizes** to acres.

```{r}
# First check what are the unit measurements reported in this case.
table(Data$f_unit_land)

#table(Data$f_unit_land_other2)
```

```{r}
                             
```

The code below creates a variable for farm sizes in both acres and hectares. When another option is given as unit of measurement (other than acres or hectares), the code needs to be manually adjusted!


```{r, include = FALSE}

#Mutate the different units to acres. We need a farm size in acres for the total farm size, and the size dedicated to focus crop.
Data <- Data %>%
  mutate(
  # First convert to numeric
    f_size = as.numeric(f_size),
    f_focus_crop_size = as.numeric(f_focus_crop_size),
    
#Convert land size to acres if in hectares
    f_size_acre = ifelse(f_unit_land == "Hectares", f_size * 2.471, f_size),
    f_focus_crop_size_acre = ifelse(f_unit_land == "Hectares", f_focus_crop_size * 2.471, f_focus_crop_size),
#Mutate the different units to hectares. We need a farm size in hectares for the total farm size, and the size dedicated to focus crop.
    f_size_hectare = ifelse(f_unit_land == "Hectares", f_size, f_size_acre / 2.471),
    f_focus_crop_size_hectare = ifelse(f_unit_land == "Hectares", f_focus_crop_size, f_focus_crop_size_acre / 2.471))
```

## 8. Some statistics

```{r}
# Number of participants who gave informed consent 
nr_participants_ic <- length(unique(Data$uuid))

# NUMBER OF SEASONS
number_of_seasons <- max(Data$f_harvest_num, na.rm=TRUE)
```

# Section 2. Actual income calculation

## 1) Transform unit measurements

Each unit measurement variables needs to be aligned. let's handle the transformation of unit measurements for production, sales, consumption and losses. We transform every variable to kg.

First check what are the unit measurement reported in this case. Use the code below, or check the google looker dashboard.

```{r}
table(Data$f_focus_measurement_prod)
#table(Data$f_focus_measurement_sold)
#table(Data$f_focus_measurement_lost)
table(Data$f_focus_crop_age)

```

To do so, we extract the numerical values from the measurement units for quantity produced, sold, consumed and lost. Always check by hand whether this is going well. We extract the numbers from the measurement units to calculate total quantities. If there are other measurement units, not related to kg, put a value for those in the variable: cal_focus_measurement_prod

For palm oil, the quantity depends on the age of the tree. That is why we asked this question as well, to adapt the cal_focus_measurement_ variable to this. The weight values are the following: 
- More than 25 years old: 15.28 kg
- 20-24 years old: 16.54 kg
- 15-19 years old: 16.79 kg
- 10-14 years old: 15.37 kg
- 5-9 years old: 12.84 kg
- 0-5 years old: 12.84 kg

```{r}
# First clean the data in f_focus_crop_age
Data <- Data %>%
  mutate(f_focus_crop_age = recode(f_focus_crop_age,
    "Other, please specify|2015" = "2011 to 2015",
    "Other, please specify|2019" = "2016 to 2020",
    "Other, please specify|2020" = "2016 to 2020",
    "Other, please specify|2020-2022" = "2021 to 2025"
  ))
```

**Measurement unit production**

```{r, include = FALSE}
Data <- Data %>%
  mutate(
    cal_focus_measurement_prod = as.numeric(case_when(
      grepl("[[:digit:]]", f_focus_measurement_prod) ~ parse_number(f_focus_measurement_prod),
      f_focus_measurement_prod %in% c("kg", "Kgs") ~ 1,
      f_focus_measurement_prod == "Tonnes" ~ 1000,
      f_focus_measurement_prod %in% c("Other, please specify|Litres", 
                                      "Other, please specify|Litres (after processing)") ~ 1,
      # Handle 'Number of bunches' based on crop age
      f_focus_measurement_prod == "Number of bunches" & f_focus_crop_age == "2001 to 2005" ~ 16.54,
      f_focus_measurement_prod == "Number of bunches" & f_focus_crop_age == "2006 to 2010" ~ 16.79,
      f_focus_measurement_prod == "Number of bunches" & f_focus_crop_age == "2011 to 2015" ~ 15.37,
      f_focus_measurement_prod == "Number of bunches" & f_focus_crop_age == "2016 to 2020" ~ 12.84,
      f_focus_measurement_prod == "Number of bunches" & f_focus_crop_age == "2021 to 2025" ~ 12.84,
      f_focus_measurement_prod == "Number of bunches" & f_focus_crop_age == "Before 2000" ~ 15.28,
      #Fallback value
      TRUE ~ 0
    ))
  )
# Create a dataframe with only the two columns to check whether the calculation went allright
check_measurement_conversion_prod <- Data %>%
  select(f_focus_measurement_prod, cal_focus_measurement_prod)

```

**Measurement unit sales**

```{r, include = FALSE}

Data <- Data %>%
  mutate(
    cal_focus_measurement_sold = as.numeric(case_when(
      grepl("[[:digit:]]", f_focus_measurement_sold) ~ parse_number(f_focus_measurement_sold),
      f_focus_measurement_sold %in% c("kg", "Kgs") ~ 1,
      f_focus_measurement_sold == "Tonnes" ~ 1000,
      f_focus_measurement_sold %in% c("Other, please specify|Litres", 
                                      "Other, please specify|Litres (after processing)") ~ 1,
      # Handle 'Number of bunches' based on crop age
      f_focus_measurement_sold == "Number of bunches" & f_focus_crop_age == "2001 to 2005" ~ 16.54,
      f_focus_measurement_sold == "Number of bunches" & f_focus_crop_age == "2006 to 2010" ~ 16.79,
      f_focus_measurement_sold == "Number of bunches" & f_focus_crop_age == "2011 to 2015" ~ 15.37,
      f_focus_measurement_sold == "Number of bunches" & f_focus_crop_age == "2016 to 2020" ~ 12.84,
      f_focus_measurement_sold == "Number of bunches" & f_focus_crop_age == "2021 to 2025" ~ 12.84,
      f_focus_measurement_sold == "Number of bunches" & f_focus_crop_age == "Before 2000" ~ 15.28,
      TRUE ~ 0
    ))
  )

# Create a dataframe with only the two columns to check whether the calculation went allright
check_measurement_conversion_sold <- Data %>%
  select(f_focus_measurement_sold, cal_focus_measurement_sold)


```

**Measurement unit focus crop lost**

```{r, include = FALSE}

Data <- Data %>%
  mutate(
    cal_focus_measurement_lost = as.numeric(case_when(
      grepl("[[:digit:]]", f_focus_measurement_lost) ~ parse_number(f_focus_measurement_lost),
      f_focus_measurement_lost %in% c("kg", "Kgs") ~ 1,
      f_focus_measurement_lost == "Tonnes" ~ 1000,
      f_focus_measurement_lost %in% c("Other, please specify|Litres", 
                                      "Other, please specify|Litres (after processing)") ~ 1,
      # Handle 'Number of bunches' based on crop age
      f_focus_measurement_lost == "Number of bunches" & f_focus_crop_age == "2001 to 2005" ~ 16.54,
      f_focus_measurement_lost == "Number of bunches" & f_focus_crop_age == "2006 to 2010" ~ 16.79,
      f_focus_measurement_lost == "Number of bunches" & f_focus_crop_age == "2011 to 2015" ~ 15.37,
      f_focus_measurement_lost == "Number of bunches" & f_focus_crop_age == "2016 to 2020" ~ 12.84,
      f_focus_measurement_lost == "Number of bunches" & f_focus_crop_age == "2021 to 2025" ~ 12.84,
      f_focus_measurement_lost == "Number of bunches" & f_focus_crop_age == "Before 2000" ~ 15.28,
      
      TRUE ~ 0
    ))
  )

# Create a dataframe with only the two columns to check whether the calculation went allright
check_measurement_conversion_lost <- Data %>%
  select(f_focus_measurement_lost, cal_focus_measurement_lost)
```

## 2) Quantities produced, sold, lost and own consumption

In this section the quantities produced, sold, lost and own consumption are cleaned. Now that the unit measurements are transformed, the quantities produced, sold, lost and own consumption can be calculated in kgs.

```{r, include = FALSE}
Data <- Data %>%
    # Convert relevant columns to numeric
  mutate(across(c(f_focus_quant_prod, cal_focus_measurement_prod, f_focus_quant_sold, cal_focus_measurement_sold, f_focus_quant_lost, cal_focus_measurement_lost, 
                  f_focus_price), as.numeric)) %>%
  ##Calculate the quantities focus crop
  mutate(cal_focus_quant_prod_kg = f_focus_quant_prod * cal_focus_measurement_prod) %>%
  mutate(cal_focus_quant_sold_kg = f_focus_quant_sold * cal_focus_measurement_sold) %>%
  mutate(cal_focus_quant_lost_kg = f_focus_quant_lost * cal_focus_measurement_lost) %>%
 # mutate(cal_focus_quant_own_consumption_kg = f_focus_own_consumption * cal_focus_measurement_own_consumption)%>%
  mutate_at(vars(starts_with("cal_focus_quant")), ~replace_na(., 0)) 
    
  # also calculate the unit price 
Data <- Data %>%
  mutate(cal_focus_price = ifelse(cal_focus_measurement_prod != 0, f_focus_price/cal_focus_measurement_sold, f_focus_price),
    cal_focus_price = ifelse(is.na(cal_focus_quant_sold_kg) | cal_focus_quant_sold_kg == 0, NA, cal_focus_price))%>%
  mutate(cal_focus_price = replace_na(cal_focus_price, 0))

Data <- Data %>%
  mutate(
    cal_focus_quant_prod_kg = as.numeric(cal_focus_quant_prod_kg),
    cal_focus_quant_sold_kg = as.numeric(cal_focus_quant_sold_kg),
    cal_focus_quant_lost_kg = as.numeric(cal_focus_quant_lost_kg),
    cal_focus_price = as.numeric(cal_focus_price)
  )

```

```{r, include = FALSE}
#Replace Price NA with median if there is record of sale
Data <- Data %>%
  mutate(cal_focus_price = ifelse(cal_focus_quant_sold_kg > 0 & is.na(cal_focus_price), median(cal_focus_price), cal_focus_price))
```

Check whether the calculations went well. We make a subset of the data with relevant variables. Click on the dataset to sort the columns to detect strange numbers and solve if necessary.

```{r, include = FALSE}

quantities_produced_subset <- Data %>% select(f_focus_measurement_prod, cal_focus_measurement_prod, f_focus_quant_prod, cal_focus_quant_prod_kg)

quantities_sold_subset <- Data %>% select(f_focus_measurement_sold, cal_focus_measurement_sold, f_focus_quant_sold, cal_focus_quant_sold_kg)

quantities_lost_subset <- Data %>% select(f_focus_measurement_lost, cal_focus_measurement_lost, f_focus_quant_lost, cal_focus_quant_lost_kg)

#quantities_own_consumption_subset <- Data %>% select(f_focus_own_consumption_measurement, #cal_focus_measurement_own_consumption, f_focus_own_consumption, cal_focus_quant_own_consumption_kg)

price_subset <- Data %>% select(f_focus_price, cal_focus_measurement_prod, cal_focus_price)
```

Tabulate some stats to check data quality, then make adjustments above in section 1

```{r}

# Create a boxplot of price
#boxplot(Data$cal_focus_price, ylab="Price of coffee per kg",
#        main="Boxplot of coffee price", col="red", las=1, cex.lab=1.2)

# Create a boxplot of quantity prod
#boxplot(Data$cal_focus_quant_prod_kg, ylab="quantity produced",
#        main="Boxplot of quantity produced", col="red", las=1, cex.lab=1.2)

# Create a boxplot of quantity sold
#boxplot(Data$cal_focus_quant_sold_kg, ylab="quantity sold",
#        main="Boxplot of quantity sold", col="red", las=1, cex.lab=1.2)

# Create a boxplot of quantity lost
#boxplot(Data$cal_focus_quant_lost_kg, ylab="quantity lost",
#        main="Boxplot of quantity lost", col="red", las=1, cex.lab=1.2)

# Filter the data and tabulate the uuid when price=0 but we have sales
#Data %>%
#  filter(cal_focus_price == 0 & cal_focus_quant_sold_kg > 0) %>%
#  select(uuid, cal_focus_price, cal_focus_quant_sold_kg)  # Select relevant columns

# Filter the data and tabulate the uuid when prod is lower than sales
#Data %>%
#  filter(cal_focus_quant_sold_kg > cal_focus_quant_prod_kg) %>%
#  select(uuid, f_focus_quant_prod, f_focus_measurement_prod, f_focus_quant_sold, f_focus_measurement_sold, f_focus_crop_size_hectare)  # Select relevant columns

#Data %>%
#  filter(cal_focus_quant_sold_kg == 0 & cal_focus_quant_prod_kg == 0) %>%
#  select(uuid, f_focus_crop_size_hectare, f_focus_quant_prod, f_focus_quant_sold, cal_focus_price)

#Data %>%
#  filter(cal_focus_quant_sold_kg == 0 & cal_focus_price > 0) %>%
#  select(uuid, cal_focus_price, cal_focus_quant_sold_kg)
```

## 3) Focus crop revenue calculation

check the revenue time period and harmonize if not

```{r}
# Check how many seasons there are to adjust the following code 
#table(Data$f_focus_rev_timeperiod)
```

```{r}
# #harmonise season if not
# Data <- Data %>%
#   mutate(f_focus_rev_timeperiod=ifelse(f_focus_rev_timeperiod=="season 1 timeframe", "season 1",
#                                        ifelse(f_focus_rev_timeperiod== "season 2 timefrmae", "season 2" ,
#                                               f_focus_rev_timeperiod)))
```

Calculate the revenue (globally and per season). make sure the name of season is updated (if not labeled in the same way) in the code below

```{r, include = FALSE}
Data <- Data %>%
  mutate(cal_focus_revenue = cal_focus_quant_sold_kg * cal_focus_price,
         cal_focus_revenue = replace_na(cal_focus_revenue, 0))
   

##Replace missing values in focus revenue by zero
Data <- Data %>%
  mutate(cal_focus_revenue = ifelse(is.na(cal_focus_revenue), 0, cal_focus_revenue)) 

Data <- Data %>%
  mutate(
    cal_focus_revenue = as.numeric(cal_focus_revenue))
```

## 4) Productivity calculation

Calculate the productivity variable by dividing the production by the farm size dedicated to the focus crop (in acre).

```{r, include = FALSE}

Data <- Data %>%
  
  mutate(across(c(cal_focus_quant_prod_kg, f_focus_crop_size_acre), as.numeric)) %>%
  mutate(f_focus_crop_size_acre = replace_na(f_focus_crop_size_acre, 0),
         cal_focus_productivity_acre = ifelse(f_focus_crop_size_acre > 0, cal_focus_quant_prod_kg/f_focus_crop_size_acre, NA))

Data <- Data %>%
  mutate(
    cal_focus_productivity_acre = as.numeric(cal_focus_productivity_acre))
```


## 6) Input costs calculation

Input costs are asked per season. This code sums up the costs for seasons to arrive at a total input cost.

```{r, include = FALSE}
#cost per season
Data <- Data %>%
  mutate_at(vars(starts_with("f_inputs_costs")), as.numeric) %>%
  mutate_at(vars(starts_with("f_inputs_costs")), ~ifelse(is.na(.), 0, .))
  
#check if some inputs variables are asked per land unit and make adjustments
# Data <- Data %>%
#   # Calculate inputs costs per season for those in per hectare
#   mutate(cal_inputs_costs_insect_season = f_inputs_costs_insect_season_per_hectare * f_focus_crop_size_hectare,
#          cal_inputs_costs_irrigation_season = f_inputs_costs_irrigation_season_per_hectare * f_focus_crop_size_hectare,
#          cal_inputs_costs_biological_control_season = f_inputs_costs_biological_control_season_per_hectare * f_focus_crop_size_hectare,
#          cal_inputs_costs_other_season = f_inputs_costs_other_season_per_hectare * f_focus_crop_size_hectare)

#calculate input costs
Data <- Data %>%
  mutate(
    cal_inputs_costs = rowSums(
      select(., starts_with("f_inputs_costs")) %>%
        mutate(across(everything(), ~as.numeric(.))),
      na.rm = TRUE
    )
  )

Data <- Data %>%
  mutate(
    cal_inputs_costs = as.numeric(cal_inputs_costs))
```


## 8) Total cost focus crop

@warning: outlier detection applied

```{r}
Data <- Data %>%
  mutate(
    cal_inputs_costs = as.numeric(cal_inputs_costs),
    f_focus_costs_other = as.numeric(f_focus_costs_other),
    f_focus_labour_costs_year = as.numeric(f_focus_labour_costs_year),

    cal_focus_cost = f_focus_labour_costs_year + cal_inputs_costs + f_focus_costs_other,
    cal_focus_cost = replace_na(cal_focus_cost, 0)
  )

```

## 9) Net-income focus crop

```{r}
Data <- Data %>%
  mutate(cal_focus_income = cal_focus_revenue - cal_focus_cost)

Data <- Data %>%
  mutate(
    cal_focus_income = as.numeric(cal_focus_income))
```

## 10) Net-income other crops

```{r}
# Clean data type and values from share variable
Data$f_othermaincrop_1_income_share <- as.numeric(Data$f_othermaincrop_1_income_share)
Data <- Data %>% mutate(f_othermaincrop_1_income_share = ifelse(f_othermaincrop_1_income_share == 10, 1, f_othermaincrop_1_income_share))

Data$f_othermaincrop_2_income_share <- as.numeric(Data$f_othermaincrop_2_income_share)
Data <- Data %>% mutate(f_othermaincrop_2_income_share = ifelse(f_othermaincrop_2_income_share == 10, 1, f_othermaincrop_2_income_share))
```

```{r}
# check for rows where cost is > 0 and income missing 
Data <- Data %>%
  mutate(f_othermaincrop_1_costs = ifelse(is.na(f_othermaincrop_1_costs), NA, f_othermaincrop_1_costs)) %>%
  mutate(f_othermaincrop2_costs = ifelse(is.na(f_othermaincrop2_costs), NA, f_othermaincrop2_costs)) 
```

```{r, include = FALSE}
# Define variables
variables_to_replace <- c("f_othermaincrop_1_costs", "f_othermaincrop2_costs")

# Convert to numeric and replace NAs with 0
Data <- Data %>%
  mutate(across(all_of(variables_to_replace), ~replace_na(as.numeric(.), 0)))

```



```{r}
Data <- Data %>%
  mutate(f_othermaincrop_1_inc_sold = replace_na(as.numeric(f_othermaincrop_1_inc_sold),0),
         f_othermaincrop_2_inc_sold = replace_na(as.numeric(f_othermaincrop_2_inc_sold),0))
```

Net income calculation:

```{r}
Data <- Data %>%
  mutate(cal_other_crop_income = f_othermaincrop_1_inc_sold - f_othermaincrop_1_costs + f_othermaincrop_2_inc_sold -  f_othermaincrop2_costs)

Data <- Data %>%
  mutate(
    cal_other_crop_income = as.numeric(cal_other_crop_income))

```

## 11) Net-income livestock calculation

```{r}
Data$f_livestock_income_share <- as.numeric(Data$f_livestock_income_share)

Data <- Data %>% mutate(f_livestock_income_share = ifelse(f_livestock_income_share == 10, 1, f_livestock_income_share))
```

check the available information that can be include in variable to replace, then adapt the code

```{r}
Data <- Data %>%
  mutate(
    cal_livestock_income = 
      as.numeric(f_livestock_revenue_total) - as.numeric(f_livestock_costs) +
      as.numeric(f_bee_revenue_total) - as.numeric(f_bee_costs)
  )

Data <- Data %>%
   mutate_at(vars(cal_livestock_income), ~replace_na(., 0))
```


## 12) Loan costs

--> agri loan cost only

```{r}
table(Data$hh_loan_agri_size)
```
```{r}
# Check whether farmers said a higher interest number compared to the total amount to be repaid
loan_mismatch_subset <- Data %>%
  filter(as.numeric(hh_loan_agri_paid_year) < as.numeric(hh_loan_interest_paid_year))

```


```{r}
variables_to_replace <- c("hh_loan_agri_size","hh_loan_interest_paid_year", "hh_loan_agri_paid_year")

# Convert to numeric and replace NAs with 0
Data <- Data %>%
  mutate(across(all_of(variables_to_replace), ~replace_na(as.numeric(.), 0)))

Data <- Data %>% 
  mutate(cal_farm_costs_loans = hh_loan_interest_paid_year) %>%
  mutate(cal_farm_costs_loans = ifelse(is.na(cal_farm_costs_loans), 0, cal_farm_costs_loans)) 

Data <- Data %>%
  mutate(
    cal_farm_costs_loans = as.numeric(cal_farm_costs_loans))
```


## 13) Off-farm and other income calculation

check and add any variable that should be considered

```{r}
variables_to_replace <- c("f_nonfarm_income_total")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(as.numeric(.), 0))

```

## 14) Farm income calculation

```{r}
Data <- Data %>%
  mutate(cal_farm_income = cal_focus_income + cal_other_crop_income + cal_livestock_income,
         cal_farm_income = replace_na(cal_farm_income, 0))

Data <- Data %>%
  mutate(
    cal_farm_income = as.numeric(cal_farm_income))
```

## 15) Actual income calculation


```{r}
# Clean data type and values from share variable
Data$f_focus_income_share <- as.numeric(Data$f_focus_income_share)
Data <- Data %>% mutate(f_focus_income_share = ifelse(f_focus_income_share == 10, 1, f_focus_income_share))
```

```{r}
Data <- Data %>% 
  # Second approach for actual income, based on the income share of farmers
  mutate(
    cal_actual_income_v2 = cal_focus_income * (10/(as.numeric(f_focus_income_share)))) %>%
  mutate(
    # Set cal_actual_income_v2 to missing (NA) if cal_focus_income is 0 or missing
    cal_actual_income_v2 = ifelse(is.na(cal_focus_income) | cal_focus_income == 0 | f_focus_quant_sold==0 | is.na(f_focus_quant_sold) | f_focus_measurement_sold==""  | f_focus_income_share == 0, NA, cal_actual_income_v2))

Data <- Data %>%
  mutate(
    cal_actual_income_v2 = as.numeric(cal_actual_income_v2))

```

```{r}
# make a check file to see whether everything went correct

subset <- Data %>% 
  select(cal_focus_quant_prod_kg, cal_focus_quant_sold_kg, cal_focus_price, cal_focus_quant_lost_kg, 
         cal_focus_revenue, cal_focus_productivity_acre, cal_inputs_costs, cal_focus_cost, cal_focus_income,cal_other_crop_income, cal_livestock_income, cal_actual_income_v2)
 
#cal_actual_income_v2 , cal_actual_income_v3)

write.xlsx(subset, "calculated_variables_script2.0.xlsx")
```

# Section 3 Descriptive statistics

## 1) Net Promotor score

```{r}
unique(Data$cs_recommendation)

```

```{r, include = FALSE}
hh_farmer_gender <- c("all farmers") 
    
    NSP_total <- Data %>%
      select(uuid, cs_recommendation) %>% 
      mutate(
        cs_recommendation = ifelse(is.na(cs_recommendation), "no input", cs_recommendation),
        detractor = as.numeric(cs_recommendation %in% c("Not likely", "Somewhat likely", "Likely")),
        promoter = as.numeric(cs_recommendation == "Very likely"),
        passive = as.numeric(cs_recommendation %in% c("Most likely", "I dont know")),
        nr_recommenders = as.numeric(cs_recommendation != "no input")
             ) %>%
      summarize(nr_promoters = sum(promoter),
                nr_detractors = sum(detractor),
                nr_passive = sum(passive),
                nr_farmers = sum(nr_recommenders)
                ) %>%
      mutate(
        nsp_total = round((nr_promoters/nr_farmers) - (nr_detractors/nr_farmers),2),
        hh_farmer_gender = hh_farmer_gender)
    
# NSP By gender 
NSP_by_gender <- Data %>%
  select(uuid, cs_recommendation, hh_farmer_gender) %>% 
  mutate(
    cs_recommendation = ifelse(is.na(cs_recommendation), "no input", cs_recommendation),
    detractor = as.numeric(cs_recommendation %in% c("Not likely", "Somewhat likely", "Likely")),
    promoter = as.numeric(cs_recommendation == "Very likely"),
    passive = as.numeric(cs_recommendation %in% c("Most likely", "I dont know")),
    nr_recommenders = as.numeric(cs_recommendation != "no input")
         ) %>%
  group_by(hh_farmer_gender) %>%
  summarize(nr_promoters = sum(promoter),
            nr_detractors = sum(detractor),
            nr_passive = sum(passive),
            nr_farmers = sum(nr_recommenders)) %>%
  mutate(nsp_gender = round((nr_promoters/nr_farmers) - (nr_detractors/nr_farmers),2))

NSP <- full_join(NSP_total, NSP_by_gender)
NSP <- NSP %>% select(hh_farmer_gender, nr_farmers, nr_promoters, nr_detractors,nr_passive, nsp_total, nsp_gender)
    
```

## 2) Anonymise data for further use

```{r, include = FALSE}
Data <- Data %>%
  rename(location_cascade_region = pi_location_cascade) %>%  ### the variable on the right differs from case to case!!!
  select(
    -c(starts_with("pi_"), -contains("county"), -one_of("pi_location_distance")), # Added -one_of("pi_location_distance")
    -c(submitter), -c(`_id`), -c(`_index`),
    -c(name_of_farmer)) %>%
  rename(pi_location_cascade_first_level = location_cascade_region)
```

## 3) Summary statistics

### All Farmers: numerical descriptives

```{r, include = FALSE}
# Get numeric-only columns
numeric_data <- Data %>% select(where(is.numeric))

# Create summary tables
mean_df <- numeric_data %>%
  summarise(across(everything(), ~round(mean(., na.rm = TRUE), 2))) %>%
  t() %>% as.data.frame() %>%
  rownames_to_column("variable") %>%
  rename(mean = V1)

sd_df <- numeric_data %>%
  summarise(across(everything(), ~round(sd(., na.rm = TRUE), 2))) %>%
  t() %>% as.data.frame() %>%
  rownames_to_column("variable") %>%
  rename(sd = V1)

min_df <- numeric_data %>%
  summarise(across(everything(), ~round(min(., na.rm = TRUE), 2))) %>%
  t() %>% as.data.frame() %>%
  rownames_to_column("variable") %>%
  rename(min = V1)

max_df <- numeric_data %>%
  summarise(across(everything(), ~round(max(., na.rm = TRUE), 2))) %>%
  t() %>% as.data.frame() %>%
  rownames_to_column("variable") %>%
  rename(max = V1)

freq_df <- numeric_data %>%
  summarise(across(everything(), ~sum(!is.na(.)))) %>%
  t() %>% as.data.frame() %>%
  rownames_to_column("variable") %>%
  rename(freq = V1)

# Merge all summaries
numerical_descriptives_all_farmers <- mean_df %>%
  left_join(sd_df, by = "variable") %>%
  left_join(min_df, by = "variable") %>%
  left_join(max_df, by = "variable") %>%
  left_join(freq_df, by = "variable") %>%
  select(variable, freq, mean, sd, min, max)

```

### All farmers: single categorical descriptives

```{r, include = FALSE}
# Label and total number of respondents
farmer_type <- "all farmers"
nr_participants_raw <- nrow(Data)  # count total participants

# Count total number of responses per variable (can be adjusted later)
all_var <- tibble(
  variable = NA_character_,
  `farmer type` = farmer_type,
  n = nr_participants_raw
)

# Identify all select_one questions
single_mc <- survey_questions %>%
  filter(str_starts(type, "select_one")) %>%
  pull(name)

# Initialize output table
single_categorical_descriptives_all_farmers <- tibble()

# Loop over each select_one variable
for (i in single_mc) {
  
  if (i %in% names(Data)) {
    
    frequencies <- Data %>%
      count(!!sym(i), name = "freq") %>%
      rename(category = !!sym(i)) %>%
      mutate(
        n = nr_participants_raw,
        `farmer type` = farmer_type,
        variable = i,
        `%` = round(freq / n * 100, 2),
        category = as.character(category)
      )
    
    single_categorical_descriptives_all_farmers <- single_categorical_descriptives_all_farmers %>%
      bind_rows(frequencies) %>%
      select(variable, `farmer type`, n, category, freq, `%`)
  }
}
```

### All farmers: multiple categorical descriptives

```{r, include = FALSE}
multiple_mc <- survey_questions %>%
  filter(grepl("^select_multiple", type)) %>%  # Filter select_multiple questions
  pull(name)  # Extract the variable names
multiple_categorical_descriptives_all_farmers <- all_var

for (j in multiple_mc) {
  print(j)
  
  # Skip the variable that do not work for this code (empty for example)
  # if (j == "f_labour_otheractivity_labour_type") {
  #   next
  # }
  # if (j == "hh_loan_one_agri_purpose") {
  #   next
  # }
  # if (j == "hh_loan_mulitple_largest_purpose") {
  #   next
  # }
  if (any(names(Data) %in% j)) {
    frequencies <- Data %>%
      select(all_of(j)) %>%
      cSplit(j, "|") %>%
      gather(key, value) %>%
      select(-key) %>%
      table() %>%
      melt(c("category"), value.name = "freq") %>%
      mutate(n = nr_participants_raw) %>%
      mutate("variable" = j) %>%
      left_join(all_var) %>%
      mutate("%" = round(freq / n, 2) * 100) %>%
      mutate(category = as.character(category))
    
    multiple_categorical_descriptives_all_farmers <- multiple_categorical_descriptives_all_farmers %>%
      full_join(frequencies) %>%
      select(variable, `farmer type`, n, category, freq, "%")
  }
}

```

### By gender: numerical descriptives

```{r, include=FALSE}
# Create gender summary (optional for joining later)
gender_var <- Data %>% 
  count(hh_farmer_gender, name = "n")

# Get only numeric columns
numeric_vars <- Data %>% select(where(is.numeric)) %>% names()

# Mean
mean_df <- Data %>%
  group_by(hh_farmer_gender) %>%
  summarise(across(all_of(numeric_vars), ~round(mean(.x, na.rm = TRUE), 2)), .groups = "drop") %>%
  pivot_longer(-hh_farmer_gender, names_to = "variable", values_to = "mean")

# SD
sd_df <- Data %>%
  group_by(hh_farmer_gender) %>%
  summarise(across(all_of(numeric_vars), ~round(sd(.x, na.rm = TRUE), 2)), .groups = "drop") %>%
  pivot_longer(-hh_farmer_gender, names_to = "variable", values_to = "sd")

# Min
min_df <- Data %>%
  group_by(hh_farmer_gender) %>%
  summarise(across(all_of(numeric_vars), ~round(min(.x, na.rm = TRUE), 2)), .groups = "drop") %>%
  pivot_longer(-hh_farmer_gender, names_to = "variable", values_to = "min")

# Max
max_df <- Data %>%
  group_by(hh_farmer_gender) %>%
  summarise(across(all_of(numeric_vars), ~round(max(.x, na.rm = TRUE), 2)), .groups = "drop") %>%
  pivot_longer(-hh_farmer_gender, names_to = "variable", values_to = "max")

# Frequency (non-NA count)
freq_df <- Data %>%
  group_by(hh_farmer_gender) %>%
  summarise(across(all_of(numeric_vars), ~sum(!is.na(.x))), .groups = "drop") %>%
  pivot_longer(-hh_farmer_gender, names_to = "variable", values_to = "freq")

# Merge all stats
numerical_descriptives_by_gender <- mean_df %>%
  left_join(sd_df, by = c("hh_farmer_gender", "variable")) %>%
  left_join(min_df, by = c("hh_farmer_gender", "variable")) %>%
  left_join(max_df, by = c("hh_farmer_gender", "variable")) %>%
  left_join(freq_df, by = c("hh_farmer_gender", "variable")) %>%
  left_join(gender_var, by = "hh_farmer_gender") %>%
  select(variable, hh_farmer_gender, n, freq, mean, sd, min, max)
```

### By gender: single categorical descriptives

```{r, include = FALSE}

single_mc <- survey_questions %>%
  filter(grepl("^select_one", type)) %>%  # Keep only single-choice questions
  filter(name != "hh_farmer_gender") %>%  # Exclude gender variable
  pull(name)  # Extract variable names

single_categorical_descriptives_by_gender <- gender_var

for(i in single_mc){
  
  if(any(names(Data) %in% i)){
    
    frequencies <- Data %>% 
      select(hh_farmer_gender, all_of(i)) %>% 
      table() %>% 
      melt(c("hh_farmer_gender","category"), value.name="freq") %>%
      mutate("variable" = i) %>%
      left_join(gender_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    single_categorical_descriptives_by_gender <- single_categorical_descriptives_by_gender %>% 
      full_join(frequencies) %>%
      select(variable, hh_farmer_gender, n, category, freq, "%")
  }
}
```

### By gender: multiple categorical descriptives

```{r, include = FALSE}

multiple_mc <- survey_questions %>%
  filter(grepl("^select_multiple", type)) %>%  # Filter select_multiple questions
  filter(name != "hh_farmer_gender") %>%  # Exclude gender variable
  pull(name)  # Extract the variable names

multiple_categorical_descriptives_by_gender <- gender_var

for(j in multiple_mc){
  print(j)
  #remove variable not working for this code
  # if (j == "f_labour_otheractivity_labour_type") {
  #   next
  # }
  # if (j == "hh_loan_mulitple_largest_purpose") {
  #   next
  # }
  if(any(names(Data) %in% j)){
    
    frequencies <- Data %>% 
      select(hh_farmer_gender, all_of(j)) %>% 
      cSplit(j, "|") %>%
      gather(key, value, -hh_farmer_gender) %>%
      select(-key) %>% table() %>%
      melt(c("hh_farmer_gender","category"), value.name="freq") %>%
      mutate("variable" = j) %>%
      left_join(gender_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    multiple_categorical_descriptives_by_gender <- multiple_categorical_descriptives_by_gender %>% 
      full_join(frequencies) %>%
      select(variable, hh_farmer_gender, n, category, freq, "%")
  }
}
```

# Section 4: Final checking

## 1) Check completeness of the codebook

All variables that are in the data should get a description in the codebook. That's why we check this and correct/ add where missing. We also check whether we have all the variables that are required for the data delivery.

```{r, include = FALSE}
# Adapt column names to expected structure
question_library <- question_library %>%
  dplyr::rename(
    label = `label::English`,
    name = name,
    type = type
  ) %>%
  mutate(across(everything(), tolower)) %>%
  select(label, name, type)

# Prepare list of unmatched items if needed
unmatched <- list()
```

## 2) Remove repeated rows

Remove repeated rows for household members and create columns for each household member (max household number is 15). Delete the unnecessary, only if household roster questions are asked, check also which ones are used. @Warning: to be updated properly with data

```{r, include = FALSE}
if("uuid" %in% colnames(Data)){     
    Data <- Data %>%
      group_by(uuid) %>%
      mutate(id = row_number())
}
#    for (nr in c(1:15)){
      
      ### ---- delete the unnecessary --- only if household roster questions are asked, check also which ones are used
#      hh_focuscrop_nr <- paste0("hh_focuscrop_", nr)
      
#      Data <- Data %>%
        #Delete the rows that do not apply
#        mutate(
#          !!sym(hh_focuscrop_nr) := ifelse(id == nr, hh_focuscrop_hrs, NA)) %>%
        
        #Delete the rows that do not apply
#        fill(!!sym(hh_focuscrop_nr)) %>%
#        fill(!!sym(hh_focuscrop_nr), .direction = "up") %>%
        
        #Delete the rows that do not apply
#        mutate(
#          !!sym(hh_focuscrop_nr) := ifelse(id >1 , NA, !!sym(hh_focuscrop_nr) ))
      
#    }
#  }
  

```

Let's check the variable names. First we need a dataframe with all variable names, and

```{r, include = FALSE}
#Make a dataframe listing all variables 
variable <-ls(Data)
data <- data.frame(variable)

## ---- Variables ending with other  ---

#Check if data variable is in codebook        
data$compare <-data$variable  %in%  survey_questions$name 

#if it's a "other variable", do not remove
data <- data %>%
  mutate(compare = ifelse(grepl('_other$',variable) & compare == FALSE , FALSE, TRUE)) %>%
  filter(compare == FALSE) 

#survey_questions <- bind_rows(survey_questions, data)

#survey_questions <- survey_questions %>%
#  arrange(variable)  %>%
#  fill(label) %>%
#  mutate(type = ifelse(!is.na(compare), "text",type)) %>%
  
#  mutate(variable_2 = str_remove(variable,'_other$')) %>%
#  mutate(same = ifelse(variable_2== lag(variable_2), 1, 0)) %>%
  
#  mutate(label = ifelse(!is.na(compare) & same == 0 , "no question available", label)) %>%
#  select(-compare, -variable_2, -same)
```

Check if it is a calculated variable or a variable from the question library.

```{r}
variable <-ls(Data)
data <- data.frame(variable)
data$compare <-data$variable  %in%  survey_questions$name 
data <- data %>%
  filter(compare == FALSE) %>%
  select(-compare)

data$compare <-data$variable  %in%  vars_transformed$variable 

#check whether there is a match by checking out the "data" dataset MANUALLY
#if a variable is not in the vars_transformed, update the excel with variable transformations
```

Append the survey questions with the calculated variables

```{r}
#One changes are applied, append the codebook with information from the calculated variables (vars_transformed)
library(dplyr)

# Assuming "data" and "vars_transformed" are your dataframes

vars_calculated <- data$variable

calculated_variables <- vars_transformed %>%
  filter(variable %in% vars_calculated)

# Delete columns and rename others in one step
calculated_variables <- calculated_variables %>%
  select(
    -transformation, # Delete the 'transformation' column
    -...4            # Delete the '...4' column
  ) %>%
  rename(
    name = variable,     # Rename 'variable' to 'name'
    label = description  # Rename 'description' to 'label'
  )

survey_questions <- survey_questions %>%
  bind_rows(calculated_variables)

```

Check whether its a household demographics variable

```{r}
#Check whether already in the codebook
variable <-ls(Data)
data <- data.frame(variable)
data$compare <-data$variable  %in%  survey_questions$variable 
data <- data %>%
  filter(compare == FALSE) 

```

If indeed household demographics variable, append the codebook with that information

```{r}
data <- data %>%
  filter(variable %in% household_demographics$variable) %>%
  select(-compare)

data <- merge(data, household_demographics, by.x = "variable")
survey_questions <- bind_rows(survey_questions, data)
```

```{r, include = FALSE}
# --- Check household demographic questions ---
  hh_demo <- survey_questions %>%
    filter(str_detect(name, 'hh_member_')) %>%
    arrange(name)

  survey_questions <- survey_questions %>%
    filter(!str_detect(name, 'hh_member_')) 
  survey_questions <- bind_rows(survey_questions, hh_demo)
  
```

Check whether we have all the required variables for the portal

```{r}
# ---- dashboard variables ----
  #All variables in delivered data into dataframe
  
  variable <-ls(Data)
  data <- data.frame(variable)
  
  #Check if data variable is dashboard list     
  vars_dashboard$compare <- vars_dashboard$latest_var %in% data$variable 
  
  data <- vars_dashboard %>%
    filter(!is.na(latest_var)) %>%
    filter(compare == FALSE)

```
```{r}

```

Prepare final list of unmatched vars

-   Check these by hand and ensure the variable names are aligned with

1)  the question library

2)  list of calculated variables

3)  Household demographic variables.

Ensure that all variables part of the cleaned data or dashboard variable names are described in the codebook.

```{r, include = FALSE}

variable <- ls(Data)
data <- data.frame(variable) %>%
  filter(
    !variable %in% c(survey_questions$name, vars_transformed$variable, question_library$name)
  )

```


## 4) Combine the inputs for data delivery

```{r}
# rename columns for portal purposes
Data <- Data %>%
  rename(
    identifier = uuid,
    repeat_no = id,
    `submission date` = endtime
  )

#make sure there is no scientific notation in the excel sheet
numeric_cols <- sapply(Data, is.numeric)
Data[numeric_cols] <- lapply(Data[numeric_cols], function(x) format(x, scientific = FALSE, trim = TRUE))


sets <- list(
    "Codebook" = survey_questions,
    "Answer options" = questions_choices,
    "Cleaned Data (anonymised)" = Data,
    "Raw Data (anonymised)" = Data_raw, 
    "Num. desc. all farmers" = numerical_descriptives_all_farmers,
    "Cat. desc. single all farmers" = single_categorical_descriptives_all_farmers,
    "Cat. desc. multi all farmers" = multiple_categorical_descriptives_all_farmers,
    "Num. desc. by gender" = numerical_descriptives_by_gender,
    "Cat. desc. single by gender" = single_categorical_descriptives_by_gender,
    "Cat. desc. multi by gender" = multiple_categorical_descriptives_by_gender,
    "Net promoter score" = NSP)
  
write.xlsx (sets, file = (paste0(data_delivery)))
```


## 6) Productivity check

Create this file the first time you run the script to clean the data.

```{r}
#Productivity check

productivity <- Data %>%
    mutate(focus_product_min_sold = cal_focus_quant_prod_kg - cal_focus_quant_sold_kg) %>%

    select(uuid,
           submitter,
           #`submission date`, ##MAKE SURE THIS VARIABLE IS NOT DELETED IN THE FIRST PIECE OF THE CODE.It should be deleted before sharing with IDH, but it should be kept when checking the productivity numbers.
           starts_with("pi_location_cascade_level"), #include the applicable variable, this differs per case
           f_unit_land,
           f_size,
           f_size_acre,
           f_focus_crop_size,
           f_focus_crop_size_acre,
           cal_focus_productivity_acre,
           f_focus_quant_prod,
           f_focus_measurement_prod,
           cal_focus_quant_prod_kg,
           f_focus_quant_sold,
           f_focus_measurement_sold,
           cal_focus_quant_sold_kg,
           focus_product_min_sold,
           f_focus_price,
           cal_focus_price,
           cal_focus_revenue,
           cal_focus_cost,
           cal_focus_income
           )
  
write.xlsx (productivity, file = "productivity_check.xlsx", overwrite = TRUE)

```
