---
title: "Data_clearning_Rwanda_endline"
output: html_document
date: "2024-03-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is developed to clean data from Farmfit cases in a standardized way.

### Preparation

Load the necessary packages.

```{r, include = FALSE}
library(here); library(readxl); library(openxlsx);
library(plyr); library(dplyr); library(tidyr); library(tidylog); library(tidyselect)
library(stringr); library(data.table); library(reshape2)
library(zoo);library(splitstackshape) ;library(tidyverse); library(magrittr)
```

Fill in the basic case information.

```{r}
case = "AIF"
year = "2024"
sdm_crop = "maize"
country = "Rwanda"
```

Make a new Github map with the project name where this R Markdown file is stored, together with the raw data and the survey form downloaded from Flow. Fill in the right file names in the code below so the variables data_filename and survey_filename refer to the correct file.

```{r}
data_filename <- "DATA_ANALYSIS-587310002-AIF-Rwanda.xlsx"
survey_filename <- "SURVEY_FORM-587310002-AIF-Rwanda.xlsx"
```

Describe the files that will be created in the script here.

```{r}
pi_filename <- paste("pi_",year,"_",case,"_",sdm_crop,".xlsx", sep="")
data_delivery <- paste(year,"_",case,"_anom_",sdm_crop,"_endline",".xlsx", sep = "")
```

Import the data files that can be found in the map Analytics-Guild -\> IDH -\> Templates and input files for data delivery. Copy them in the project map.

```{r}
question_library <- read_excel("question library format v4.1.1.xlsx",
                               sheet = "Full Survey", skip = 2) 
vars_transformed <- read_excel("variables with transformation_2.xlsx") 
household_demographics  <- read_excel("household_demographics_var_2.xlsx") 
vars_dashboard <- read_excel("variable names input dashboard_2.xlsx") 

```

Define functions that will be needed later in the script.

```{r}
# DATA TYPE Set factors to integers
factor_to_int <- function(x){
  as.numeric(as.character(x))
}
# Find and replace outliers
outlier_detection <- function(x){
  ifelse(
    x > (mean(x, na.rm=TRUE) + sd(x, na.rm=TRUE)*3) |
      x < (mean(x, na.rm=TRUE) - sd(x, na.rm=TRUE)*3), 
    9997,
    x
  )
}

options(scipen = 999)  
# count values without NA
count_n <- function(x){sum(!is.na(x))}
```

### Import data

We read data from Flow twice. On Data we will perform cleaning. Data_raw will be added to the data delivery as the full raw data file without transformations. Only the personal information variables are removed.

```{r, include = FALSE}
# Data for transformation
Data <- read_excel(data_filename)

# Raw data 
Data_raw <- read_excel(data_filename)
Data_raw <- Data_raw %>%
  select(-starts_with("pi_"), -name_of_farmer, -mobile_number_farmer)

# Read file with survey structure
survey_questions <- read_excel((survey_filename),sheet ="Full Survey", skip=2)
```


Adjust the survey file to be ready for analysis. We now have a clear overview of the survey questions asked in this case.

```{r, include = FALSE}
survey_questions %<>% 
  # Fill excel merged cells from the section
  mutate(Title = na.locf(Title, na.rm = FALSE)) %>%
  # Select right columns
  select("Title", "Variable name", "Text",
         "Question type", "Options",
         "Allow multiple") %>%
  # Trim trailing spaces
  mutate_if(is.character, str_trim) %>%
  # Rename columns
  rename("section" = "Title", 
         "variable" = "Variable name", 
         "question" = "Text",
         "type" = "Question type", 
         "options" = "Options",
         "multiple" = "Allow multiple") %>%
  # All variables to lowercase
  mutate_all(tolower) %>%
  # Filter the empty variables
  filter(!is.na(variable)) 
```

What are the numerical columns? How many participants?

```{r}
# Collect the different variables, numerical questions:
numerical_columns <- survey_questions %>%
  filter(type == "number") %>%
  select("variable") %>% pull()

## ---- Count initial number of participants
nr_participants_raw <- length(unique(Data$Identifier))
```


### Changes to dataset and codebook

First quick data cleaning steps:

-   Date format

-   Remove irrelevant columns

-   Names to lower case

-   Remove farmers that do not want to participate

-   Data types

-   "other" columns

```{r, include = FALSE}
Data <- Data %>%
  
  # Set the date to date format
  mutate(`Submission Date` = as.Date(`Submission Date`, format = "%d-%m-%Y")) %>%
  
  # Remove irrelevant columns
  # ONLY DELETE SUBMITTER AFTER THE PRODUCTIVITY CHECK! YOU WANT THIS VARIABLE IN FOR OULIER CHECKING
  select(-c(`Display Name`,`Device identifier`,`Instance`,
            #`Submitter`, 
            `Form version`)) %>%
  select(-contains("--option--")) %>%
  
  # All column names to lowercase
  rename_all(funs(tolower)) %>%
  
  # Rename option variables
  rename_all( funs(gsub("--other--", "_other2", ., perl=T))) %>%

  # All variables to lowercase
  mutate_all(tolower) %>%
  
  # Remove to farmers that didn't participate
  mutate(ic_informed_consent = na.locf(ic_informed_consent, na.rm = FALSE)) %>%
  filter(ic_informed_consent == "accepted to participate")  %>%
  
  # Numerical columns to numeric data type
  mutate_at(vars(numerical_columns), funs(as.numeric)) %>%
    
    #General case
    mutate(f_harvest_num = as.numeric(f_harvest_num)) %>%
    ## In some cases, the crop has been added to this variable name. Variable will then be f_[CROP]_harvest_num. This occurs when there are 2 focus crops in the survey. Adjustment is required in this case, like in the example below for a cocoa case. 
    #mutate(f_cocoa_harvest_num = as.numeric(f_cocoa_harvest_num)) %>% 
  
  # Remove all "other" option text ("other, please specify")
  # mutate_each(funs(str_remove(., "other, please specify"))) %>%
  mutate_if(is.character, funs(gsub("other, please specify", NA,.))) %>% 
  mutate_if(is.character, funs(gsub("\\|$","",.))) %>% 
  
  # Remove trailing spaces
  mutate_if(is.character, str_trim) %>% 
  
  # Remove punctuation and special characters - EXCEPT for "|"
  mutate_if(is.character, funs(gsub("[^\\|[:^punct:]]", "", ., perl=T)))
```

@NEW: FOR THIS CASE, WE NEED TO RUN THIS CODE BY HERE. IF THIS APPLIES TO YOUR CASE, PLEASE CHECK YOUR DATA MANUALLY BEFORE RUNNING, THERE MAY BE SOME VALUES THAT COULD BE CLEANED DIFFERENTELY
Put values that are 9999/"i don't know" or 9998/"i prefer not to say/9997 to NA.

```{r, include=FALSE}
Data <- Data %>%
    mutate_if(is.numeric, list(~na_if(., 9999))) %>%
    mutate_if(is.numeric, list(~na_if(., 9998))) %>%
    mutate_if(is.numeric, list(~na_if(., 9997))) %>%
    mutate_if(is.character, list(~na_if(., "i don't know"))) %>%
    mutate_if(is.character, list(~na_if(., "i prefer not to say"))) 
```


Adjust the **farm sizes** to acres.

```{r}
# First check what are the unit measurements reported in this case.
table(Data$f_unit_land)
table(Data$f_unit_land_other2)
```

The code below creates a variable for farm sizes in both acres and hectares. 
When another option is given as unit of measurement (other than acres or hectares), the code needs to be manually adjusted!

```{r, include = FALSE}
#First let's clean variable
Data <- Data %>%
    #clean: for those with more than 1000 and unit is not squared meters, replace by squared meters (for this case)
  mutate(f_unit_land=ifelse(f_size>=1000 & f_unit_land!="squared meters", "squared meters",
                            felse(f_size>=100 & f_size<1000 & f_unit_land!="hectares", "acres", f_unit_land)))

  
Data <- Data %>%
  mutate(f_focus_crop_size = ifelse(identifier == "rchpqf3qhavk" & f_focus_crop_size == 5, NA,
                          ifelse(identifier == "00hw14waq6wf" & f_focus_crop_size == 350, NA,
                                 ifelse(identifier == "0s7cm0x1xv7e" & f_focus_crop_size == 1000, NA,
                                        ifelse(identifier == "p51u7r5p7e80" & f_focus_crop_size == 10000, NA,
                                               ifelse(identifier == "s5sxqc3a88s3" & f_focus_crop_size == 50, f_size,
                                                      ifelse(identifier == "5ve877bbegyc" & f_focus_crop_size == 50, f_size,
                                                             ifelse(identifier == "w1537yqrybf7" & f_focus_crop_size == 50, f_size,
                                 f_focus_crop_size
                          ))))))))

```
  
```{r, include = FALSE}
# Mutate the different units to acres. We need a farm size in acres for the total farm size, and the size dedicated to focus crop.

Data <- Data %>%
  mutate(f_unit_land_other2=ifelse(f_unit_land_other2=="kg umunzani", "squared meters", f_unit_land_other2))%>%
  # Combine unit columns
  unite("f_unit_land", c(f_unit_land, f_unit_land_other2), 
        na.rm = TRUE, remove = FALSE, sep = " ") %>%
  # Convert land size to acres 
  mutate(f_size_acre = case_when(
    f_unit_land == "hectares" ~ f_size * 2.471,
    f_unit_land == "squared meters" ~ f_size / 4047,
    TRUE ~ f_size
  )) %>%
  
  # Convert focus crop size to acres
  mutate(f_focus_crop_size_acre = case_when(
    f_unit_land == "hectares" ~ f_focus_crop_size * 2.471,
    f_unit_land == "squared meters" ~ f_focus_crop_size / 4047,
    TRUE ~ f_focus_crop_size
  )) %>%
  
  # Fill missing values in focus crop size
  mutate(f_focus_crop_size_acre = na.locf(f_focus_crop_size_acre, na.rm = FALSE)) %>%
  
  # Mutate the different units to hectares.
  mutate(f_size_hectare = case_when(
    f_unit_land == "hectares" ~ f_size,
    f_unit_land == "squared meters" ~ f_size / 10000,
    TRUE ~ f_size_acre / 2.471
  )) %>%
  
  mutate(f_focus_crop_size_hectare = case_when(
    f_unit_land == "hectares" ~ f_focus_crop_size,
    f_unit_land == "squared meters" ~ f_focus_crop_size / 10000,
    TRUE ~ f_focus_crop_size_acre / 2.471
  ))
```



Mutate the birth year variables to **age.** Adjust the year to the current time. Not all birthyear variables are included in each case. Check the survey_questions table to see whether they are included in this case or not.

```{r, include = FALSE}
Data <- Data %>%
  mutate(cal_hh_farmer_age = 2024 - hh_farmer_birthyear,
           cal_hh_head_age = 2024-hh_head_birthyear) 
```

### Other variables

Identify the "other" variables, and combine the values with their parent. Then the "other" variables are removed.

```{r, include = FALSE}
variables_other <- Data %>% 
  select(ends_with("_other"), ends_with("_other2")) %>% 
  names() 

for(other in variables_other){
  
  if(any(names(Data) %in% gsub("_other$|_other2$", "", other))){
    
    # Fuzzy match? - many spelling mistakes in "other" questions
    
    # Combine "other" variables with their parents
    Data <- Data %>%
      unite(!!gsub("_other$|_other2$", "", other), 
            c(gsub("_other$|_other2$", "", other), all_of(other)), 
            sep="|", remove=FALSE, na.rm=TRUE) %>%
      select(-other)
  }
}
```

### Context specific adjustments

Check for obvious outliers first by running the full code EXCEPT for running the outlier function below. Download the productivity_check.xlsx. file that is created at the end of the script. Use the filter options and arrange columns from high to low, low to high, etc. to detect outliers manually. Also use the intake form! Highlight the cells you have questions about or that are obvious outliers. Consult the data collection supervisors to make adjustments. The changes in the data can be made in the code block below.

Note that you should keep the variable "submitter" for the time being so we can trace back which enumerator has entried that data. This variable is normally deleted in the script above (cmd+F "submitter"). Make sure you delete this variable in the end when sharing the data.

#### Adjustments for farm size

```{r}
# THIS IS EXAMPLE CODE! ADAPT THE CODE TO THE NECESSARY ADJUSTMENTS IN THE CASE, OR DELETE CODE IF NO CLEANING IS REQUIRED

Data <- Data %>%
  mutate(f_focus_crop_size = ifelse(identifier == "rchpqf3qhavk" & f_focus_crop_size == 5, NA,
                          ifelse(identifier == "00hw14waq6wf" & f_focus_crop_size == 350, NA,
                                 ifelse(identifier == "0s7cm0x1xv7e" & f_focus_crop_size == 1000, NA,
                                        ifelse(identifier == "p51u7r5p7e80" & f_focus_crop_size == 10000, NA,
                                               ifelse(identifier == "s5sxqc3a88s3" & f_focus_crop_size == 50, f_size,
                                                      ifelse(identifier == "5ve877bbegyc" & f_focus_crop_size == 50, f_size,
                                                             ifelse(identifier == "w1537yqrybf7" & f_focus_crop_size == 50, f_size,
                                 f_focus_crop_size
                          ))))))))

```


#### Focus crop measurement units

```{r}
#let's deal with missing units
Data <-Data %>%
  mutate(
    #production
    f_focus_measurement_prod = ifelse(identifier%in% c("fvn3mf5g10n1", "b2s069vt91fh","7xnk8jm61mje") ,"90 kg bag", 
                                      ifelse(identifier=="xxhn497d79e6","50 kg bag",
                                      ifelse(identifier %in% c("8v5hx685t58h", "yf160kfb4q9h") ,"130 kg bag",
                                      ifelse(identifier%in% c("89x7a8rsj9cf", "vpt6xjwsu7wd", "x4k0q51788un", "0n8uapbwwxh5", "660fypr41hx9", "a30w0ra75eq9") , "100 kg bag",
                                      ifelse(identifier=="qfcnwr2ye66s", "110 kg bag",
                                      ifelse(identifier %in% c("qtcusunudgav", "k7mrepywua1u"),"120 kg bag",
                                      ifelse(identifier=="bcc1wpjc6g88","140 kg bag",
                                      ifelse(identifier=="pj0ge1eqerrd","50 kg bag",
                                      ifelse(identifier %in% c("gmcp756u39qb","4r6hvfsmfyyq","cjthf5cjx5a1","ahj5ewk2phmd"),"kg",
                                      f_focus_measurement_prod))))))))),
    #sold
    f_focus_measurement_sold=ifelse(identifier%in%c("jmvxjaa8stw9","ewvufrv186ws","vr8ehbs0s6tc","0ncageb2s4ju"), "kg", f_focus_measurement_sold)
  )
```

```{r}
#Harmonise units
Data <- Data %>%
  mutate(
    #prod
    f_focus_measurement_prod = ifelse(f_focus_measurement_prod %in% c("100","100 bags kg","100 kg","100 kg bags","100 kgs","100kg","100kgs","10kg"),"100 kg bag",
                                      ifelse(f_focus_measurement_prod %in% c("120","120 kgs","120kg"),"120 kg bag",
                                             ifelse(f_focus_measurement_prod %in% c("130","130 bags of kg","130 kgs") , "130 kg bag",
                                                    ifelse(f_focus_measurement_prod %in% c("140","140 kgs"),"140 kg bag",
                                                           ifelse(f_focus_measurement_prod %in% c("150 kgs","1500 kg bags"),"150 kg bag",
                                            ifelse(f_focus_measurement_prod %in% c("1kg", "kilogram", "kilograms", "kg umunzani","umunzani"),"kg",
                                                   ifelse(f_focus_measurement_prod =="50kg bag","50 kg bag",
                                      f_focus_measurement_prod))))))),

    #sold
    f_focus_measurement_sold = ifelse(f_focus_measurement_sold %in% c("100","100 bags kg","100 kg","100 kg bags","100 kgs","100kg"),"100 kg bag",
                                      ifelse(f_focus_measurement_sold=="110kg","110 kg bag",
                                             ifelse(f_focus_measurement_sold %in% c("120 kgs","120kg"),"120 kg bag",
                                                    ifelse(f_focus_measurement_sold %in% c("130","130 bags of kg","130 kgs","130kg") , "130 kg bag",
                                                          ifelse(f_focus_measurement_sold %in% c("140","140 kgs"),"140 kg bag",
                                                                ifelse(f_focus_measurement_sold %in% c("150 kgs","1500 kg bags"),"150 kg bag",
                                                                       ifelse(f_focus_measurement_sold=="400 kg bags","40 kg bag",
                                            ifelse(f_focus_measurement_sold %in% c("1kg", "kilogram", "kilograms", "kg umunzani","umunzani","umunzani kg"),"kg",
                                                                 ifelse(f_focus_measurement_prod =="50kg bag","50 kg bag",
                                      f_focus_measurement_sold))))))))),

    #own consumption
    f_focus_own_consumption_measurement =ifelse(f_focus_own_consumption_measurement %in% c("100 kgs","100 kg bags"), "100 kg bag",
                                                ifelse(f_focus_own_consumption_measurement == "120kg", "120 kg bag",
                                                       ifelse(f_focus_own_consumption_measurement == "130 kgs", "130 kg bag",
                                                              ifelse(f_focus_own_consumption_measurement == "140 kgs", "140 kg bag",
                                                                     ifelse(f_focus_own_consumption_measurement %in% c("kilogram","kilograms", "umunzani"), "kg",
                                                                            ifelse(f_focus_measurement_prod =="50kg bag","50 kg bag",
                                                              f_focus_own_consumption_measurement)))))),
  )

# Missing unit for prod and sold but we have price in kg after checking (generally under 1000). we do unit =kg.
identifier_to_replace <- subset(Data, f_focus_measurement_prod=="" & f_focus_price <1000, select = identifier)

Data <- Data %>%
  mutate(f_focus_measurement_prod = ifelse(identifier %in% identifier_to_replace$identifier, "kg", f_focus_measurement_prod),
         f_focus_measurement_sold = ifelse(identifier %in% identifier_to_replace$identifier, "kg", f_focus_measurement_sold))

#missing unit sold but we have unit price in kg
Data <- Data %>%
  mutate(f_focus_measurement_sold = ifelse(identifier %in% c("ft2ejbjw2byq", "fkr72531ne7w"), "kg", f_focus_measurement_sold))
```  

#### Focus crop quantities

```{r}
#we have ingemeri as a unit, equivalent 1.5 kg. let's do unit=kg and quantity*1.5
Data <- Data %>%
  mutate(f_focus_quant_sold=ifelse(f_focus_measurement_sold=="ingemeri", f_focus_quant_sold * 1.5, f_focus_quant_sold),
         f_focus_measurement_sold=ifelse(f_focus_measurement_sold=="ingemeri", "kg", f_focus_measurement_sold)
  )

#convert high kg quantities into 100 kg bag quantities and harmonise units
#prod
identifier_to_replace <- subset(Data, f_focus_quant_prod >= 100 & f_focus_measurement_prod == "kg", select = identifier)


Data <- Data %>%
  mutate(f_focus_quant_prod = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_quant_prod / 100, f_focus_quant_prod),
         f_focus_measurement_prod = ifelse(identifier %in% identifier_to_replace$identifier, "100 kg bag", f_focus_measurement_prod)
         )%>%
  mutate(f_focus_quant_prod = ifelse(identifier =="yh9a5seyc1d0", f_focus_quant_prod * 25, f_focus_quant_prod),
         f_focus_measurement_prod = ifelse(identifier =="yh9a5seyc1d0", "kg", f_focus_measurement_prod)
         )

#sold
identifier_to_replace <- subset(Data, f_focus_quant_sold >= 100 & f_focus_measurement_sold == "kg", select = identifier)

Data <- Data %>%
  mutate(f_focus_quant_sold = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_quant_sold / 100, f_focus_quant_sold),
         f_focus_measurement_sold = ifelse(identifier %in% identifier_to_replace$identifier, "100 kg bag", f_focus_measurement_sold),
         #at the same time, let's harmonise price for 100 kg
         f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 100, f_focus_price)
         )

#Consumption
identifier_to_replace <- subset(Data, f_focus_own_consumption >= 100 & f_focus_own_consumption_measurement == "kg", select = identifier)

Data <- Data %>%
  mutate(f_focus_own_consumption = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_own_consumption / 100, f_focus_own_consumption),
         f_focus_own_consumption_measurement = ifelse(identifier %in% identifier_to_replace$identifier, "100 kg bag", f_focus_own_consumption_measurement)
         )

#losses
identifier_to_replace <- subset(Data, f_focus_quant_lost >= 100 & f_focus_measurement_lost == "kg", select = identifier)

Data <- Data %>%
  mutate(f_focus_quant_lost = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_quant_lost / 100, f_focus_quant_lost),
         f_focus_measurement_lost = ifelse(identifier %in% identifier_to_replace$identifier, "100 kg bag", f_focus_measurement_lost)
         )
```

```{r}
#in some cases we have price but not the quantity sold. but we have quantity sold in kg (which we convert in 100kg bag). so we replace sold=production. some price was given per kg. we then do times the coefficient of the unit (100 and 50)
identifier_to_replace <- subset(Data, f_focus_quant_sold ==0 & f_focus_price !=0, select = identifier)

Data <- Data %>%
  mutate(f_focus_quant_sold = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_quant_prod, f_focus_quant_sold),
         f_focus_measurement_sold = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_measurement_prod, f_focus_measurement_sold),
         f_focus_price= ifelse(identifier %in% identifier_to_replace$identifier & f_focus_measurement_sold=="100 kg bag", f_focus_price * 100, 
                               ifelse(identifier %in% identifier_to_replace$identifier & f_focus_measurement_sold=="50 kg bag", f_focus_price * 50, 
                                      #one of them had no problem with price, let's keep the original value
                                      ifelse(identifier =="9d1gd1u9pbk8", 60000, 
                               f_focus_price)))
         )
#outilier for quantity sold (100kg bag) and quant prod (low) in kg: price in line with kg, so we do quant sold=prod and unit = 100 kg bag
Data <- Data %>%
  mutate(f_focus_quant_sold = ifelse(identifier =="exns4dsmjdey", f_focus_quant_prod, f_focus_quant_sold),
         f_focus_measurement_prod = ifelse(identifier =="exns4dsmjdey", "100 kg bag", f_focus_measurement_prod))

#quantity sold > prod, let do regarding the crop size
Data <- Data %>%
  #sold<-prod
  mutate(f_focus_quant_sold = ifelse(identifier %in% c("2e00npfbragd", "7xnk8jm61mje", "rjurqexyvu2j", "xsuwh0vhs6cf", "a30w0ra75eq9"), f_focus_quant_prod, f_focus_quant_sold),
         #prod<-sold
         f_focus_quant_prod = ifelse(identifier %in% c("906y2hheaatr", "r7fhgdfkurju"), f_focus_quant_sold, f_focus_quant_prod),
         #sold unit=120
                  f_focus_measurement_sold = ifelse(identifier =="9djks2k9y7yh", "120 kg bag",
                                                    ifelse(identifier =="2e00npfbragd", "50 kg bag",
                                                    f_focus_measurement_sold)),
         f_focus_measurement_prod = ifelse(identifier =="exns4dsmjdey", "100 kg bag", f_focus_measurement_sold)

         )
```

#### Adjustments for price

```{r}
# THIS IS EXAMPLE CODE! ADAPT THE CODE TO THE NECESSARY ADJUSTMENTS IN THE CASE, OR DELETE CODE IF NO CLEANING IS REQUIRED
#We have farmers with quantities in 90 kg but price is for kg. more generally, data shows less than 1000 for price in kg bag. let's then correct the prices for 90 kg bag by multiplying by 90. we do the same for other units
#90 kg bag
identifier_to_replace <- subset(Data, f_focus_measurement_sold =="90 kg bag" & f_focus_price <=1000, select = identifier)
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 90, f_focus_price))

#100 kg bag
identifier_to_replace <- subset(Data, f_focus_measurement_sold =="100 kg bag" & f_focus_price <=1000, select = identifier)
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 100, f_focus_price))

# 120 kg bag
identifier_to_replace <- subset(Data, f_focus_measurement_sold =="120 kg bag" & f_focus_price <=1000, select = identifier)
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 120, f_focus_price))

# 130 kg bag
identifier_to_replace <- subset(Data, f_focus_measurement_sold =="130 kg bag" & f_focus_price <=1000, select = identifier)
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 130, f_focus_price))

# 140 kg bag
identifier_to_replace <- subset(Data, f_focus_measurement_sold =="140 kg bag" & f_focus_price <=1000, select = identifier)
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 140, f_focus_price))

# 40 kg bag
identifier_to_replace <- subset(Data, f_focus_measurement_sold =="40 kg bag" & f_focus_price <=1000, select = identifier)
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 40, f_focus_price))

# 50 kg bag
identifier_to_replace <- subset(Data, f_focus_measurement_sold =="50 kg bag" & f_focus_price <=1000, select = identifier)
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% identifier_to_replace$identifier, f_focus_price * 50, f_focus_price))

##//////////////////
#calculate median (this is used so far on the script)
median_price_kgs<- Data %>%
  filter(f_focus_measurement_sold == "kg") %>%
  summarize(median_price_kgs = median(f_focus_price, na.rm = TRUE)) 

#price not in line with the unit (kg)
Data <- Data %>%
  mutate(f_focus_price=ifelse(identifier=="reyqk7bbs7hd", median_price_kgs$median_price_kgs, f_focus_price))

# some changes were already made when dealing with the unit sold. 
#specific changees for some who already has prices for 100 kg, we'll keep the original value and for outlier, we replace by median value
median_price_100kg<- Data %>%
  filter(f_focus_measurement_sold == "100 kg bag") %>%
  summarize(median_price_100kg = median(f_focus_price, na.rm = TRUE))

Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier =="hmy0dx5tyndg", 20000, 
                                ifelse(identifier=="a30w0ra75eq9", 1000000,
                                      ifelse(identifier=="rvqudu81ud6a", 120000, 
                                            ifelse(identifier=="dedx7uxqqcwq", 325000, 
                                                   
                                                   #oultliers
                                                   ifelse(identifier %in% c("emj4key9puef","2e00npfbragd","k7mrepywua1u"), median_price_100kg$median_price_100kg,  
                                f_focus_price))))))

#Missing price, do median price
Data <- Data %>%
  mutate(f_focus_price = ifelse(identifier %in% c("cjv21cfx18h2", "kxx8dgdmaud2"), median_price_100kg$median_price_100kg, f_focus_price))
```


#### Other specific Adjustments 
```{r}

# if quant sold ==0 then unit==NA. same for prod
Data <- Data %>%
  mutate(f_focus_measurement_sold=ifelse(f_focus_quant_sold==0, NA_character_ , f_focus_measurement_sold),
         f_focus_measurement_prod=ifelse(f_focus_quant_prod==0, NA_character_ , f_focus_measurement_prod),
         )

```

### Handle outliers.

```{r, include = FALSE}

# OUTLIERS:
#Identify numerical columns again because of change in names above
numerical_columns <- survey_questions %>%
filter(type == "number") %>%
select("variable") %>% pull()

```

Only run the next chunk of code AFTER you did the productivity check! See explanation above.

```{r, include = FALSE}
Data <- Data %>%
    mutate_at(vars(numerical_columns), funs(outlier_detection))
```

Put values that are 9999/"i don't know" or 9998/"i prefer not to say/9997 to NA.

```{r, include=FALSE}
#NOTE: for this case, this function is moved to above
#Data <- Data %>%
#    mutate_if(is.numeric, list(~na_if(., 9999))) %>%
#    mutate_if(is.numeric, list(~na_if(., 9998))) %>%
#    mutate_if(is.numeric, list(~na_if(., 9997))) %>%
#    mutate_if(is.character, list(~na_if(., "i don't know"))) %>%
#    mutate_if(is.character, list(~na_if(., "i prefer not to say"))) 
```

### Some statistics

```{r}
# Number of participants who gave informed consent 
nr_participants_ic <- length(unique(Data$identifier))

# NUMBER OF SEASONS
number_of_seasons <- max(Data$f_harvest_num, na.rm=TRUE)
```

# Actual income calculation

### 1) Transform unit measurements

Each unit measurement variables needs to be aligned. We transform every variable to kg.

First check what are the unit measurement reported in this case. Use the code below, or check the Lumen dashboard.

```{r}
table(Data$f_focus_measurement_prod)
table(Data$f_focus_measurement_sold)
```

To do so, we extract the numerical values from the measurement units for quantity produced, sold, consumed and lost. Always check by hand whether this is going well. We extract the numbers from the measurement units to calculate total quantities. If there are other measurement units, not related to kg, put a value for those in the variable: cal_focus_measurement_prod

**Measurement unit production**

```{r, include = FALSE}
Data <- Data %>%
##Measurement unit for production
  mutate(cal_focus_measurement_prod := ifelse(grepl("[[:digit:]]", f_focus_measurement_prod), 
      extract_numeric(f_focus_measurement_prod), 
      ifelse(f_focus_measurement_prod %in%c("kg","kgs","kilogram","kilograms"), 1, 0)))
```

**Measurement unit sales**

```{r, include = FALSE}
Data <- Data %>%
##Measurement unit for sales focus crop
  mutate(cal_focus_measurement_sold := ifelse(
      grepl("[[:digit:]]", f_focus_measurement_sold), 
      extract_numeric(f_focus_measurement_sold), 
      ifelse(f_focus_measurement_sold %in%c("kg","kgs","kilogram","kilograms"), 1, 0)))

```

**Measurement unit focus crop lost**

```{r}
Data <- Data %>%
# measurement unit for lost
  mutate(cal_focus_lost_measurement := ifelse(
    grepl("[[:digit:]]", f_focus_measurement_lost), 
    extract_numeric(f_focus_measurement_lost), 
    ifelse(f_focus_measurement_lost %in%c("kg","kgs","kilogram","kilograms"), 1, 0)))
```

**Measurement unit focus crop own consumption**

```{r}
Data <- Data %>%
# measurement unit for lost
  mutate(cal_focus_measurement_own_consumption := ifelse(
    grepl("[[:digit:]]", f_focus_own_consumption_measurement), 
    extract_numeric(f_focus_own_consumption_measurement), 
    ifelse(f_focus_own_consumption_measurement %in%c("kg","kgs","kilogram","kilograms"), 1, 0))) 
```


### 2) Quantities produced, sold, lost and own consumption

In this section the quantities produced, sold, lost and own consumption are cleaned.
Now that the unit measurements are transformed, the quantities produced, sold, lost and own consumption can be calculated.


```{r, include = FALSE}
##Calculate the quantities focus crop
Data <- Data %>%
  mutate(cal_focus_quant_prod_kg = f_focus_quant_prod * cal_focus_measurement_prod) %>%
  mutate(cal_focus_quant_sold_kg = f_focus_quant_sold * cal_focus_measurement_sold) %>%
  mutate(cal_focus_quant_lost_kg = f_focus_quant_lost * cal_focus_lost_measurement) %>%
  mutate(cal_focus_quant_own_consumption_kg = f_focus_own_consumption * cal_focus_measurement_own_consumption)%>%
  mutate_at(vars(starts_with("cal_focus_quant")), ~replace_na(., 0)) 
    
  # also calculate the unit price 
Data <- Data %>%
  mutate(cal_focus_price = ifelse(cal_focus_measurement_prod != 0, f_focus_price/cal_focus_measurement_sold, f_focus_price),
    cal_focus_price = ifelse(is.na(cal_focus_quant_sold_kg) | cal_focus_quant_sold_kg == 0, NA, cal_focus_price))%>%
  mutate(cal_focus_price = replace_na(cal_focus_price, 0))
```

Check whether the calculations went well. We make a subset of the data with relevant variables. Click on the dataset to sort the columns to detect strange numbers and solve if necessary.

```{r, include = FALSE}

quantities_produced_subset <- Data %>% select(f_focus_measurement_prod, cal_focus_measurement_prod, f_focus_quant_prod, cal_focus_quant_prod_kg)

quantities_sold_subset <- Data %>% select(f_focus_measurement_sold, cal_focus_measurement_sold, f_focus_quant_sold, cal_focus_quant_sold_kg)

quantities_lost_subset <- Data %>% select(f_focus_measurement_lost, cal_focus_lost_measurement, f_focus_quant_lost, cal_focus_quant_lost_kg)

quantities_own_consumption_subset <- Data %>% select(f_focus_own_consumption_measurement, cal_focus_measurement_own_consumption, f_focus_own_consumption, cal_focus_quant_own_consumption_kg)

price_subset <- Data %>% select(f_focus_price, cal_focus_measurement_prod, cal_focus_price)
```

```{r}
# Check how many seasons there are to adjust the following code 
table(Data$f_focus_rev_timeperiod)
```

```{r}
Data <- Data %>%
  mutate(cal_focus_price = ifelse(cal_focus_quant_sold_kg > 0 & is.na(cal_focus_price), median_price_kgs$median_price_kgs, cal_focus_price))
```

### 3) Focus crop revenue

```{r, include = FALSE}
Data <- Data %>%
  mutate(cal_focus_revenue = cal_focus_quant_sold_kg * cal_focus_price) 
   
# Create extra columns to define the quantities per season
Data <- Data %>%
  # Create new columns based on the season specified in f_focus_rev_timeperiod
  mutate(
    cal_focus_quant_prod_kg_season_1 = ifelse(f_focus_rev_timeperiod == "season 1 timeframe", cal_focus_quant_prod_kg, NA),
    cal_focus_quant_prod_kg_season_2 = ifelse(f_focus_rev_timeperiod == "season 2 timefrmae", cal_focus_quant_prod_kg, NA),
    cal_focus_quant_prod_kg_whole_year = ifelse(f_focus_rev_timeperiod == "past 12 months whole year", cal_focus_quant_prod_kg, NA), #   
    cal_focus_quant_sold_kg_season_1 = ifelse(f_focus_rev_timeperiod == "season 1 timeframe", cal_focus_quant_sold_kg, NA),
    cal_focus_quant_sold_kg_season_2 = ifelse(f_focus_rev_timeperiod == "season 2 timefrmae", cal_focus_quant_sold_kg, NA),
    cal_focus_quant_sold_kg_whole_year = ifelse(f_focus_rev_timeperiod == "past 12 months whole year", cal_focus_quant_sold_kg, NA),
#
    cal_focus_quant_own_consumption_kg_season_1 = ifelse(f_focus_rev_timeperiod == "season 1 timeframe", cal_focus_measurement_own_consumption, NA),
    cal_focus_quant_own_consumption_kg_season_2 = ifelse(f_focus_rev_timeperiod == "season 2 timefrmae", cal_focus_measurement_own_consumption, NA),
    cal_focus_quant_own_consumption_kg_whole_year = ifelse(f_focus_rev_timeperiod == "past 12 months whole year", cal_focus_measurement_own_consumption, NA),
#
    cal_focus_quant_lost_kg_season_1 = ifelse(f_focus_rev_timeperiod == "season 1 timeframe", cal_focus_quant_lost_kg, NA),
    cal_focus_quant_lost_kg_season_2 = ifelse(f_focus_rev_timeperiod == "season 2 timefrmae", cal_focus_quant_lost_kg, NA),
    cal_focus_quant_lost_kg_whole_year = ifelse(f_focus_rev_timeperiod == "past 12 months whole year", cal_focus_quant_lost_kg, NA)
  )

##X) STANDARD APPROACH Inbetween, adjust for repeated question groups (IF APPLICABLE)  ----  
     ##Adress values for those that have multiple seasons

Data <- Data %>%
  #Identify duplicates
  group_by(identifier, f_focus_rev_timeperiod) %>% 
  mutate(count_dup = n()) %>% # Count duplicates within each group
  mutate(count_dup = ifelse(is.na(f_focus_rev_timeperiod), NA, count_dup)) %>% # Check whether we don't make mistakes here. For some cases, farmers enter twice wrongly, while in other cases entering the same values for 2 seasons is ok.
  
  #Identify those with multiple seasons
  group_by(identifier, f_focus_rev_timeperiod) %>% 
  mutate(count_seasons = n()) %>% # Count the number of seasons for each group
  mutate(count_seasons = ifelse(is.na(f_focus_rev_timeperiod) | count_seasons > 1, NA, count_seasons)) %>%

  # Sum the quantities for those that entered data for multiple seasons
  group_by(identifier) %>%
  mutate(
    cal_focus_quant_prod_kg_season_1 = sum(na.omit(cal_focus_quant_prod_kg_season_1), na.rm = TRUE),
    cal_focus_quant_prod_kg_season_2 = sum(na.omit(cal_focus_quant_prod_kg_season_2), na.rm = TRUE),
    cal_focus_quant_prod_kg_whole_year = sum(na.omit(cal_focus_quant_prod_kg_whole_year), na.rm = TRUE),
    cal_focus_quant_sold_kg_season_1 = sum(na.omit(cal_focus_quant_sold_kg_season_1), na.rm = TRUE),
    cal_focus_quant_sold_kg_season_2 = sum(na.omit(cal_focus_quant_sold_kg_season_2), na.rm = TRUE),
    cal_focus_quant_sold_kg_whole_year = sum(na.omit(cal_focus_quant_sold_kg_whole_year), na.rm = TRUE), 
    cal_focus_quant_own_consumption_kg_season_1 = sum(na.omit(cal_focus_quant_own_consumption_kg_season_1), na.rm = TRUE),
    cal_focus_quant_own_consumption_kg_season_2 = sum(na.omit(cal_focus_quant_own_consumption_kg_season_2), na.rm = TRUE),
    cal_focus_quant_own_consumption_kg_whole_year = sum(na.omit(cal_focus_quant_own_consumption_kg_whole_year), na.rm = TRUE),
    cal_focus_quant_lost_kg_season_1 = sum(na.omit(cal_focus_quant_lost_kg_season_1), na.rm = TRUE),
    cal_focus_quant_lost_kg_season_2 = sum(na.omit(cal_focus_quant_lost_kg_season_2), na.rm = TRUE),
    cal_focus_quant_lost_kg_whole_year = sum(na.omit(cal_focus_quant_lost_kg_whole_year), na.rm = TRUE)
  ) %>% 
  
  #Replace values with 0 for those that entered the data twice 
  mutate(`repeat no` = as.numeric(`repeat no`)) %>%
  mutate(cal_focus_quant_prod_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_prod_kg)) %>%
  mutate(cal_focus_quant_sold_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_sold_kg)) %>%
  mutate(cal_focus_quant_own_consumption_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_own_consumption_kg )) %>%
  mutate(cal_focus_quant_lost_kg = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_quant_lost_kg)) %>%
  mutate(cal_focus_revenue = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_focus_revenue)) %>%

  #Sum the quantities for those that entered data for multiple seasons
  group_by(identifier, count_seasons) %>%
  mutate(cal_focus_quant_prod_kg =  sum(cal_focus_quant_prod_kg)) %>%
  mutate(cal_focus_quant_sold_kg =  sum(cal_focus_quant_sold_kg)) %>%
  mutate(cal_focus_quant_own_consumption_kg =  sum(cal_focus_quant_own_consumption_kg)) %>%
  mutate(cal_focus_quant_lost_kg =  sum(cal_focus_quant_lost_kg)) %>%
  mutate(cal_focus_revenue = sum(cal_focus_revenue)) %>%

  #Ensure we have 1 row of data for each farmer
  mutate(cal_focus_quant_prod_kg = ifelse(`repeat no` >1 , NA, cal_focus_quant_prod_kg)) %>%
  mutate(cal_focus_quant_sold_kg = ifelse(`repeat no` >1  , NA, cal_focus_quant_sold_kg)) %>%
  mutate(cal_focus_quant_own_consumption_kg = ifelse(`repeat no` >1  , NA, cal_focus_quant_own_consumption_kg )) %>%
  mutate(cal_focus_quant_lost_kg = ifelse(`repeat no` >1 , NA, cal_focus_quant_lost_kg)) %>%
  mutate(cal_focus_revenue = ifelse (`repeat no`>1, NA, cal_focus_revenue)) %>%
  ungroup() %>%
  select(-count_dup, -count_seasons)

##Replace missing values in focus revenue by zero
Data <- Data %>%
  mutate(cal_focus_revenue = ifelse(is.na(cal_focus_revenue), 0, cal_focus_revenue)) 

```

### 4) Productivity

Calculate the productivity variable by dividing the production by the farm size dedicated to the focus crop.

```{r, include = FALSE}

Data <- Data %>%
  mutate(cal_focus_productivity_acre = cal_focus_quant_prod_kg/f_focus_crop_size_acre) 
```

### 5) Labour costs

Labour cost is asked per season. We need to sum the costs given for each season.

```{r}
#Labour cost per season
Data <- Data %>%
  mutate(cal_labour_cost = f_focus_labour_costs_season)

# FILL NA VALUES WITH 0
Data$cal_labour_cost <- Data$cal_labour_cost %>% replace(is.na(.), 0)
```

```{r}
table(Data$f_focus_labour_rev_timeperiod)
```

```{r, include = FALSE}

# Create extra columns to define the quantities per season
Data <- Data %>%
  # Create new columns based on the season specified in f_focus_labour_rev_timeperiod
  mutate(
    cal_labour_cost_season_1 = ifelse(f_focus_labour_rev_timeperiod == "season 1 insert timeframe", cal_labour_cost, NA),
    cal_labour_cost_season_2 = ifelse(f_focus_labour_rev_timeperiod == "season 2 insert timeframe", cal_labour_cost, NA),
    cal_labour_cost_season_3 = ifelse(f_focus_labour_rev_timeperiod == "season 3 insert timeframe", cal_labour_cost, NA))


## Inbetween, adjust for repeated questions groups (IF APPLICABLE) ----  
  
##Adress values for those that have multiple seasons
Data <- Data %>%
#Identify duplicates
group_by(identifier, f_focus_labour_rev_timeperiod) %>% mutate(count_dup = n()) %>% #check whether we don't make mistakes here. For some cases farmer enter twice wrongly, in other cases entering same values for 2 seasons is ok.
mutate(count_dup = ifelse(is.na(f_focus_labour_rev_timeperiod), NA, count_dup)) %>%
#Identify those with multiple seasons
group_by(identifier, f_focus_labour_rev_timeperiod) %>% mutate(count_seasons = n()) %>%
mutate(count_seasons = ifelse(is.na(f_focus_labour_rev_timeperiod) | count_seasons > 1, NA, count_seasons)) %>%

    #Sum the labour cost for those that entered data for multiple seasons
group_by(identifier) %>%
  mutate(
    cal_labour_cost_season_1 = sum(na.omit(cal_labour_cost_season_1), na.rm = TRUE),
    cal_labour_cost_season_2 = sum(na.omit(cal_labour_cost_season_2), na.rm = TRUE),
    cal_labour_cost_season_3 = sum(na.omit(cal_labour_cost_season_3), na.rm = TRUE)) %>%

#Replace values with 0 for those that entered the data twice
mutate(`repeat no` = as.numeric(`repeat no`)) %>%
mutate(cal_labour_cost = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_labour_cost)) %>%

#Sum the values for those that entered data for multiple seasons
group_by(identifier, count_seasons) %>%
mutate(cal_labour_cost =  sum(cal_labour_cost, na.rm = TRUE)) %>%

#Ensure we have 1 row of data for each farmer
mutate(cal_labour_cost = ifelse(`repeat no` >1 , NA, cal_labour_cost)) %>%
ungroup() %>% select(-count_dup, -count_seasons)
```

### 6) Input costs

Input costs are asked per season. This code sums up the costs for seasons to arrive at a total input cost. 


```{r, include = FALSE}
#cost per season
Data <- Data %>%
  mutate_at(vars(starts_with("f_inputs_costs")), as.numeric) %>%
  mutate_at(vars(starts_with("f_inputs_costs")), ~ifelse(is.na(.), 0, .))
  
Data <- Data %>%
  mutate(cal_inputs_costs = Data %>% 
           select(starts_with("f_inputs_costs")) %>%
           rowSums(na.rm=TRUE)) 
```

```{r}
table(Data$f_focus_inputs_rev_timeperiod)
```

```{r, include = FALSE}

# Create extra columns to define the quantities per season
Data <- Data %>%
  # Create new columns based on the season specified in f_focus_inputs_rev_timeperiod
  mutate(
    cal_inputs_costs_season_1 = ifelse(f_focus_inputs_rev_timeperiod == "season 1 insert timeframe", cal_inputs_costs, NA),
    cal_inputs_costs_season_2 = ifelse(f_focus_inputs_rev_timeperiod == "season 2 insert timeframe", cal_inputs_costs, NA),
    cal_inputs_costs_season_3 = ifelse(f_focus_inputs_rev_timeperiod == "season 3 insert timeframe", cal_inputs_costs, NA))

# In between ,adjust for RQG
Data <- Data %>%
  #Identify duplicates
  group_by(identifier, f_focus_inputs_rev_timeperiod) %>% mutate(count_dup = n()) %>%
  mutate(count_dup = ifelse(is.na(f_focus_inputs_rev_timeperiod), NA, count_dup)) %>%
  #Identify those with multiple seasons
  group_by(identifier, f_focus_inputs_rev_timeperiod) %>% mutate(count_seasons = n()) %>%
  mutate(count_seasons = ifelse(is.na(f_focus_inputs_rev_timeperiod) | count_seasons > 1, NA, count_seasons)) %>%
  
  #Sum the labour cost for those that entered data for multiple seasons
group_by(identifier) %>%
  mutate(
    cal_inputs_costs_season_1 = sum(na.omit(cal_inputs_costs_season_1), na.rm = TRUE),
    cal_inputs_costs_season_2 = sum(na.omit(cal_inputs_costs_season_2), na.rm = TRUE),
    cal_inputs_costs_season_3 = sum(na.omit(cal_inputs_costs_season_3), na.rm = TRUE)) %>%
  
  #Replace values with 0 for those that entered the data twice
  mutate(`repeat no` = as.numeric(`repeat no`)) %>%
  mutate(cal_inputs_costs = ifelse(`repeat no` >1  & count_dup > 1, 0, cal_inputs_costs)) %>%
  
  #Sum the values for those that entered data for multiple seasons
  group_by(identifier, count_seasons) %>%
  mutate(cal_inputs_costs =  sum(cal_inputs_costs, na.rm = TRUE)) %>%
  
  #Ensure we have 1 row of data for each farmer
  mutate(cal_inputs_costs = ifelse(`repeat no` >1 , NA, cal_inputs_costs)) %>%
  ungroup() %>%
  select(-count_dup, -count_seasons)
```


```{r}
input_costs_subset <- Data %>% select(c(f_inputs_costs_chemicals_season, f_inputs_costs_fertilizer_season, f_inputs_costs_seeds_season, cal_inputs_costs))
```


### 7) Other costs

```{r}
# Fill all NA values 
variables_to_replace <- c("f_focus_costs_irrigation", "f_focus_costs_energy",
                          "f_costs_land_focuscrop", "f_focus_costs_storage",
                          "f_focus_costs_marketing")

# Replace NA values with 0 for the specified variables using dplyr
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

# Calculation
Data <- Data %>%
  mutate(cal_focus_other_costs = f_focus_costs_irrigation + f_focus_costs_energy + f_costs_land_focuscrop + f_focus_costs_storage + f_focus_costs_marketing)
```

### 8) Total cost focus crop

```{r}
Data <- Data %>%
    mutate(cal_focus_cost = cal_labour_cost + cal_inputs_costs + cal_focus_other_costs) 
```

### 9) Net-income focus crop

```{r}
Data <- Data %>%
  mutate(cal_focus_income = cal_focus_revenue - cal_focus_cost)
```

### 10) Net-income other crops

```{r}
# Clean data type and values from share variable
Data$f_othermaincrop_1_income_share <- as.numeric(Data$f_othermaincrop_1_income_share)
Data <- Data %>% mutate(f_othermaincrop_1_income_share = ifelse(f_othermaincrop_1_income_share == 10, 1, f_othermaincrop_1_income_share))

Data$f_othermaincrop_2_income_share <- as.numeric(Data$f_othermaincrop_2_income_share)
Data <- Data %>% mutate(f_othermaincrop_2_income_share = ifelse(f_othermaincrop_2_income_share == 10, 1, f_othermaincrop_2_income_share))
```

```{r}
# check for rows where cost is >0 and income missing 
Data <- Data %>%
  mutate(f_othermaincrop_1_costs_labour = ifelse(is.na(f_othermaincrop_1_inc_sold), NA, f_othermaincrop_1_costs_labour)) %>%
  mutate(f_othermaincrop_1_costs_inputs = ifelse(is.na(f_othermaincrop_1_inc_sold), NA, f_othermaincrop_1_costs_inputs)) %>%
  mutate(f_othermaincrop_1_costs_transport = ifelse(is.na(f_othermaincrop_1_inc_sold), NA, f_othermaincrop_1_costs_transport)) %>%
  
  mutate(f_othermaincrop_2_costs_labour = ifelse(is.na(f_othermaincrop_2_inc_sold), NA, f_othermaincrop_2_costs_labour)) %>%
  mutate(f_othermaincrop_2_costs_inputs = ifelse(is.na(f_othermaincrop_2_inc_sold), NA, f_othermaincrop_2_costs_inputs)) %>%
  mutate(f_othermaincrop_2_costs_transport = ifelse(is.na(f_othermaincrop_2_inc_sold), NA, f_othermaincrop_2_costs_transport)) 
```

```{r, include = FALSE}
# Fill NA's 
variables_to_replace <- c("f_othermaincrop_1_costs_labour", "f_othermaincrop_1_costs_inputs",
                          "f_othermaincrop_1_costs_transport", "f_othermaincrop_2_costs_labour"                       ,"f_othermaincrop_2_costs_inputs", "f_othermaincrop_2_costs_transport")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

# calculation
Data <- Data %>%
  mutate(cal_othermaincrop_1_cost = f_othermaincrop_1_costs_labour + f_othermaincrop_1_costs_inputs + f_othermaincrop_1_costs_transport) %>%
  mutate(cal_othermaincrop_2_cost = f_othermaincrop_2_costs_labour + f_othermaincrop_2_costs_inputs + f_othermaincrop_2_costs_transport)

```

Following code calculates production value for other crops. This is only done in Living income studies! Check whether the code applies to this case.

```{r, include = FALSE}
#variables_to_replace <- c("f_othermaincrop_1_quant_prod", "f_othermaincrop_1_price",
#                          "f_othermaincrop_2_quant_prod", "f_othermaincrop_2_price")
#Data <- Data %>%
#  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

#Data <- Data %>%
#  mutate(cal_othermaincrop_production_value = f_othermaincrop_1_quant_prod * f_othermaincrop_1_price + f_othermaincrop_2_quant_prod #* f_othermaincrop_2_price)
```

Revenues from other crops:

```{r}
# Clean data type and values from share variable
Data$f_other_crop_income_share <- as.numeric(Data$f_other_crop_income_share)
Data <- Data %>% mutate(f_other_crop_income_share = ifelse(f_other_crop_income_share == 10, 1, f_other_crop_income_share))
```

```{r}
variables_to_replace <- c("f_othermaincrop_1_inc_sold", "f_othermaincrop_2_inc_sold",
                          "f_other_crop_income")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

Data <- Data %>%
  mutate(cal_othermaincrop_1_inc_sold = f_othermaincrop_1_inc_sold) %>%
  mutate(cal_othermaincrop_2_inc_sold = f_othermaincrop_2_inc_sold) %>%
  mutate(cal_other_crop_income = f_other_crop_income)
```

Net income calculation:

```{r}
Data <- Data %>%
  mutate(cal_other_crop_income = cal_othermaincrop_1_inc_sold - cal_othermaincrop_1_cost + cal_othermaincrop_2_inc_sold -  cal_othermaincrop_2_cost + cal_other_crop_income)
```

```{r}
# Subset to check calculations 
cal_otherinc_subset <- Data %>% select(c(cal_othermaincrop_1_inc_sold, cal_othermaincrop_1_cost, cal_othermaincrop_2_inc_sold, cal_othermaincrop_2_cost, cal_other_crop_income))
```

### 11) Net-income livestock
```{r}
Data$f_livestock_income_share <- as.numeric(Data$f_livestock_income_share)

Data <- Data %>% mutate(f_livestock_income_share = ifelse(f_livestock_income_share == 10, 1, f_livestock_income_share))
```
```{r}
Data <- Data %>% 
  mutate(f_livestock_income_total = ifelse(f_livestock_income_share == 0, NA, f_livestock_income_total))
```

```{r}
#variables_to_replace <- c("f_livestock_income_total", "f_livestock_nr_hired_labourers",
#                          "f_livestock_days_hiredlabour", "f_livestock_wages_hiredlabour",                               #"f_livestock_costs_other")
#Data <- Data %>%
#  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))
#Code for this cas
Data <- Data %>%
  mutate_at(vars(f_livestock_income_total), ~replace_na(., 0))

Data <- Data %>%
  mutate(cal_livestock_revenue = f_livestock_income_total) 
#%>%
#  mutate(cal_livestock_labour_cost = (f_livestock_nr_hired_labourers * f_livestock_days_hiredlabour * #f_livestock_wages_hiredlabour)) %>%
#  mutate(cal_livestock_costs_other = f_livestock_costs_other) %>%
#  mutate(cal_livestock_cost = cal_livestock_costs_other+ cal_livestock_labour_cost)

# Net income
Data <- Data %>%
  mutate(cal_livestock_income = cal_livestock_revenue
         #- cal_livestock_cost
         )

```

### 12) General farm costs


```{r}
#variables_to_replace <- c("f_equip_costs", "hh_loan_one_size",
                          #"hh_loan_one_value_interest", 
                          #"hh_loan_two_size"
                          #, "hh_loan_two_value_interest")

Data <- Data %>%
#  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))%>%
  mutate_at(vars(f_equip_costs), ~replace_na(., 0))

Data <- Data %>% 
  mutate(cal_equipment_costs = f_equip_costs) %>%
  mutate(cal_equipment_costs = ifelse(is.na(cal_equipment_costs), 0, cal_equipment_costs)) 
#%>%
  
#  mutate(loan_1_cost = ifelse(hh_loan_one_value_interest == 0, 0, hh_loan_one_size * (10/hh_loan_one_value_interest))) %>%
#  mutate(loan_2_cost = ifelse(hh_loan_two_value_interest == 0, 0, hh_loan_two_size * (10/hh_loan_two_value_interest))) %>%
  
#  mutate(cal_farm_costs_loans = loan_1_cost + loan_2_cost) %>%
#  mutate(cal_farm_costs_loans = ifelse(is.na(cal_farm_costs_loans), 0, cal_farm_costs_loans)) 

# total costs 
Data <- Data %>% mutate(cal_farm_costs_general = cal_equipment_costs
                        #+ cal_farm_costs_loans
                        )

# Delete intermediate variables 
#Data <- Data %>% select(-c(loan_1_cost, loan_2_cost))
```

### 13) Off-farm and other income

```{r}
#"f_income_other_month"
variables_to_replace <- c("f_nonfarm_enterpr_income_month", "f_income_offfarmlabour_month")
Data <- Data %>%
  mutate_at(vars(all_of(variables_to_replace)), ~replace_na(., 0))

Data <- Data %>%
  mutate(cal_offfarm_income = 12*(f_nonfarm_enterpr_income_month + f_income_offfarmlabour_month))
```

### 14) Farm income calculation

```{r}
#Data <- Data %>%
#  mutate(cal_farm_income = cal_focus_income + cal_other_crop_income + cal_livestock_income - cal_farm_costs_general)
```

### 15) Actual income calculation

There are different approaches to calculating the actual income. The first version is the default. Use this approach, unless communicated otherwise. 

```{r}
#Data <- Data %>% 
  # First approach for actual income, based on as many direct inputed values of the farmer.
#  mutate(
#    cal_actual_income = cal_focus_income + cal_other_crop_income + cal_livestock_income - cal_farm_costs_general + #cal_offfarm_income) %>%
#  mutate(
    # Set cal_actual_income to missing (NA) if cal_focus_income is 0 or missing
#    cal_actual_income = ifelse(is.na(cal_focus_income) | cal_focus_income == 0 | f_focus_quant_sold==0 | is.na(f_focus_quant_sold) #| f_focus_measurement_sold=="" , NA, cal_actual_income))
```

```{r}
# Clean data type and values from share variable
#Data$f_focus_income_share <- as.numeric(Data$f_focus_income_share)
#Data <- Data %>% mutate(f_focus_income_share = ifelse(f_focus_income_share == 10, 1, f_livestock_income_share))
```

```{r}
#Data <- Data %>% 
  # Second approach for actual income, based on the income share of farmers
#  mutate(
#    cal_actual_income_v2 = cal_focus_income * (10/(as.numeric(f_focus_income_share)))) %>%
#  mutate(
    # Set cal_actual_income_v2 to missing (NA) if cal_focus_income is 0 or missing
#    cal_actual_income_v2 = ifelse(is.na(cal_focus_income) | cal_focus_income == 0 | f_focus_quant_sold==0 | #is.na(f_focus_quant_sold) | f_focus_measurement_sold==""  | f_focus_income_share == 0, NA, cal_actual_income_v2))

```

```{r}
# make a check file to see whether everything went correct

subset <- Data %>% 
  select(cal_focus_quant_prod_kg, cal_focus_quant_sold_kg, cal_focus_price, 
         cal_focus_quant_own_consumption_kg, cal_focus_quant_lost_kg, 
         cal_focus_revenue, cal_focus_productivity_acre,cal_labour_cost, cal_inputs_costs, cal_focus_other_costs, cal_focus_cost, cal_focus_income, cal_othermaincrop_1_cost, cal_othermaincrop_2_cost, #cal_othermaincrop_production_value,
cal_othermaincrop_1_inc_sold, cal_othermaincrop_2_inc_sold,cal_other_crop_income, cal_livestock_income, cal_equipment_costs,#cal_farm_costs_loans, 
cal_farm_costs_general, cal_offfarm_income)
#cal_actual_income, 
#cal_actual_income_v2 , cal_actual_income_v3)

write.xlsx(subset, "calculated_variables_script2.0.xlsx")
```

## Net Promotor score

```{r, include = FALSE}
hh_farmer_gender <- c("all farmers")
    
    NSP_total <- Data %>%
      select(`repeat no`, cs_recommendation) %>% 
      filter(`repeat no` == 1)%>%
      mutate(cs_recommendation = ifelse(is.na(cs_recommendation), "no input",cs_recommendation)) %>%
      mutate(detractor = ifelse(cs_recommendation == "not likely"| cs_recommendation == "somewhat likely"| cs_recommendation == "likely",1,0),
             promoter = ifelse(cs_recommendation == "very likely", 1,0),
             passive = ifelse(cs_recommendation == "most likely"| cs_recommendation == "i dont know",1,0),
             nr_recommenders = ifelse(cs_recommendation == "no input", 0,1)) %>%
      summarize(nr_promoters = sum(promoter),
                nr_detractors = sum(detractor),
                nr_passive = sum(passive),
                nr_farmers = sum(nr_recommenders)) %>%
      mutate(nsp_total = round((nr_promoters/nr_farmers) - (nr_detractors/nr_farmers),2),
             hh_farmer_gender = hh_farmer_gender)
    
    NSP_by_gender <- Data %>%
      select(`repeat no`, cs_recommendation, hh_farmer_gender) %>% 
      filter(`repeat no` == 1)%>%
      mutate(cs_recommendation = ifelse(is.na(cs_recommendation), "no input",cs_recommendation)) %>%
      mutate(detractor = ifelse(cs_recommendation == "not likely"| cs_recommendation == "somewhat likely"| cs_recommendation == "likely",1,0),
             promoter = ifelse(cs_recommendation == "very likely", 1,0),
             passive = ifelse(cs_recommendation == "most likely"| cs_recommendation == "i dont know",1,0),
             nr_recommenders = ifelse(cs_recommendation == "no input", 0,1)) %>%
      group_by(hh_farmer_gender) %>%
      summarize(nr_promoters = sum(promoter),
                nr_detractors = sum(detractor),
                nr_passive = sum(passive),
                nr_farmers = sum(nr_recommenders)) %>%
      mutate(nsp_gender = round((nr_promoters/nr_farmers) - (nr_detractors/nr_farmers),2))
    
    NSP <- full_join(NSP_total, NSP_by_gender)
    NSP <- NSP %>% select(hh_farmer_gender, nr_farmers, nr_promoters, nr_detractors,nr_passive, nsp_total, nsp_gender)
    
```

## Anonymise data for further use

```{r, include = FALSE}
Data <- Data %>%
          rename(location_cascade_region = pi_location_cascade_district) %>%  ### the variable on the right differs from case to case!!!
          select(
      -c(starts_with("pi_"), -contains("county")),
      -c(name_of_farmer)) %>%
          rename(pi_location_cascade_first_level = location_cascade_region)
```


# Summary statistics

### All Farmers: numerical descriptives

```{r, include = FALSE}
farmer_type <- c("all farmers")
all_var <- Data %>% 
  filter(`repeat no` == 1) %>%
  select(`repeat no`) %>% 
  table() %>% 
  melt(c("farmer type"), value.name="n") %>%
  mutate(`farmer type` = farmer_type)

numerical_descriptives_all_farmers <- Data %>%
  select_if(is.numeric) %>%
  summarise_each(funs(round(mean(., na.rm = TRUE),2))) %>%
  melt(value.name="mean") %>%
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(round(sd(., na.rm = TRUE),2))) %>%
      melt(value.name="sd")) %>%
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(min(., na.rm = TRUE))) %>%
      melt(value.name="min")
  ) %>% 
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(round(max(., na.rm = TRUE),2))) %>%
      melt(value.name="max")
  ) %>% 
  left_join(
    Data %>% 
      select_if(is.numeric) %>%
      summarise_each(funs(count_n))  %>%
      melt(value.name="freq")
  ) %>%
  left_join(survey_questions) %>%
  select("variable", "freq", "mean","sd", "min","max") 

```

### All farmers: single categorical descriptives

```{r, include = FALSE}
single_mc <- survey_questions %>% 
  filter(type=="option") %>% 
  filter(is.na(multiple)) %>% 
  pull(variable)

single_categorical_descriptives_all_farmers <- all_var

for(i in single_mc){
  
  if(any(names(Data) %in% i)){
    
    frequencies <- Data %>% 
      filter(`repeat no`==1) %>%
      select(all_of(i)) %>%       
      table() %>% 
      melt(c("category"),value.name = "freq") %>%
      mutate(n = nr_participants_raw) %>%
      mutate(`farmer type` = farmer_type) %>%
      mutate("variable" = i) %>%
      left_join(all_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    single_categorical_descriptives_all_farmers <- single_categorical_descriptives_all_farmers %>% 
      full_join(frequencies) %>%
      select(variable, `farmer type`, n, category, freq, "%")
    
  }
}
```

### All farmers: multiple categorical descriptives

```{r, include = FALSE}
multiple_mc <- survey_questions %>%
  filter(type == "option") %>%
  filter(multiple == "yes") %>%
  pull(variable)

multiple_categorical_descriptives_all_farmers <- all_var

for (j in multiple_mc) {
  print(j)
  
  # Skip the variable "f_labour_otheractivity_labour_type"
  if (j == "f_labour_otheractivity_labour_type") {
    next
  }
  if (j == "hh_loan_one_agri_purpose") {
    next
  }
  if (j == "hh_loan_mulitple_largest_purpose") {
    next
  }
  if (any(names(Data) %in% j)) {
    frequencies <- Data %>%
      select(all_of(j)) %>%
      cSplit(j, "|") %>%
      gather(key, value) %>%
      select(-key) %>%
      table() %>%
      melt(c("category"), value.name = "freq") %>%
      mutate(n = nr_participants_raw) %>%
      mutate("variable" = j) %>%
      left_join(all_var) %>%
      mutate("%" = round(freq / n, 2) * 100) %>%
      mutate(category = as.character(category))
    
    multiple_categorical_descriptives_all_farmers <- multiple_categorical_descriptives_all_farmers %>%
      full_join(frequencies) %>%
      select(variable, `farmer type`, n, category, freq, "%")
  }
}

```

### By gender: numerical descriptives

```{r, include=FALSE}
gender_var <- Data %>% 
  select(hh_farmer_gender) %>% 
  table() %>% 
  melt("hh_farmer_gender", value.name="n")

numerical_descriptives_by_gender <- Data %>%
  group_by(hh_farmer_gender) %>%
  select_if(is.numeric) %>%
  summarise_each(funs(round(mean(., na.rm = TRUE),2))) %>%
  melt(value.name="mean") %>%
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(round(sd(., na.rm = TRUE),2))) %>%
      melt(value.name="sd")) %>%
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(min(., na.rm = TRUE))) %>%
      melt(value.name="min")
  ) %>% 
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(round(max(., na.rm = TRUE),2))) %>%
      melt(value.name="max")
  ) %>% 
  left_join(
    Data %>% 
      group_by(hh_farmer_gender) %>%
      select_if(is.numeric) %>%
      summarise_each(funs(count_n))  %>%
      melt(value.name="freq")
  ) %>%
  left_join(survey_questions) %>%
  left_join(gender_var) %>%
  select("variable","hh_farmer_gender","n", "freq", "mean","sd", "min","max") 

```

### By gender: single categorical descriptives

```{r, include = FALSE}
single_mc <- survey_questions %>% 
  filter(type=="option") %>% 
  filter(is.na(multiple)) %>% 
  filter(variable != "hh_farmer_gender") %>%
  pull(variable)

single_categorical_descriptives_by_gender <- gender_var

for(i in single_mc){
  
  if(any(names(Data) %in% i)){
    
    frequencies <- Data %>% 
      select(hh_farmer_gender, all_of(i)) %>% 
      table() %>% 
      melt(c("hh_farmer_gender","category"), value.name="freq") %>%
      mutate("variable" = i) %>%
      left_join(gender_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    single_categorical_descriptives_by_gender <- single_categorical_descriptives_by_gender %>% 
      full_join(frequencies) %>%
      select(variable, hh_farmer_gender, n, category, freq, "%")
  }
}
```

### By gender: multiple categorical descriptives

```{r, include = FALSE}
multiple_mc <- survey_questions %>% 
  filter(type=="option") %>% 
  filter(multiple == "yes") %>%
  pull(variable)

multiple_categorical_descriptives_by_gender <- gender_var

for(j in multiple_mc){
  print(j)
  if (j == "f_labour_otheractivity_labour_type") {
    next
  }
  if (j == "hh_loan_mulitple_largest_purpose") {
    next
  }
  if(any(names(Data) %in% j)){
    
    frequencies <- Data %>% 
      select(hh_farmer_gender, all_of(j)) %>% 
      cSplit(j, "|") %>%
      gather(key, value, -hh_farmer_gender) %>%
      select(-key) %>% table() %>%
      melt(c("hh_farmer_gender","category"), value.name="freq") %>%
      mutate("variable" = j) %>%
      left_join(gender_var) %>%
      mutate("%" = round(freq/n,2)*100) %>%
      mutate(category = as.character(category))
    
    multiple_categorical_descriptives_by_gender <- multiple_categorical_descriptives_by_gender %>% 
      full_join(frequencies) %>%
      select(variable, hh_farmer_gender, n, category, freq, "%")
  }
}
```

# Check completeness of the codebook

All variables that are in the data should get a description in the codebook. That's why we check this and correct/ add where missing. We also check whether we have all the variables that are required for the data delivery.

```{r, include = FALSE}
question_library <- question_library %>%
    dplyr::rename(section = "Title",
                  question = "Text", 
                  variable = "Variable name",
                  type = "Question type",
                  options = "Options",
                  multiple = "Allow multiple") %>%
    mutate_all(tolower) %>%
    select(section,
           question,
           variable,
           type,
           options,
           multiple) %>%
    fill(section)
  
unmatched <- list()
```

# Remove repeated rows

Remove repeated rows for household members and create columns for each household member (max household number is 15). Delete the unnecessary, only if household roster questions are asked, check also which ones are used.

```{r, include = FALSE}
if("identifier" %in% colnames(Data)){     
    Data <- Data %>%
      group_by(identifier) %>%
      mutate(id = row_number())
    
    for (nr in c(1:15)){
      
      ### ---- delete the unnecessary --- only if household roster questions are asked, check also which ones are used
      hh_focuscrop_nr <- paste0("hh_focuscrop_", nr)
      
      Data <- Data %>%
        #Delete the rows that do not apply
        mutate(
          !!sym(hh_focuscrop_nr) := ifelse(id == nr, hh_focuscrop_hrs, NA)) %>%
        
        #Delete the rows that do not apply
        fill(!!sym(hh_focuscrop_nr)) %>%
        fill(!!sym(hh_focuscrop_nr), .direction = "up") %>%
        
        #Delete the rows that do not apply
        mutate(
          !!sym(hh_focuscrop_nr) := ifelse(id >1 , NA, !!sym(hh_focuscrop_nr) ))
      
    }
}
  Data <- Data %>%
    filter(id == 1) %>%
    select(-id,
           #-hh_member_birthyear,
           #-hh_member_gender,
           #-hh_member_education,
           
           -hh_focuscrop_hrs)
  
```

Let's check the variable names. First we need a dataframe with all variable names, and

```{r, include = FALSE}
#Make a dataframe listing all variables 
variable <-ls(Data)
data <- data.frame(variable)

## ---- Variables ending with other  ---

#Check if data variable is in codebook        
data$compare <-data$variable  %in%  survey_questions$variable 

#if it's a "other variable", do not remove
data <- data %>%
  mutate(compare = ifelse(grepl('_other$',variable) & compare == FALSE , FALSE, TRUE)) %>%
  filter(compare == FALSE) 

survey_questions <- bind_rows(survey_questions, data)
survey_questions <- survey_questions %>%
  arrange(variable)  %>%
  fill(question,
       section) %>%
  mutate(type = ifelse(!is.na(compare), "free text",type)) %>%
  
  mutate(variable_2 = str_remove(variable,'_other$')) %>%
  mutate(same = ifelse(variable_2== lag(variable_2), 1, 0)) %>%
  
  mutate(section = ifelse(!is.na(compare) & same == 0 , "unknown", section),
         question = ifelse(!is.na(compare) & same == 0 , "no question available", question)) %>%
  select(-compare, -variable_2, -same)
```

Check if it is a calculated variable or a variable from the question library.

```{r}
variable <-ls(Data)
data <- data.frame(variable)
data$compare <-data$variable  %in%  survey_questions$variable 
data <- data %>%
  filter(compare == FALSE) %>%
  select(-compare)

data$compare <-data$variable  %in%  vars_transformed$variable 

#check whether there is a match by checking out the "data" dataset MANUALLY
#if a variable is not in the vars_transformed, update the excel with variable transformations
```

Append the survey questions with the calculated variables

```{r}
#One changes are applied, append the codebook with information from the calculated variables (vars_transformed)
library(dplyr)

# Assuming "data" and "vars_transformed" are your dataframes

vars_calculated <- data$variable

calculated_variables <- vars_transformed %>%
  filter(variable %in% vars_calculated)

survey_questions <- survey_questions %>%
  bind_rows(calculated_variables)

```

Check whether its a household demographics variable

```{r}
#Check whether already in the codebook
variable <-ls(Data)
data <- data.frame(variable)
data$compare <-data$variable  %in%  survey_questions$variable 
data <- data %>%
  filter(compare == FALSE) 

```

If indeed household demographics variable, append the codebook with that information

```{r}
data <- data %>%
  filter(variable %in% household_demographics$variable) %>%
  select(-compare)

data <- merge(data, household_demographics, by.x = "variable")
survey_questions <- bind_rows(survey_questions, data)
```

```{r, include = FALSE}
# --- Check household demographic questions ---
  hh_demo <- survey_questions %>%
    filter(str_detect(variable, 'hh_member_')) %>%
    arrange(variable) %>%
    fill(section, options)
  
  survey_questions <- survey_questions %>%
    filter(!str_detect(variable, 'hh_member_')) 
  survey_questions <- bind_rows(survey_questions, hh_demo)
  
```

Check whether we have all the required variables for the portal

```{r}
# ---- dashboard variables ----
  #All variables in delivered data into dataframe
  
  variable <-ls(Data)
  data <- data.frame(variable)
  
  #Check if data variable is dashboard list     
  vars_dashboard$compare <- vars_dashboard$latest_var %in% data$variable 
  
  data <- vars_dashboard %>%
    filter(!is.na(latest_var)) %>%
    filter(compare == FALSE) %>%
    filter(is.na(note))
  

```

Prepare final list of unmatched vars

-   Check these by hand and ensure the variable names are aligned with

1)  the question library

2)  list of calculated variables

3)  Household demographic variables.

Ensure that all variables part of the cleaned data or dashboard variable names are described in the codebook.

```{r, include = FALSE}
variable <- ls(Data)
data <- data.frame(variable) %>%
  filter(
    !variable %in% c(survey_questions$variable, vars_transformed$variable, question_library$variable)
  )

```

# Add match - column

When you are dealing with an endline case, we create a "match" column to indicate whether or not the farmer is recurring. 
PLEASE IGNORE THIS SECTION WHEN YOU ARE DEALING WITH A BASELINE CASE! 

In the 5200 Farmfit Github folder, you can find a folder named "PDC 2.0 sampling". Look here for the pi_ file of the baseline and copy to the case folder. We load in this dataset to match with current farmers and see which ones are recurring. 


NB:If the pi_ file is not available for baseline, use the raw data and generate it. here is and example
```{r}
#Data_baseline<-read_excel("DATA_ANALYSIS-baseline-Alluvial.xlsx")
#Data_baseline <- Data_baseline %>%
#  select(-contains("--option--")) %>%
#  rename_all(funs(tolower)) %>%
#  filter(`repeat no`==1) %>%
#  select(identifier,
#         name_of_farmer,
#         mobile_number_farmer,
#         contains("pi_"), -contains("ppi")) 

#pi_info<- list("Personal information" = Data_baseline)

#write.xlsx(pi_info, file="Pi_alluvial_baseline.xlsx")
```

First load in the baseline farmers and only keep the phone numbers. 

```{r}
baseline_farmers <- read_excel("pi_aif_rwanda_baseline.xlsx")
# Only keep the phone number column for matching
baseline_farmers <- baseline_farmers %>% select(identifier, mobile_number_farmer)

# Retain only unique rows based on mobile_number_farmer
baseline_farmers <- baseline_farmers %>%
  distinct(mobile_number_farmer, .keep_all = TRUE)
```

Now we match the Data with these phone numbers. 

```{r}
Data <- Data %>% left_join(baseline_farmers, by="mobile_number_farmer")
```
And add a column to indicate the match
```{r}
# create now column to indicate whether there is a match or not 
Data$cal_match <- ifelse(is.na(Data$identifier.y), FALSE, TRUE)
# Remove and rename columns that changed due to merging
Data <- Data %>% select(-identifier.y)
colnames(Data)[colnames(Data) == "identifier.x"] <- "identifier"
```

```{r}
# Delete phone number from dataset for anonymisation
Data <- Data %>%
          select(
      -c(mobile_number_farmer)) 
```

# Combine the inputs for data delivery

```{r}

sets <- list(
    "Codebook" = survey_questions,
    "Cleaned Data" = Data,
    "Raw Data (anonymised)" = Data_raw, 
    "Num. desc. all farmers" = numerical_descriptives_all_farmers,
    "Cat. desc. single all farmers" = single_categorical_descriptives_all_farmers,
    "Cat. desc. multi all farmers" = multiple_categorical_descriptives_all_farmers,
    "Num. desc. by gender" = numerical_descriptives_by_gender,
    "Cat. desc. single by gender" = single_categorical_descriptives_by_gender,
    "Cat. desc. multi by gender" = multiple_categorical_descriptives_by_gender,
    "Net promoter score" = NSP)
  
write.xlsx (sets, file = (paste0(data_delivery)))
```

# Get personal information sheet

```{r, include = FALSE}
## ---- Extract personal information ----
Data_raw <- read_excel(data_filename)

Data_raw <- Data_raw %>%
  select(-contains("--option--")) %>%
  rename_all(funs(tolower)) %>%
  filter(`repeat no`==1) %>%
  select(identifier,
         name_of_farmer,
         mobile_number_farmer,
         contains("pi_")) #-contains("ppi")) 


pi_info<- list("Personal information" = Data_raw)

write.xlsx(pi_info, file=pi_filename) #pi_filename was created in the beginning of the script


```

# Productivity check

Create this file the first time you run the script to clean the data.  

```{r}
#Productivity check

productivity <- Data %>%
    mutate(focus_product_min_sold = cal_focus_quant_prod_kg - cal_focus_quant_sold_kg) %>%

    select(identifier,
           #submitter,
           #`submission date`, ##MAKE SURE THIS VARIABLE IS NOT DELETED IN THE FIRST PIECE OF THE CODE.It should be deleted before sharing with IDH, but it should be kept when checking the productivity numbers.
           starts_with("pi_location_cascade_level"), #include the applicable variable, this differs per case
           f_unit_land,
           f_size,
           f_size_acre,
           f_focus_crop_size,
           f_focus_crop_size_acre,
           cal_focus_productivity_acre,
           f_focus_quant_prod,
           f_focus_measurement_prod,
           cal_focus_quant_prod_kg,
           f_focus_quant_sold,
           f_focus_measurement_sold,
           cal_focus_quant_sold_kg,
           focus_product_min_sold,
           f_focus_price,
           f_focus_own_consumption,
           f_focus_own_consumption_measurement,
           f_focus_quant_lost,
           f_focus_measurement_lost
           )
  
write.xlsx (productivity, file = "productivity_check.xlsx", overwrite = TRUE)

```
