
---
title: "Farmer_list_matching"
output: html_document
date: "2023-07-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the personal information file of the PDC 1 from the case.

```{r}
baseline <- read_excel("pi_smart_logistics.xlsx")
baseline$identifier <- gsub("-", "", baseline$identifier)

full_data <- read_excel("2021-03-19_smart logistics_anom.xlsx")
full_data <- full_data %>% select(c("identifier", "hh_farmer_gender"))

baseline <- baseline %>% left_join( full_data, by="identifier")

```

```{r}
table(baseline$hh_farmer_gender)
```

Import the available farmer list for endline.

```{r, include = FALSE}
farmer_list_endline_machakos <- read_excel("IDH ENDLINE SURVEY FARMERS LIST (1).xlsx", sheet = 1)
farmer_list_endline_nakuru <- read_excel("IDH ENDLINE SURVEY FARMERS LIST (1).xlsx", sheet = 2)
```

```{r}
table(farmer_list_endline_machakos$GENDER)
```

```{r}
table(farmer_list_endline_nakuru$GENDER)
```

We continue here with Nakaru for matching.

```{r, include = FALSE}
# remove (254) from the mobile number column
farmer_list_endline_nakuru$CONTACTS <- gsub("\\(254\\)", "", farmer_list_endline_nakuru$CONTACTS)

# Rename the CONTACTS column to mobile_number_farmer
colnames(farmer_list_endline_nakuru)[colnames(farmer_list_endline_nakuru) == "CONTACTS"] <- "mobile_number_farmer"

```

## Matching 

We can try and match farmers with a left join function

```{r}
merged_data <- baseline %>% left_join(farmer_list_endline_nakuru, by="mobile_number_farmer")
```

```{r}
# create now column to indicate whether there is a match or not 
merged_data$match <- ifelse(is.na(merged_data$`#`), FALSE, TRUE)
```

```{r}
table(merged_data$match, merged_data$hh_farmer_gender)
```



## Match to retrieve sample = starting from farmer list 

```{r}
# Merge baseline on mobile number 
farmer_list_endline_nakuru <- farmer_list_endline_nakuru %>% left_join(baseline, by = "mobile_number_farmer")

farmer_list_endline_nakuru <- farmer_list_endline_nakuru %>% select(-c("pi_location_cascade_county", "pi_location_cascade_sub_county", "pi_location_cascade_ward","pi_location_cascade_village","pi_location_other", "pi_location_other_first_admin", "pi_location_other_second_admin", "pi_location_other_third_admin", "pi_location_other_village", "pi_geolocation_1_latitude", "pi_geolocation_1_longitude", "pi_geolocation_1_elevation","hh_farmer_gender", "name_of_farmer"))

farmer_list_endline_nakuru$match <- ifelse(is.na(farmer_list_endline_nakuru$identifier), FALSE, TRUE)
colnames(farmer_list_endline_nakuru)[colnames(farmer_list_endline_nakuru) == "identifier"] <- "match_phone"

```


Make the sample

```{r}
sample <- farmer_list_endline_nakuru[farmer_list_endline_nakuru$match == TRUE,]

```

Make random sample of 112 male farmers Nakuru

```{r}
male_nakuru <- farmer_list_endline_nakuru[(farmer_list_endline_nakuru$GENDER=="Male" & farmer_list_endline_nakuru$match == FALSE),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 112

# Get the total number of rows in the dataframe
total_rows <- nrow(male_nakuru)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
male_nakuru_sample <- male_nakuru[random_row_indices, ]
```

Make random sample of 109 female farmers from Nakuru

```{r}
female_nakuru <- farmer_list_endline_nakuru[(farmer_list_endline_nakuru$GENDER=="Female" & farmer_list_endline_nakuru$match == FALSE),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 109

# Get the total number of rows in the dataframe
total_rows <- nrow(female_nakuru)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
female_nakuru_sample <- female_nakuru[random_row_indices, ]
```

Make random sample of 6 male farmers from Machakos

```{r}
#rename column for matching later
colnames(farmer_list_endline_machakos)[colnames(farmer_list_endline_machakos) == "CONTACTS"] <- "mobile_number_farmer"

```

```{r}
male_machakos <- farmer_list_endline_machakos[(farmer_list_endline_machakos$GENDER=="Male"),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 6

# Get the total number of rows in the dataframe
total_rows <- nrow(male_machakos)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
male_machakos_sample <- male_machakos[random_row_indices, ]

male_machakos_sample$match_phone <- NA
male_machakos_sample$match <- FALSE

```

Make random sample of 40 female farmers from Machakos

```{r}
female_machakos <- farmer_list_endline_machakos[(farmer_list_endline_machakos$GENDER=="Female"),]

# Set the seed for reproducibility (optional)
set.seed(123)

# Number of rows to randomly pick
n_rows_to_pick <- 40

# Get the total number of rows in the dataframe
total_rows <- nrow(female_machakos)

# Randomly select 51 row indices without replacement
random_row_indices <- sample(1:total_rows, n_rows_to_pick, replace = FALSE)

# Extract the randomly selected rows from the dataframe
female_machakos_sample <- female_machakos[random_row_indices, ]

female_machakos_sample$match_phone <- NA
female_machakos_sample$match <- FALSE
```

Merge the different random sets with the recurring farmers 

```{r}
sample <- rbind(sample, male_nakuru_sample, female_nakuru_sample, male_machakos_sample, female_machakos_sample)
```

Export the full farmer list and the sample.

```{r}
#full farmer list
farmer_list_endline_nakuru_2 <- farmer_list_endline_nakuru %>% select(-c(match_phone, match))
full_list <- rbind(farmer_list_endline_nakuru_2, farmer_list_endline_machakos)
```

```{r}
write.xlsx(full_list, "Farmer_list_smart_logistics_endline.xlsx")
write.xlsx(sample, "sample_smart_logistics_endline.xlsx")
```

